---
title: "Relative Deprivation in `r params$ward_name`"
always_allow_html: true
output:
  html_document:
    df_print: paged
    includes:
      in_header: rmd_report_includes.html
    mathjax: null
    self_contained: yes
    theme: flatly
    toc: yes
    toc_depth: 2
    toc_float: yes
  word_document:
    toc: no
date: "Last updated: `r format(Sys.time(), '%d %B %Y')`"
params:
  ward_name: Bucklow-St Martins
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE)

# Load the required libraries
library(tidyverse) ; library(ggplot2) ; library(ggspatial) ; library(ggthemes) ; library(ggrepel) ; library(sf) ; library(kableExtra) ; library(english) ; library(janitor) ; library(scales)


# Load in the ggplot2 Lab theme
source("https://www.trafforddatalab.io/assets/theme/ggplot2/theme_lab.R")


# Set up the chart highlight colour
chart_highlight <- "#18BC9E"


# Function to draw barcode plots
barcode_plot <- function (domain) {
  ggplot() +
  geom_segment(data = filter(df_iod_trafford, !lsoa11cd %in% lsoas_ward, index_domain == domain), 
               aes(x = rank, xend = rank, y = 0, yend = 1), 
               colour = "#bdbdbd", size = 0.25, alpha = 1) +
  geom_segment(data = filter(df_iod_trafford, lsoa11cd %in% lsoas_ward, index_domain == domain), 
               aes(x = rank, xend = rank, y = 0, yend = 1), 
               colour = chart_highlight, size = 0.25, alpha = 1) +
  geom_point(data = filter(df_iod_trafford, lsoa11cd %in% lsoas_ward, index_domain == domain),
               aes(x = rank, y = 1.1), pch = 6, colour = chart_highlight, size = 0.25) +
  geom_point(aes(x = c(0, 32844), y = -0.05), pch = 15, colour = "#bdbdbd", size = 1) +
  scale_x_continuous(limits = c(0, 32844), expand = c(0.005, 0.005), 
                     breaks = c(3284, 6569, 9853, 13138, 16422, 19706, 22991, 26275, 29560, 32844),
                     labels = c(1:10)) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL,
      tag = "Each line represents an LSOA") +
  theme_lab() +
  theme(panel.grid.major = element_blank(),
        plot.title = element_text(size = 26, face = "plain", hjust = 0.5, margin = margin(10, 0, 10, 0)),
        plot.subtitle = element_text(hjust = 0.5, vjust = 5),
        axis.text.x = element_text(face = "bold"),
        axis.text.y = element_blank(),
        legend.position = "none",
        plot.tag.position = c(0.01, -0.05),
        plot.tag = element_text(size = 10, colour = "#757575", hjust = 0))
}


# Function to return numbers expressed as words if the number of words is 2 or less, or the original number. This is for academic non-technical report writing. Uses packages: english (John Fox, Bill Venables, Anthony Damico and Anne Pier Salverda), stringr (Hadley Wickham)
number_for_reports <- function (num) {
  # Convert the input to words
  result <- words(num) # NOTE: very important to use 'words(n)' not 'english(n)' in R markdown as 'words' returns type character. Using 'english' doesn't work and simply returns the number - see: https://cran.r-project.org/web/packages/english/vignettes/the-english-patient.pdf
  
  # We only want to return numbers as words if they can be expressed with no more than 2 words
  if ((str_count(result, " ") + 1) > 2) {
    result <- num
  }
  else if (num < 100) {
    result <- str_replace(result, " ", "-") # replace a space if it exists with a hyphen if the number is below 100
  }
  
  result  # return the result
}


# Get the LSOAs which are best-fit within our chosen ward as a vector
lsoas_ward <- read_csv("https://www.trafforddatalab.io/spatial_data/lookups/lsoa_to_ward_best-fit_lookup.csv") %>%
  filter( lad17nm == 'Trafford',
          wd17nm == params$ward_name) %>%
  .$lsoa11cd

# Get the spatial data for the LSOAs within our chosen ward
sf_lsoas_ward <- st_read("https://www.trafforddatalab.io/spatial_data/lsoa/2011/trafford_lsoa_generalised.geojson") %>%
  filter(area_code %in% lsoas_ward) %>%
  # Create new variables for the lon & lat coordinates of the centroids of the LSOAs
  mutate(lon = map_dbl(geometry, ~st_centroid(.x)[[1]]),
         lat = map_dbl(geometry, ~st_centroid(.x)[[2]]))


# Get the spatial data for the boundary of our chosen ward
sf_ward <- st_read("https://www.trafforddatalab.io/spatial_data/ward/2017/trafford_ward_generalised.geojson") %>%
  filter(area_name == params$ward_name)


# English Indices of Deprivation 2019 #
# Source: Ministry of Housing, Communities and Local Government
# Publisher URL: https://www.gov.uk/government/statistics/announcements/english-indices-of-deprivation-2019
# Licence: Open Government Licence 3.0
df_iod_trafford <- read_csv("https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/833982/File_7_-_All_IoD2019_Scores__Ranks__Deciles_and_Population_Denominators.csv") %>% 
  clean_names() %>% 
  filter(local_authority_district_name_2019 == "Trafford") %>% 
  select(lsoa11cd = 1, lad19nm = 4, 5:34) %>% 
  gather(variable, value, -lsoa11cd, -lad19nm) %>% 
  mutate(measure = case_when(str_detect(variable, "score") ~ "score", 
                             str_detect(variable, "decile") ~ "decile", 
                             str_detect(variable, "rank") ~ "rank"),
         index_domain = case_when(str_detect(variable, "index_of_multiple_deprivation") ~ "Index of Multiple Deprivation", 
                                  str_detect(variable, "employment") ~ "Employment",
                                  str_detect(variable, "education") ~ "Education",
                                  str_detect(variable, "health") ~ "Health",
                                  str_detect(variable, "crime") ~ "Crime",
                                  str_detect(variable, "barriers") ~ "Barriers to Housing and Services",
                                  str_detect(variable, "living") ~ "Living Environment",
                                  str_detect(variable, "idaci") ~ "Income Deprivation Affecting Children",
                                  str_detect(variable, "idaopi") ~ "Income Deprivation Affecting Older People",
                                  TRUE ~ "Income"),
         index_domain = fct_relevel(factor(index_domain), 
                                    "Index of Multiple Deprivation", 
                                    "Income", "Employment", "Education", "Health", "Crime",
                                    "Barriers to Housing and Services", "Living Environment",
                                    "Income Deprivation Affecting Older People",
                                    "Income Deprivation Affecting Children")) %>% 
  select(lsoa11cd, lad19nm, measure, value, index_domain) %>% 
  spread(measure, value) %>% 
  mutate(year = "2019")
```

This report explores the [2019 English Indices of Deprivation](https://www.gov.uk/government/statistics/english-indices-of-deprivation-2019) for the small areas (Lower-Layer Super Output Areas or LSOAs) within `r params$ward_name`. It compares these small areas relative to Trafford using the rank and score and highlights the indicators and data sources behind these measures.

## Key findings

TO BE COMPLETED

## Background
The [English Indices of Deprivation](https://www.gov.uk/government/statistics/english-indices-of-deprivation-2019) (IoD) were published by the [Ministry of Housing, Communities & Local Government](https://www.gov.uk/government/organisations/ministry-of-housing-communities-and-local-government) in September 2019. The Indices measure relative levels of deprivation in 32,844 Lower-layer Super Output Areas (LSOAs) in England. Each LSOA contains approximately 1,500 residents or 650 households. They are ranked from 1 (most deprived) to 32,844 (least deprived) and assigned to a decile ranging from 1 (most deprived 10%) to 10 (least deprived 10%). Local authorities in England are also ranked from 1 to 317 based on a population weighted average rank or score of the LSOA data.

Relative deprivation is measured using 39 indicators which are grouped across seven different domains:

* **Income** - the proportion of the population experiencing deprivation relating to low income, including those people-out-of-work, and those who have low earnings    
* **Employment** - the proportion of the working age population in an area involuntarily excluded from the labour market     
* **Education, Skills and Training** - the lack of attainment and skills in the local population   
* **Health and Disability** - the risk of premature death and the impairment of quality of life through poor physical or mental health     
* **Crime** - the risk of personal and material victimisation at a local level  
* **Barriers to Housing and Services** - the physical and financial accessibility of housing and local services     
* **Living Environment** - the quality of the local environment    

Each domain is combined and weighted to create the **Index of Multiple Deprivation** (IMD), an overall measure of relative deprivation. 

A further 2 supplementary indices of income deprivation are also provided:    

* **Income Deprivation Affecting Children Index (IDACI)** - the proportion of all children aged 0 to 15 living in income deprived families          
* **Income Deprivation Affecting Older People Index (IDAOPI)** - the proportion of all those aged 60 or over who experience income deprivation

#### Please note      
* The Indices of Deprivation measure *relative* deprivation so a change in rank between 2019 and 2015 releases does not necessarily indicate a change in *absolute* levels of deprivation.
* The Indices of Deprivation cannot be used to quantify how deprived or affluent an area is, identify deprived people or to measure an absolute change in deprivation over time. However, it can be used to look at changes in relative deprivation between iterations of the data and to identify the most deprived small areas.
* The 39 indicators are based on data from the most recent time point available. This means that there is not one consistent time point for all the data. Although published in 2019, the latest Indices of Deprivation data are based largely on administrative data from 2015 and 2016. Similarly the 2015 publication is mainly based on data from 2012. Therefore there may be recent developments which are not yet reflected in the information published.
* The score for the Income and Employment domains as well as the two sub-domains for Income (IDACI AND IDAOPI) directly relates to the percentage of the relevant population experiencing that type of deprivation. See [The English Indices of Deprivation 2019 Research Report](https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/833947/IoD2019_Research_Report.pdf) for more information.

#### Further resources
* Ward level summaries of the 2019 Indices of Deprivation for Trafford are available to view interactively in linked charts and maps [here](https://www.trafforddatalab.io/ward_profiler/).
* Relative deprivation in Trafford compared to other Local Authorities within Greater Manchester [report](https://www.trafforddatalab.io/analysis/IoD2019/report/).
* A web application which interactively visualises the proportion of LSOAs in the most deprived decile and the spatial distribution of relative deprivation for every local authority district in England is available [here](https://trafforddatalab.shinyapps.io/imd19).

#### Relationship between LSOAs and electoral wards
Ward boundaries and LSOA boundaries do not always perfectly align and so it is impossible to aggregate from LSOA to ward exactly. Instead a 'best-fit' methodology is used by the [Office for National Statistics (ONS)](http://geoportal.statistics.gov.uk/datasets/500d4283cbe54e3fa7f358399ba3783e_0) to assign LSOAs to the most appropriate ward. This may mean that some parts of the LSOAs discussed within this report fall outside of the ward boundary and some areas within the boundary are not covered as they have been assigned to another ward. Using this method, the ward of `r params$ward_name` encompasses the following `r length(lsoas_ward)` LSOAs:

```{r lsoa_map,  echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.align="center", fig.cap="© CARTO. Contains Ordnance Survey data © Crown copyright and database right 2019"}
ggplot() +
  # Add map tiles in background
  annotation_map_tile(type = "cartodark") +
  # Add the ward boundary
  geom_sf(data = sf_ward, fill = NA, alpha = 0.5, colour = "#2C3E50", size = 1.5) +
  # Add the LSOA boundaries
  geom_sf(data = sf_lsoas_ward, fill = NA, alpha = 1, colour = "#18BC9E", size = 0.6) +
  # Add a point to the centroids
  geom_sf(data = sf_lsoas_ward %>% # pipe the dataset into the data attribute
            st_drop_geometry() %>% # remove the current geometry...
            select(lon, lat) %>%   # ...and all other fields except lon and lat and then...
            st_as_sf(coords = c("lon", "lat")) %>% # ...create a new geometry from them
            st_set_crs(4326), # set the correct coordinate reference system
          colour = "#18BC9E", fill = "#ffffff", shape = 21, size = 1, stroke = 0.5) +
  # Add the LSOA names to the centroids
  geom_text_repel(data = sf_lsoas_ward, aes(x = lon, y = lat, label = area_name), colour = "#FFFFFF", size = 3, fontface = "bold") +
  # Remove all other plot labels
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL) +
  # Set the coordinate reference system
  coord_sf(crs = st_crs(4326), datum = NA) +
  # Apply the themes
  theme_void(base_family = "Open Sans") +
  theme_map() +
  theme(legend.position = "none",
        panel.background = element_blank(),
        axis.text = element_blank())
```

```{r lsoa_table, echo=FALSE, message=FALSE, warning=FALSE}

# Create a table showing the LSOAs within the ward
df_lsoa_codes_and_names <- read_csv("https://www.trafforddatalab.io/spatial_data/lookups/statistical_lookup.csv") %>%
  filter(lsoa11cd %in% lsoas_ward) %>%
  select(lsoa11cd, lsoa11nm) %>%
  distinct(lsoa11cd, .keep_all = TRUE) %>%
  arrange(lsoa11cd) %>%
  rename("LSOA code" = lsoa11cd,
         "LSOA name" = lsoa11nm)

kable(df_lsoa_codes_and_names) %>%
  kable_styling(bootstrap_options = "striped")
```

## Index of Multiple Deprivation
As mentioned above, the 7 domains are combined and weighted to create the Index of Multiple Deprivation which is an overall measure of relative deprivation. The plot below shows where the `r length(lsoas_ward)` LSOAs within `r params$ward_name` lie relative to all the LSOAs in Trafford, and in which decile.

<h4 class="plotTitle">LSOAs in `r params$ward_name` relative to Trafford on IMD</h4>
```{r barcode_imd, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.align='center', fig.cap="Source: 2019 Indices of Deprivation, MHCLG"}
barcode_plot(domain = "Index of Multiple Deprivation")
```

TO BE COMPLETED

## Income
The definition of low income used includes both those people that are out-of-work, and those that are in work but who have low earnings (and who satisfy the respective means tests). A combined count of income deprived individuals is produced by summing the following non-overlapping indicators:

* Adults and children in Income Support families
* Adults and children in income-based Jobseeker's Allowance families
* Adults and children in income-based Employment and Support Allowance families
* Adults and children in Pension Credit (Guarantee) families
* Adults and children in Working Tax Credit and Child Tax Credit families not already counted
* Adults and children in Universal Credit families where no adult is in 'Working - requirements' conditionality regime
* Asylum seekers in England in receipt of subsistence support, accommodation support, or both

This domain has a weighting of 22.5% when combined with the others to form the Index of Multiple Deprivation 2019.

<h4 class="plotTitle">LSOAs in `r params$ward_name` relative to Trafford on Income</h4>
```{r barcode_income, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.align='center', fig.cap="Source: 2019 Indices of Deprivation, MHCLG"}
barcode_plot(domain = "Income")
```

TO BE COMPLETED

## Employment
This domain measures the proportion of working age population who are involuntarily excluded from the labour market, including those who would like to work but are unable to do so due to unemployment, sickness or disability, or caring responsibilities. This comprises the following non-overlapping indicators:

* Claimants of Jobseeker's Allowance (both contribution-based and income-based), women aged 18 to 59 and men aged 18 to 64
* Claimants of Employment and Support Allowance, women aged 18 to 59 and men aged 18 to 64
* Claimants of Incapacity Benefit, women aged 18 to 59 and men aged 18 to 64
* Claimants of Severe Disablement Allowance, women aged 18 to 59 and men aged 18 to 64
* Claimants of Carer's Allowance, women aged 18 to 59 and men aged 18 to 64
* Claimants of Universal Credit in the 'Searching for work' and 'No work requirements' conditionality groups

This domain has a weighting of 22.5% when combined with the others to form the Index of Multiple Deprivation 2019.

<h4 class="plotTitle">LSOAs in `r params$ward_name` relative to Trafford on Employment</h4>
```{r barcode_employment, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.align='center', fig.cap="Source: 2019 Indices of Deprivation, MHCLG"}
barcode_plot(domain = "Employment")
```

TO BE COMPLETED

## Education, Skills and Training
The indicators within this domain fall into two sub-domains: Children and Young People and Adult Skills. The Children and Young People sub-domain comprises:

* Key Stage 2 attainment
* Key Stage 4 attainment
* Secondary school absence
* Staying on in education post 16
* Entry to higher education

The Adult Skills sub-domain is a non-overlapping count of the following two indicators:

* The proportion of working age adults with no or low qualifications (women aged 25 to 59 and men aged 25 to 64)
* The proportion of working adults who cannot speak English or cannot speak English well (women aged 25 to 59 and men aged 25 to 64)

This domain has a weighting of 13.5% when combined with the others to form the Index of Multiple Deprivation 2019.

<h4 class="plotTitle">LSOAs in `r params$ward_name` relative to Trafford on Education</h4>
```{r barcode_education, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.align='center', fig.cap="Source: 2019 Indices of Deprivation, MHCLG"}
barcode_plot(domain = "Education")
```

TO BE COMPLETED

## Health and Disability
Measuring the risk of premature death and the impairment of quality of life through poor physical or mental health through the following four indicators:

* Years of potential life lost
* Comparative illness and disability ratio
* Acute morbidity
* Mood and anxiety disorders

This domain has a weighting of 13.5% when combined with the others to form the Index of Multiple Deprivation 2019.

<h4 class="plotTitle">LSOAs in `r params$ward_name` relative to Trafford on Health</h4>
```{r barcode_health, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.align='center', fig.cap="Source: 2019 Indices of Deprivation, MHCLG"}
barcode_plot(domain = "Health")
```

TO BE COMPLETED

## Crime
This domain measures the rate of recorded crimes per 1,000 at risk population for:

* Violence: number of recorded violent crimes (18 recorded crime types in 2016/17; 20 recorded crime types in 2017/18)
* Burglary: number of recorded burglaries (4 recorded crime types)
* Theft: number of recorded thefts (5 recorded crime types)
* Criminal damage: number of recorded crimes (8 recorded crime types)

This domain has a weighting of 9.3% when combined with the others to form the Index of Multiple Deprivation 2019.

<h4 class="plotTitle">LSOAs in `r params$ward_name` relative to Trafford on Crime</h4>
```{r barcode_crime, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.align='center', fig.cap="Source: 2019 Indices of Deprivation, MHCLG"}
barcode_plot(domain = "Crime")
```

TO BE COMPLETED

## Barriers to Housing and Services
The indicators which comprise this domain are split into two sub-domains: Geographical Barriers and Wider Barriers. Geographical Barriers takes into account road distances to a:

* Post Office
* Primary School
* General Store
* Supermarket

Wider Barriers takes into account the following:

* Household overcrowding
* Homelessness
* Housing affordability

This domain has a weighting of 9.3% when combined with the others to form the Index of Multiple Deprivation 2019.

<h4 class="plotTitle">LSOAs in `r params$ward_name` relative to Trafford on Barriers to Housing & Services</h4>
```{r barcode_barriers, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.align='center', fig.cap="Source: 2019 Indices of Deprivation, MHCLG"}
barcode_plot(domain = "Barriers to Housing and Services")
```

TO BE COMPLETED

## Living Environment
This domain is also derived from two sub-domains: Indoors and Outdoors.  

Indoors sub-domain:

* Houses without central heating
* Housing in poor condition: Social and private homes failing to meet the [Decent Homes standard](https://www.gov.uk/government/publications/a-decent-home-definition-and-guidance).

Outdoors sub-domain:

* Air quality: A measure of air quality based on emissions rates for four pollutants
* Road traffic accidents involving injury to pedestrians and cyclists: A measure of road traffic accidents involving injury to pedestrians and cyclists among the resident and workplace population

This domain has a weighting of 9.3% when combined with the others to form the Index of Multiple Deprivation 2019.

<h4 class="plotTitle">LSOAs in `r params$ward_name` relative to Trafford on Living Environment</h4>
```{r barcode_living, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.align='center', fig.cap="Source: 2019 Indices of Deprivation, MHCLG"}
barcode_plot(domain = "Living Environment")
```

TO BE COMPLETED

## Income Deprivation Affecting Children
The proportion of children aged 0 to 15 living in income deprived households.

<h4 class="plotTitle">LSOAs in `r params$ward_name` relative to Trafford on IDACI</h4>
```{r barcode_idaci, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.align='center', fig.cap="Source: 2019 Indices of Deprivation, MHCLG"}
barcode_plot(domain = "Income Deprivation Affecting Children")
```

TO BE COMPLETED

## Income Deprivation Affecting Older People
The proportion of people aged 60 and over living in income deprived households.

<h4 class="plotTitle">LSOAs in `r params$ward_name` relative to Trafford on IDAOPI</h4>
```{r barcode_idaopi, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.align='center', fig.cap="Source: 2019 Indices of Deprivation, MHCLG"}
barcode_plot(domain = "Income Deprivation Affecting Older People")
```

TO BE COMPLETED