---
title: "Claimant Count and Universal Credit in Trafford"
lang: "en-GB"
output: html_document
---

```{r setup, include=FALSE, results="hide"}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse) ; library(httr) ; library(readxl) ; library(lubridate)
library(zoo) ;library(sf) ; library(jsonlite) ; library(scales)
library(reactable) ; library(htmltools) ; library(ggrepel)

```


```{r CCdata}

base <- "Jan 2020"
basech <- "January 2020"

#Claimant Count
# Source: ONS
# URL: https://www.nomisweb.co.uk/sources/cc
# Licence: Open Government Licence

cc_trafford <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_162_1.data.csv?geography=1811939363&date=latestMINUS41-latest&gender=0&age=0&measure=1,2&measures=20100") %>%
    select(period = DATE_NAME,
         area_code = GEOGRAPHY_CODE,
         area_name = GEOGRAPHY_NAME,
         measure = MEASURE_NAME,
         value = OBS_VALUE)

latest <- as.character(max(as.yearmon(cc_trafford$period)))
latestm12 <- as.character(max(as.yearmon(cc_trafford$period))-1)

latestch <- "June 2023"
latestm12ch <- "June 2022"

cc_traf <- cc_trafford %>%
  mutate(period=as.yearmon(period, "%b %Y"),
         measure=ifelse(measure=="Claimant count","count","rate")) %>%
  select(-area_code) %>%
  spread(measure,value) %>%
  mutate(change=round((count*100/count[1])-100,1))

```

<main>
<section>

# Claimant Count in Trafford

The [Claimant Count](https://www.nomisweb.co.uk/sources/cc) indicates the number of people claiming benefits principally for the reason of being unemployed. The Claimant Count in `r latestch` was `r filter(cc_traf,period==as.yearmon(latest, "%b %Y"))$change`% higher than in `r basech`. The rate of claimants to residents aged 16 to 64 reached `r filter(cc_traf,period==as.yearmon(latest, "%b %Y"))$rate`% in `r latestch` representing an increment of `r filter(cc_traf,period==as.yearmon(latest, "%b %Y"))$rate-cc_traf$rate[1]`% from `r basech`. The number of claimants has been decreasing from April 2021 but started to increase again from January 2023.


 <div class="container-fluid">
  <div class="row">
   <div class = "col-sm-4">

```{r CCTraffPCHTable}

table1 <- cc_traf %>% 
  mutate(period = as.Date(period,frac=0)) %>%
  select(-area_name)

tbl1 <- reactable(
  table1,
  pagination = TRUE,
  defaultSorted = "period",
  defaultSortOrder = "desc",
  fullWidth = FALSE,
  defaultColDef = colDef(headerClass = "header", align = "right", minWidth = 60, format = colFormat(separators = TRUE)),
  columns = list(
    period = colDef(name = "Date", align = "left", cell = function(value) date_format("%b %Y")(value)),
   # area_name = colDef(name = "Area", align = "left"),
    change = colDef(name = "Change %",
                    format = colFormat(digits = 1))
    ),
  compact = TRUE,
  class = "table"
)

div(class = "div",
div(class= "tableTitle", "Change in the Claimant Count for Trafford compared to January 2020"),
tbl1
)

```

   </div>
  
   <div class = "col-sm-8">
```{r CCTrafPCHPLineP}

cc_traffordp <- cc_traf %>% mutate(date=as.Date(period,frac=0)) %>%
  mutate(change = round(change,0)) %>%
  mutate(index = case_when(change == 0 ~ "Label",
                           period == latest ~ "Label",
                           change -lag(change)<=0 & change - lead(change)<0 ~ "Label",
                           change -lag(change)>=0 & change - lead(change)>0 ~ "Label",
                           change -lead(change)<(-20) | change - lag(change)>20 ~ "Label",
                           TRUE ~ "No label"))

ggplot(cc_traffordp,aes(x = date, y = round(change,0))) + 
  geom_line(aes(group=1), size = 1, alpha = 1, color="#006d2c") + #"#66c2a4", "#006d2c" dark
  geom_point(size = 3, alpha = 1, color="#006d2c") + 
    geom_text(data=cc_traffordp,
            aes(label = paste0(change, "%"), color = index),
            fontface = 1.5, size = 2.5, hjust = 0.3, vjust = -1.3)+
  scale_color_manual(values = c("#006d2c", "transparent")) +
  scale_x_date(date_breaks="2 month", date_labels="%b %y", expand = c(0.03,0.03)) +
  scale_y_continuous(limits = c(0,135)) +
  labs(title = "Claimant Count percentage change",
       subtitle = paste("Trafford, ", base, " - ", latest),
       caption = "Source: ONS",
       x = NULL,
       y = NULL,
       colour = NULL) +
  #theme_lab() +
  theme_minimal() +
  theme(plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        panel.grid.minor = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.text.y = element_blank(),
        legend.position = "none")

```

   </div>
  </div>
 
```{r CCMSOAData}
#Claimant Count
# Source: ONS
# URL: https://www.nomisweb.co.uk/sources/cc
# Licence: Open Government Licence

 #House of Commons Library MSOA Names
# URL: https://visual.parliament.uk/msoanames
lookup <- read_csv("https://houseofcommonslibrary.github.io/msoanames/MSOA-Names-Latest.csv") %>%
  filter(Laname=="Trafford")

# Mid-2020 population estimates by ward
# Source: Nomis / ONS
# URL: https://www.nomisweb.co.uk/datasets/pestsyoaoa
# Licence: Open Government Licence 3.0

msoa_pop <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_2010_1.data.csv?geography=1245709510...1245709537&date=latest&gender=0&c_age=203&measures=20100") %>%
    select(area_code = GEOGRAPHY_CODE,
         pop16_64 = OBS_VALUE)

# claimant_ward <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_162_1.data.csv?geography=1660945005...1660945019,1660945021,1660945020,1660945022...1660945025&date=latestMINUS12,latest&gender=0&age=0&measure=1,2&measures=20100")
#https://www.nomisweb.co.uk/api/v01/dataset/NM_162_1.data.csv?geography=1245709510...1245709537&date=latest&gender=0&age=0&measure=1&measures=20100

claimant_msoa <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_162_1.data.csv?geography=1245709510...1245709537&date=latestMINUS12,latest&gender=0&age=0&measure=1&measures=20100")

claimant_area <- claimant_msoa %>%
  select(period = DATE_NAME,
         area_code = GEOGRAPHY_CODE,
         area_code2 = GEOGRAPHY_NAME,
         measure = MEASURE_NAME,
         value = OBS_VALUE) %>%
  mutate(period = as.yearmon(period, "%b %Y")) %>%
  left_join(lookup%>%select(msoa11cd,msoa11nm,msoa11hclnm), by = c("area_code2" = "msoa11nm")) %>%
  rename(area_name = msoa11hclnm)

c_count <- claimant_area %>%
  spread(period,value) %>%
  mutate(additional_claims= get(latest) - get(latestm12),
         `Share` = round(get(latest)/sum(get(latest)) *100,1),
         percentage_change = round(additional_claims/get(latestm12) *100,0))

c_rate <- c_count %>%
    left_join(msoa_pop, by="area_code") %>%
    mutate(latestm12_r = round(get(latestm12)*100/pop16_64,1), latest_r=round(get(latest)*100/pop16_64,1),Change=round(latest_r-latestm12_r,1))
  

```
 
The Trafford MSOA with the largest share of claims in `r latestch` was `r filter(c_count, Share == max(Share)) %>% pull(area_name)` with `r filter(c_count, Share == max(Share))$Share`% of all claimants within Trafford, almost the double of claims than Gorse Hill, the MSOA with the second largest number of claims. 8 MSOAs out of 28 have more claims in `r latestch`  compared to `r latestm12ch`. `r arrange(c_count, desc(additional_claims))$area_name[1]` have the largest increase in number of claims with `r abs(arrange(c_count, desc(additional_claims))$additional_claims[1])` more, followed by `r arrange(c_count, desc(additional_claims))$area_name[2]` (`r arrange(c_count, desc(additional_claims))$additional_claims[2]`), `r arrange(c_count, desc(additional_claims))$area_name[3]` (`r arrange(c_count, desc(additional_claims))$additional_claims[3]`) and  `r arrange(c_count, desc(additional_claims))$area_name[4]` (`r arrange(c_count, desc(additional_claims))$additional_claims[4]`). `r arrange(c_count, additional_claims)$area_name[1]` and `r arrange(c_count, additional_claims)$area_name[2]` have the largest reduction in number of claims with `r abs(arrange(c_count, additional_claims)$additional_claims[1])` less, followed by `r arrange(c_count, additional_claims)$area_name[3]` (`r arrange(c_count, additional_claims)$additional_claims[3]`).


```{r CCMSOATable}

table2 <- c_rate %>% rename(latestm12_c=latestm12, latest_c=latest) %>%
  select(-area_code2, -measure, -msoa11cd, -pop16_64, -area_code)

cell_function <- function(value, colTable){
        width_neg_tot <- ifelse(min(colTable, na.rm = TRUE) < 0,abs(min(colTable, na.rm = TRUE)),0)
        width_pos_tot <- ifelse(max(colTable, na.rm = TRUE) >= 0,max(colTable, na.rm = TRUE),0)
        width_bar_tot <- width_neg_tot + width_pos_tot
        width_neg_percent <- paste0(width_neg_tot * 100 /width_bar_tot, "%")
        width_pos_percent <- paste0(width_pos_tot * 100 /width_bar_tot, "%")
        width_neg <- ifelse(value < 0, paste0(abs(value) * 100 /width_neg_tot, "%"),0) 
        width_pos <- ifelse(value >= 0, paste0(value * 100 / width_pos_tot, "%"),0)
        value <- format(value, big.mark = ",")
        value <- format(value, width = 9, justify = "right")
        bar <- div(
          class = "bar-chart",
          style = list(marginRight = "6px", marginLeft = "6px"),
          div(class = "bar-neg-container",style = list(width = width_neg_percent, borderRight = "1px solid rgba(0, 0, 0, 0.2)"),
            div(class = "bar", style = list(width = width_neg, backgroundColor = "#66c2a4"))),
          div(class = "bar-pos-container",style = list(width = width_pos_percent, borderLeft = "1px solid rgba(0, 0, 0, 0.2)"),
            div(class = "bar", style = list(width = width_pos, backgroundColor = "#006d2c")))
        )
        div(class = "bar-cell", span(class = "number", value), bar)
}

tbl2 <- reactable(
  table2,
  pagination = FALSE,
  defaultSorted = "area_name",
  defaultSortOrder = "asc",
  defaultColDef = colDef(headerClass = "header", align = "left", format = colFormat(separators = TRUE)),
  columns = list(
    area_name = colDef(name = "MSOA", align = "left"),
    latestm12_c = colDef(name = latestm12, maxWidth = 60, align = "right"),
    latest_c = colDef(name = latest, maxWidth = 60, align = "right"),
    additional_claims = colDef(
      name = "Additional claims",
      cell = function(value) {
        cell_function(value,table2$additional_claims)
      }
    ),
    Share = colDef(name = "Share Jun 23", maxWidth = 80, align = "left"),
    percentage_change = colDef(name = "% Change", maxWidth = 80, align = "right"),
    latestm12_r = colDef(name = latestm12, maxWidth = 60, align = "left"),
    latest_r = colDef(name = latest, maxWidth = 60, align = "left"),
    `Change` = colDef(
        cell = function(value) {
          cell_function(value,table2$`Change`)
      }
    )
    ),
  columnGroups = list(
    colGroup(name = "Count", columns = c("latestm12_c", "latest_c", "additional_claims", "Share", "percentage_change")),
    colGroup(name = "Rate", columns = c("latestm12_r", "latest_r", "Change"))
  ),
  compact = TRUE,
  class = "large-table"
)

div(class = "div",
div(class= "large-table tableTitle", "Change in the Claimant Count for Trafford's MSOAs"),
tbl2
)


```

The MSOAs with higher claimant rate in `r latestch` were `r arrange(table2, desc(latest_r))$area_name[1]` with `r arrange(table2, desc(latest_r))$latest_r[1]`%,  `r arrange(table2, desc(latest_r))$area_name[2]` with `r arrange(table2, desc(latest_r))$latest_r[2]`%,  `r arrange(table2, desc(latest_r))$area_name[3]` with `r arrange(table2, desc(latest_r))$latest_r[3]`% and `r arrange(table2, desc(latest_r))$area_name[4]` with `r arrange(table2, desc(latest_r))$latest_r[4]`%. From the top 4 MSOAs with the highest rates, Old Trafford is the only one with a higher rate when comparing May 2022 to May 2023. The MSOAs where the claimant rate have decreased more from `r latestm12ch` to `r latestch` are `r arrange(table2, Change)$area_name[1]` with a decrease in rate of `r arrange(table2, Change)$Change[1]`%, `r arrange(table2, Change)$area_name[2]` with `r arrange(table2, Change)$Change[2]`%, `r arrange(table2, Change)$area_name[3]` with `r arrange(table2, Change)$Change[3]`% followed by `r arrange(table2, Change)$area_name[4]` with `r arrange(table2, Change)$Change[4]`%.

  <div class="row">
   <div class = "col-sm-6">

```{r CCMSOADot, fig.height=7}

df_traff <-  cc_trafford %>%
  mutate(period = as.yearmon(period, format="%B %Y")) %>%
  filter(period %in% c(as.yearmon(latest),as.yearmon(latestm12))) 

# df_traff_plot<- claimant_area %>%
#    mutate(period = as.yearmon(period, format="%b %Y")) %>%
#    bind_rows(df_traff %>% filter(measure == "")) %>%
#   filter(measure=="Claimants as a proportion of residents aged 16-64") %>%
#    mutate(period = factor(period, levels = c(latestm12,latest)),
#          area_name = fct_reorder(factor(area_name), ifelse(period == latest, value, NA), na.rm = TRUE),
#          area_name = fct_relevel(factor(area_name), "Trafford"))

cc_plot_dot <- table2 %>% 
  select (area_name,{{latestm12}}:=latestm12_r, {{latest}}:=latest_r) %>%
  gather(period,value,-one_of("area_name")) %>%
  mutate(period=as.yearmon(period, format="%B %Y")) %>%
  bind_rows(df_traff %>% filter(measure == "Claimants as a proportion of residents aged 16-64") %>% select(area_name,period,value) %>% filter(period %in% c(as.yearmon(latest),as.yearmon(latestm12))) )%>%
   mutate(period = factor(period, levels = c(latestm12,latest)),
         area_name = fct_reorder(factor(area_name), ifelse(period == latest, value, NA), na.rm = TRUE),
         area_name = fct_relevel(factor(area_name), "Trafford"))


ggplot(cc_plot_dot, aes(value,area_name)) +
  geom_line(color = "grey", size=1) +
  geom_point(aes(color = period), size = 4) +
  scale_x_continuous(labels = function(x){ paste0(x, "%") }, limits = c(0, 8)) +
  scale_colour_manual(values = c( "#66c2a4", "#006d2c")) +
  labs(x = "percentage of residents aged 16 to 64", y = NULL,
       title = "Claimant Rate", 
       subtitle = paste0("Trafford MSOAs, ", latestm12, " and ", latest), 
       caption = "Source: ONS")+
  theme_minimal() +
  theme(plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
        panel.grid.major.y = element_blank(),
        axis.text.y = element_text(hjust = 1),
        legend.position = "bottom", 
        legend.title = element_blank())

```

   </div>
  
   <div class = "col-sm-6">

```{r CCmsoaMap, fig.height=6}

c_rate_msoa <- st_read("msoa_boundaries.geojson", quiet = TRUE) %>% 
  left_join(., select(c_rate,latest_r,area_code)  , by = "area_code") 

ggplot() +
  geom_sf(data = c_rate_msoa, aes(fill = latest_r), alpha = 1, colour = "#FFFFFF",  size = 0.5) +
  scale_fill_gradientn(colours = c("#edf8fb","#b2e2e2","#66c2a4","#2ca25f","#006d2c"), 
                       na.value = "#f0f0f0",
                       guide = guide_legend(keyheight = unit(3, units = "mm"), 
                                            keywidth=unit(12, units = "mm"), 
                                            label.position = "bottom", 
                                            title = "Rate",
                                            title.position = 'top', 
                                            nrow = 1)) +
  labs(x = NULL, y = NULL, title = "Claimant Rate",
       subtitle = paste("Trafford MSOAs, ", latest),
       caption = "Source: ONS\nContains OS data © Crown copyright and database right 2020") +
  theme(plot.margin = unit(c(0.5,0,1,0), "cm"),
        plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
        axis.text = element_blank(),
        plot.caption = element_text(size=6,margin = margin(t = 55)),
        legend.position = c(0.3,-0.05),
        legend.title = element_text(size = 9),
        legend.text = element_text(size = 8),
        plot.background = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        axis.ticks = element_blank())

``` 

   </div>
  </div> 
 
```{r CCbyAgeTableData}

#Claimant Count
# Source: ONS
# URL: https://www.nomisweb.co.uk/sources/cc
# Licence: Open Government Licence

raw_cc_age <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_162_1.data.csv?geography=1811939363&date=latestMINUS41-latest&gender=0&age=0,2,10,11,3,12...16,4,17...20&measure=1&measures=20100") 


cc_age <- raw_cc_age %>%
  select(period = DATE_NAME,
         area_code = GEOGRAPHY_CODE,
         area_name = GEOGRAPHY_NAME,
         age = AGE_NAME,
         age_type = AGE_TYPE,
         value = OBS_VALUE) %>%
  mutate(period = as.yearmon(period, "%b %Y"),
         Age =  gsub("Aged |All categories: Age ", "", age)) %>%
    select(period, age_type, Age, value) %>%
  spread(period,value) %>%
  mutate(additional_claims= get(latest)-get(latestm12),
         percentage_change = round(additional_claims/get(latestm12) *100,0))

age_bands <- cc_age %>%
  filter(age_type=="Ageband") %>%
  mutate(Age=factor(Age, levels=c("16-24", "25-49", "50+", "16+"))) %>%
  arrange(Age) %>%
  select(-age_type)

quinary <- cc_age %>%
  filter(age_type=="Quinary") %>%
  mutate_if(is.numeric, list(~na_if(., Inf)))

table3 <- quinary %>% select(Age, `latestm12`= latestm12, `latest` = latest, additional_claims, percentage_change)
table4 <- age_bands %>% select(Age, `latestm12`= latestm12, `latest` = latest, additional_claims, percentage_change)

```


In Trafford from  `r latestm12ch` to `r latestch` 5 age bands has increased the number of claimants, the age band 18 to 24 (65 more), 40 to 44 (35 more), 55 to 59 (15 more) and 30 to 34 and 16 to 17 (5 more). When considering the reduction of claims the age band `r filter(table3, additional_claims == min(additional_claims)) %>% pull(Age)` years had the largest reduction with `r filter(table3, additional_claims == min(additional_claims)) %>% pull(additional_claims)` less claims.

  <div class="row">
   <div class = "col-sm-6">

```{r CCbyAgeTable}


tbl3 <- reactable(
  table3,
  pagination = FALSE,
  defaultSorted = "Age",
  defaultSortOrder = "asc",
  defaultColDef = colDef(headerClass = "header", align = "right", format = colFormat(separators = TRUE)),
  columns = list(
    Age = colDef(maxWidth = 50, align = "left"),
    latestm12 = colDef(name = latestm12, maxWidth = 60),
    latest = colDef(name = latest, maxWidth = 60),
    additional_claims = colDef(
      name = "Additional claims",
      cell = function(value) {
        cell_function(value,table3$additional_claims)
      }
    ),
    percentage_change = colDef(
      name = "% Change",
      cell = function(value) {
        cell_function(value,table3$percentage_change)
      }
    )
    ),
  compact = TRUE,
  class = "table"
)

div(class = "div",
div(class= "large-table tableTitle", "Change in the Claimant Count by age bands in Trafford"),
tbl3
)

```

   </div>
  
   <div class = "col-sm-6">

```{r CCbyAgeTable2}

tbl4 <- reactable(
  table4,
  pagination = FALSE,
  defaultSorted = "Age",
  defaultSortOrder = "asc",
  defaultColDef = colDef(headerClass = "header", align = "right", format = colFormat(separators = TRUE)),
  columns = list(
    Age = colDef(name = "Age", maxWidth = 60, align = "left"),
    latestm12 = colDef(name = latestm12, maxWidth = 60),
    latest = colDef(name = latest, maxWidth = 60),
    additional_claims = colDef(
      name = "Additional claims",
      cell = function(value) {
        cell_function(value,table4$additional_claims)
      }
    ),
    percentage_change = colDef(
      name = "% Change",
      cell = function(value) {
        cell_function(value, table4$percentage_change)
      }
    )
    ),
  compact = TRUE,
  class = "table"
)

div(class = "div",
div(class= "large-table tableTitle", "Change in the Claimant Count by broad age bands in Trafford"),
tbl4
)


```

   </div>
  </div>
 
At the start of 2020 the age band 25-34 already had the highest number of claims, however after the effects of the Covid-19 pandemic which started in March 2020 this figure increased more sharply compared to the other age bands. The claimant rate as a proportion of Trafford's population within each age band showed then a higher increase in the age bands of 16-24 and 25-34 compared to the others. From January 2023 the count and rate of all age bands started to increase but the figures are fluctuating from April 2023.

  <div class="row">
   <div class = "col-sm-6">

```{r CCAgeMLine}

plotdf <- quinary %>%
  select(-c(age_type, additional_claims, percentage_change)) %>%
  mutate(Age=fct_collapse(Age, `16-24` = c("16-17", "18-24"),
                       `25-34` = c("25-29", "30-34"),
                       `35-44` = c("35-39", "40-44"),
                       `45-54` = c("45-49", "50-54"),
                       `55+` = c("55-59", "60-64","65+"))) %>%
  group_by(Age) %>%
  summarise_all(sum) %>%
  gather(date,value, base:latest) %>%
  mutate(date=as.Date(as.yearmon(date), format="%b %Y")) %>%
  mutate(label = if_else(date == max(date), Age, factor(NA)))
  

 ggplot(plotdf, aes(x = date, y = value, colour = Age)) +
  geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +
  geom_line(size = 1, show.legend = FALSE) +
  geom_text_repel(aes(x = max(date) + 60, label = label, fontface = "bold", hjust = 1), direction = "y", show.legend = FALSE) +
  scale_x_date(breaks = c(min(plotdf$date), as.Date(as.yearmon("Apr 2020"),format="%b %Y"), as.Date(as.yearmon("Jan 2021"),format="%b %Y"), as.Date(as.yearmon("Jan 2022"),format="%b %Y"), max(plotdf$date)), date_labels = "%b %y") +
  scale_y_continuous(expand = c(0.005, 0.005)) +
  labs(title = "Claimant Count by age band",
       subtitle = paste("Trafford, ", base, " - ",latest),
       caption = "Source: ONS",
       x = NULL,
       y = NULL,
       colour = NULL) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
    plot.margin = unit(rep(0.5,4), "cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks.x = element_line(colour = "#333333", size = 0.5))  +
  coord_cartesian(clip = "off") 


```

   </div>
  
   <div class = "col-sm-6">

```{r CCAgeRate}

# Source:Mid-2019 population estimates for local authorities in England
# URL: https://www.nomisweb.co.uk/datasets/pestsyoala
# Licence: Open Government Licence v3.0


pop_trafford_age_raw <- read_csv("http://www.nomisweb.co.uk/api/v01/dataset/NM_2002_1.data.csv?geography=1811939363&date=latest&gender=0&c_age=101...191&measures=20100&select=date_name,geography_name,geography_code,gender_name,c_age_name,measures_name,obs_value,obs_status_name")

pop_trafford_age <- pop_trafford_age_raw %>%
    select(age = C_AGE_NAME, n = OBS_VALUE) %>%
    mutate(age = parse_number(age),
          ageband = cut(age,
                        breaks = c(0,15,24,34,44,54,64,120),
                        labels = c("0-15","16-24","25-34","35-44","45-54","55-64","65+"),
                        right = TRUE,
                        include.lowest = TRUE)) %>%
   group_by(ageband) %>%
   summarise(population = sum(n))

#write_csv(pop_trafford_age_raw, "pop_trafford_age_raw.csv")

plotdf2 <- quinary %>%
  select(-c(age_type, additional_claims, percentage_change)) %>%
  mutate(Age=fct_collapse(Age, `16-24` = c("16-17", "18-24"),
                       `25-34` = c("25-29", "30-34"),
                       `35-44` = c("35-39", "40-44"),
                       `45-54` = c("45-49", "50-54"),
                       `55-64` = c("55-59", "60-64"),
                       `65+` = "65+")) %>%
  filter(!Age == "65+") %>%
  group_by(Age) %>%
  summarise_all(sum) %>%
  gather(date,value, base:latest) %>%
  left_join(pop_trafford_age, by = c("Age" = "ageband")) %>%
  mutate(rate = round((value/population)*100,1),
         date=as.Date(as.yearmon(date), format="%b %Y")) %>%
  mutate(label = if_else(date == max(date), Age, factor(NA)))

 ggplot(plotdf2, aes(x = date, y = rate, colour = Age)) +
  geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +
  geom_line(size = 1, show.legend = FALSE) +
  geom_text_repel(aes(x = max(date) + 60, label = label, fontface = "bold", hjust = 1), direction = "y", show.legend = FALSE) +
  #scale_color_startrek() +
  scale_x_date(breaks = c(min(plotdf$date), as.Date(as.yearmon("Apr 2020"),format="%b %Y"), as.Date(as.yearmon("Jan 2021"),format="%b %Y"), max(plotdf$date)), date_labels = "%b %y") +
  scale_y_continuous(expand = c(0.005, 0.005)) +
  labs(title = "Claimant rate by age band",
       subtitle = paste("Trafford, ", base, " - ",latest),
       caption = "Source: ONS",
       x = NULL,
       y = NULL,
       colour = NULL) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
    plot.margin = unit(rep(0.5,4), "cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks.x = element_line(colour = "#333333", size = 0.5))  +
  coord_cartesian(clip = "off") 

```

   </div>
  </div>

More than half of Trafford's claimants are residents of 7 of the 28 MSOAs. More than half of the claimants are between 25 and 40 years.

  <div class="row">
   <div class = "col-sm-6">

```{r CCMSOAAgeTable}
#Claimant Count
# Source: ONS
# URL: https://www.nomisweb.co.uk/sources/cc
# Licence: Open Government Licence


cc_msoa_age <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_162_1.data.csv?geography=1245709510...1245709537&date=latest&gender=0&age=10...20&measure=1&measures=20100") %>%
  select(period = DATE_NAME, area_code = GEOGRAPHY_CODE, area_code2 = GEOGRAPHY_NAME, ageband = AGE_NAME, n = OBS_VALUE) %>%
    left_join(lookup%>%select(msoa11cd,msoa11nm,msoa11hclnm), by = c("area_code2" = "msoa11nm")) %>%
  rename(area_name = msoa11hclnm) %>%
  mutate(period = as.Date(paste("01", period), format = "%d %b %Y"),
         ageband=gsub("Aged ", "", ageband)) %>%
  mutate(ageband = case_when(
    ageband == "16-17" ~ "16-17",
    ageband == "18-24" ~ "18-24",
    ageband %in% c("25-29","30-34") ~ "25-34",
    ageband %in% c("35-39", "40-44") ~ "35-44",
    ageband %in% c("45-49", "50-54") ~ "45-54",
    ageband %in% c("55-59", "60-64") ~ "55-64",
    TRUE ~ "65+"
  )) %>%
  group_by(period, ageband, area_code, area_name) %>%
  summarise(n = sum(n)) %>%
  ungroup()


cc_msoa_aget<- cc_msoa_age %>%
  select(MSOA=area_name,ageband,n) %>%
  spread(ageband,n)

tbl5 <- reactable(
  cc_msoa_aget,
  pagination = FALSE,
  defaultSorted = "MSOA",
  defaultSortOrder = "asc",
  defaultColDef = colDef(headerClass = "header", align = "left", maxWidth = 45, format = colFormat(separators = TRUE)),
  columns = list(
    MSOA = colDef(maxWidth = 200, align = "left")
    ),
  class = "xsmall-table"
)

div(class = "div",
div(class= "large-table tableTitle", "Claimant Count for Trafford's MSOAs by age band"),
tbl5
)

```

   </div>
  
   <div class = "col-sm-6">

```{r CCMSOAAgeMMK, fig.height=6}

cc_msoa_agep<- cc_msoa_age %>%
  spread(ageband,n) %>%
  mutate("16-24" = `16-17`+`18-24`) %>%
  select(period:area_name,`16-24`,`25-34`:`65+`) %>%
  gather(ageband,n,`16-24`:`65+`) %>%
  group_by(period, area_code, area_name) %>%
  mutate(ageband = fct_rev(ageband),
         total = sum(n), percent = n/total) 

stats <- cc_msoa_agep %>% 
  summarise(totalm =sum(n)) %>%
  ungroup() %>%
  mutate(perM = totalm*100/sum(totalm))

stats2 <- cc_msoa_agep %>% 
  ungroup() %>%
  group_by(period, ageband) %>%
    summarise(totalm =sum(n)) %>%
  mutate(perM = totalm*100/sum(totalm))

ggplot(cc_msoa_agep, aes(x = area_name, y = percent, width = total, fill = ageband)) +
  geom_col(position = "stack", colour = NA) +
  facet_grid(~fct_reorder(area_name,total, .desc = T), scales = "free_x", space = "free_x") +
  scale_y_continuous(expand = c(0.005, 0.005), breaks = c(0, 0.5, 1), labels = percent) +
  scale_fill_brewer(palette = "Set3") +
  labs(x = NULL, y = NULL, 
       title = "Claimant Count by Trafford's MSOAs and age band",
       subtitle = as.yearmon(unique(cc_msoa_agep$period)),
       caption = "The width of the columns is proportional to the number of claimants. Source: DWP.",
       fill = NULL) +
  theme(plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
    panel.spacing.x = unit(0.01, "npc"),
    panel.grid.major = element_blank(),
    strip.text.x = element_blank(),
    legend.position = "right",
    axis.text.x = element_text(size=8, angle = 90, hjust = 1, vjust=0.5),
    panel.background = element_blank(),
    axis.ticks = element_blank())

```

   </div>
  </div>
 </div>
</section>

<section>

 <div class="container-fluid">
 
## Universal Credit claims in Trafford

```{r UCTraffPCHData}

# add your API key
api_key <- ""
# API endpoint
path <- "https://stat-xplore.dwp.gov.uk/webapi/rest/v1/table"

# Source:Mid-2019 population estimates for local authorities in England
# URL: https://www.nomisweb.co.uk/datasets/pestsyoala
# Licence: Open Government Licence v3.0

pop_trafford <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_2002_1.data.csv?geography=1811939363&date=latest&gender=0&c_age=203&measures=20100") %>%
    select(area_code = GEOGRAPHY_CODE,
         pop16_64 = OBS_VALUE)

#Universal Credit
# Source: DWP
# URL: https://stat-xplore.dwp.gov.uk/webapi/metadata/UC_Monthly/UC_Monthly.html
# Licence: Open Government Licence

query <- list(database = unbox("str:database:UC_Monthly"),
              measures = "str:count:UC_Monthly:V_F_UC_CASELOAD_FULL",
              dimensions = c("str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE",
                             "str:field:UC_Monthly:F_UC_DATE:DATE_NAME") %>% matrix(),
              recodes = list(
                `str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE` = list(
                  map = list("str:value:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE:V_C_MASTERGEOG11_LA_TO_REGION:E08000009")),
                `str:field:UC_Monthly:F_UC_DATE:DATE_NAME` = list(
                  map = as.list(paste0("str:value:UC_Monthly:F_UC_DATE:DATE_NAME:C_UC_DATE:",c(202001,202002,202003,202004,202005,202006,202007,202008,202009,202010,202011,202012,202101,202102,202103,202104,202105,202106,202107,202108,202109,202110,202111,202112,202201,202202,202203,202204,202205,202206,202207,202208,202209,202210,202211,202212,202301,202302,202303,202304,202305,202306))))
              )) %>% toJSON()
request <- POST(
  url = path,
  body = query,
  config = add_headers(APIKey = api_key),
  encode = "json")
response <- fromJSON(content(request, as = "text"), flatten = TRUE)
# extract list items and convert to a dataframe
tabnames <- response$fields$items %>% map(~.$labels %>% unlist)
values <- response$cubes[[1]]$values
dimnames(values) <- tabnames

trafford_uc <- values %>%
  as.data.frame.table(values, stringsAsFactors = FALSE) %>%
  as_tibble() %>%
  set_names(c(response$fields$label,"value")) %>%
mutate(period=as.yearmon(`Month`, "%B %Y")) %>%
  select(area_name = "National - Regional - LA - OAs",period,value) %>%
  mutate(rate= round(value*100/pop_trafford$pop16_64,1),
         change=round((value*100/value[1])-100,1))


```

The [Universal Credit](https://stat-xplore.dwp.gov.uk/webapi/metadata/UC_Monthly/UC_Monthly.html) claims in Trafford have increased `r filter(trafford_uc, period == latest) %>% pull(change)`% from `r basech` to `r latestch`. The rate of claims as a proportion of people age 16 to 64 increased form `r filter(trafford_uc, period == base) %>% pull(rate)`% to `r filter(trafford_uc, period == latest) %>% pull(rate)`% from `r basech` to `r latestch`.


  <div class="row">
   <div class = "col-sm-4">

```{r UCTrafPCHTable}

table6 <-  trafford_uc %>% 
  mutate(period = as.Date(period,frac=0)) %>%
  select(-area_name)

 tbl6 <- reactable(
  table6,
  pagination = TRUE,
  defaultSorted = "period",
  defaultSortOrder = "desc",
  defaultColDef = colDef(headerClass = "header", align = "right", minWidth = 60, format = colFormat(separators = TRUE)),
  columns = list(
    period = colDef(name = "Date", align = "left", cell = function(value) date_format("%b %Y")(value))
    #area_name = colDef(name = "Area", maxWidth = 100, align = "left")
    ),
  compact = TRUE,
  class = "table"
)

div(class = "div",
div(class= "tableTitle", "Change in the Universal Credit claims for Trafford"),
tbl6
)

```

   </div>
  
   <div class = "col-sm-8">

```{r UCTrafPCHPlot}

uc_plot_ch <- trafford_uc %>% mutate(date=as.Date(period,frac=0)) %>%
  mutate(change = round(change,0)) %>%
  mutate(index = case_when(change == 0 ~ "Label",
                           period == latest ~ "Label",
                           change -lag(change)<=0 & change - lead(change)<0 ~ "Label",
                           change -lag(change)>=0 & change - lead(change)>0 ~ "Label",
                           change -lead(change)<(-20) | change - lag(change)>20 ~ "Label",
                           TRUE ~ "No label"))

ggplot(uc_plot_ch,aes(x = date, y = change)) + 
  geom_line(aes(group=1), size = 1, alpha = 1, color="#006d2c") + 
  geom_point(size = 3, alpha = 1, color="#006d2c") + 
    geom_text(data=uc_plot_ch %>% filter(!period==base),
            aes(label = paste0(change, "%"), color = index),
            fontface = 1.5, size = 2.5, hjust = 0.3, vjust = -1.3)+
  scale_color_manual(values = c("#006d2c", "transparent")) +
  scale_x_date(date_breaks="2 month", date_labels="%b %y", expand = c(0.03,0.03)) +
  scale_y_continuous(limits = c(0,105)) +
  labs(title = "Universal Credit claims percentage change",
       subtitle = paste0("Trafford, ", base, " - ", latest),
       caption = "Source: DWP",
       x = NULL,
       y = NULL,
       colour = NULL) +
  #theme_lab() +
  theme_minimal() +
  theme(plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
        panel.grid.minor = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        axis.text.y = element_blank(),
        legend.position = "none")

```

   </div>
  </div>
 
```{r UCMSOATableData}
 



#Universal Credit
# Source: DWP
# URL: https://stat-xplore.dwp.gov.uk/webapi/metadata/UC_Monthly/UC_Monthly.html
# Licence: Open Government Licence


query <- list(database = unbox("str:database:UC_Monthly"),
              measures = "str:count:UC_Monthly:V_F_UC_CASELOAD_FULL",
              dimensions = c("str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE",
                             "str:field:UC_Monthly:F_UC_DATE:DATE_NAME") %>% matrix(),
              recodes = list(
                `str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE` = list(
                  map = as.list(paste0("str:value:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE:V_C_MASTERGEOG11_MSOA_TO_LA:E0", seq(2001259, 2001286, 1)))),
                `str:field:UC_Monthly:F_UC_DATE:DATE_NAME` = list(
                  map = as.list(paste0("str:value:UC_Monthly:F_UC_DATE:DATE_NAME:C_UC_DATE:",c(202206,202306))))
              )) %>% toJSON()
request <- POST(
  url = path,
  body = query,
  config = add_headers(APIKey = api_key),
  encode = "json")
response <- fromJSON(content(request, as = "text"), flatten = TRUE)
# extract list items and convert to a dataframe
tabnames <- response$fields$items %>% map(~.$labels %>% unlist)
values <- response$cubes[[1]]$values
dimnames(values) <- tabnames

uc <- as.data.frame.table(values, stringsAsFactors = FALSE) %>%
  as_tibble() %>%
  set_names(c(response$fields$label,"value")) %>%
  left_join(lookup%>%select(msoa11cd,msoa11nm,msoa11hclnm), by = c("National - Regional - LA - OAs" = "msoa11nm")) %>%
  mutate(period=format(as.yearmon(`Month`, "%B %Y"),"%b %Y")) %>%
  select(area_code=msoa11cd, area_name = msoa11hclnm, period, value) %>%
  filter(period %in% c(latestm12,latest)) %>%
  spread(period,value) %>%
  mutate(additional_claims= get(latest)-get(latestm12),
        percentage_change = round(additional_claims/get(latestm12) *100,0),
        `Share` = round(get(latest)/sum(get(latest)) *100,1)) %>%
  left_join(msoa_pop, by="area_code") %>%
  mutate(latestm12_rate = round(get(latestm12)*100/pop16_64,1), latest_rate=round(get(latest)*100/pop16_64,1),Change=round(latest_rate-latestm12_rate,1))

table7 <- uc %>%
  select(-c(area_code,pop16_64))
 
```
 
The Trafford MSOA with the largest share of claims in `r latestch` was `r filter(table7, Share == max(Share)) %>% pull(area_name)` with `r filter(table7, Share == max(Share))$Share`% of all Universal Credit claims in Trafford. 24 out of the 28 MSOAs have more claims in `r latestch`  compared to `r latestm12ch`. The MSOAs with more additional claims when comparing `r latestm12ch` to `r latestch` were `r arrange(table7,desc(additional_claims))$area_name[1]` (`r arrange(table7,desc(additional_claims))$additional_claims[1]`), `r arrange(table7,desc(additional_claims))$area_name[2]` (`r arrange(table7, desc(additional_claims))$additional_claims[2]`) and `r arrange(table7, desc(additional_claims))$area_name[3]` (`r arrange(table7,desc(additional_claims))$additional_claims[3]`). The MSOAs with more increase in percentage change from `r latestm12ch` to `r latestch` were `r arrange(table7, desc(percentage_change))$area_name[1]` with `r arrange(table7, desc(percentage_change))$percentage_change[1]`% change, followed by `r arrange(table7, desc(percentage_change))$area_name[2]` with `r arrange(table7, desc(percentage_change))$percentage_change[2]`% change. The MSOAs with reductions in claims when comparing `r latestm12ch` to `r latestch` are `r arrange(table7,additional_claims)$area_name[1]` (`r arrange(table7,additional_claims)$additional_claims[1]`), `r arrange(table7,additional_claims)$area_name[2]` (`r arrange(table7, additional_claims)$additional_claims[2]`) and  `r arrange(table7,additional_claims)$area_name[3]` (`r arrange(table7, additional_claims)$additional_claims[3]`). The MSOAs with reduction in percentage change from `r latestm12ch` to `r latestch` were `r arrange(table7, percentage_change)$area_name[1]` with `r arrange(table7, percentage_change)$percentage_change[1]`% change, followed by `r arrange(table7, percentage_change)$area_name[3]` and `r arrange(table7, percentage_change)$area_name[2]` with `r arrange(table7, percentage_change)$percentage_change[2]`%.


```{r UCMSOATable}

#The MSOAs that increased the number of claims from `r latestm12ch` to `r latestch` were `r arrange(table7, desc(additional_claims))$area_name[1]` with `r arrange(table7, desc(additional_claims))$additional_claims[1]`  and `r arrange(table7, desc(additional_claims))$area_name[2]` with `r arrange(table7, desc(additional_claims))$additional_claims[2]`.

tbl7 <- reactable(
  table7,
  pagination = FALSE,
  defaultSorted = "area_name",
  defaultSortOrder = "asc",
  defaultColDef = colDef(headerClass = "header", align = "right", maxWidth = 60, format = colFormat(separators = TRUE)),
  columns = list(
    area_name = colDef(name = "MSOA", maxWidth = 220, align = "left"),
    additional_claims = colDef(
      name = "Additional claims",
      maxWidth = 175,
      align = "left",
      cell = function(value) {
        cell_function(value,table7$additional_claims)
      }
    ),
    percentage_change = colDef(name = "% Change", maxWidth = 80),
    Share = colDef(name = "Share Jun 23", maxWidth = 80, align = "left"),
    latestm12_rate = colDef(name = latestm12, maxWidth = 60, align = "left"),
    latest_rate = colDef(name = latest, maxWidth = 60, align = "left"),
    `Change` = colDef(
      maxWidth = 175,
      cell = function(value) {
        cell_function(value,table7$Change)
      }
    )
    ),
  columnGroups = list(
    colGroup(name = "Count", columns = c(latestm12, latest, "additional_claims", "percentage_change", "Share")),
    colGroup(name = "Rate", columns = c("latestm12_rate", "latest_rate", "Change"))
  ),
  compact = TRUE,
  class = "large-table"
)

div(class = "div",
div(class= "large-table tableTitle", "Change in the Universal Credit claims for Trafford's MSOAs"),
tbl7
)

```

The MSOAs with higher rates of Universal Credit claims in `r latestch` were `r filter(table7, latest_rate == max(latest_rate)) %>% pull(area_name)` with `r filter(table7, latest_rate == max(latest_rate))$latest_rate`% and `r arrange(table7, desc(latest_rate))$area_name[2]` with `r arrange(table7, desc(latest_rate))$latest_rate[2]`%. The MSOAs where the rate of Universal Credit claims has increased more from `r latestm12ch` to `r latestch` were `r arrange(table7, desc(Change))$area_name[1]` with `r arrange(table7, desc(Change))$Change[1]`% and `r arrange(table7, desc(Change))$area_name[2]` with `r arrange(table7, desc(Change))$Change[2]`% increase. The MSOA where the rate of Universal Credit claims has decreased more from `r latestm12ch` to `r latestch` was `r arrange(table7, Change)$area_name[1]` with `r arrange(table7, Change)$Change[1]`%, `r arrange(table7, Change)$area_name[2]` with `r arrange(table7, Change)$Change[2]`%, and `r arrange(table7, Change)$area_name[3]` with `r arrange(table7, Change)$Change[3]`%.

  <div class="row">
  <div class = "col-sm-6">

```{r UCMSOADotPlot, fig.height=7}

uc_plot_dot <- table7 %>% 
  select (area_name,{{latestm12}}:=latestm12_rate, {{latest}}:=latest_rate) %>%
  gather(period,value,-one_of("area_name")) %>%
  mutate(period=as.yearmon(period, format="%B %Y")) %>%
  bind_rows(trafford_uc %>% select(area_name,period,value=rate) %>% filter(period %in% c(as.yearmon(latest),as.yearmon(latestm12))) )%>%
   mutate(period = factor(period, levels = c(latestm12,latest)),
         area_name = fct_reorder(factor(area_name), ifelse(period == latest, value, NA), na.rm = TRUE),
         area_name = fct_relevel(factor(area_name), "Trafford"))

ggplot(uc_plot_dot, aes(value,area_name)) +
  geom_line(color = "grey", size=0.5) +
  geom_point(aes(color = period), size = 3) +
  scale_x_continuous(labels = function(x){ paste0(x, "%") }, limits = c(0, 30)) +
  scale_colour_manual(values = c( "#66c2a4", "#006d2c")) +
  labs(x = "percentage of residents aged 16 to 64", y = NULL,
       title = "Universal Credit claims rate", 
       subtitle = paste0("Trafford MSOAs, ", latestm12, " and ", latest), 
       caption = "Source: DWP")+
  theme_minimal() +
  theme(plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
        panel.grid.major.y = element_blank(),
        axis.text.y = element_text(hjust = 1),
        legend.position = "bottom", 
        legend.title = element_blank())

```

   </div>
  
   <div class = "col-sm-6">

```{r UCMSOAMap, fig.height=6}

msoa_uc <- st_read("msoa_boundaries.geojson", quiet = TRUE) %>% 
  left_join(., select(uc,latest_rate,area_code), by = "area_code")

ggplot() +
  geom_sf(data = msoa_uc, aes(fill = latest_rate), alpha = 1, colour = "#FFFFFF",  size = 0.5) +
  scale_fill_gradientn(colours = c("#edf8fb","#b2e2e2","#66c2a4","#2ca25f","#006d2c"), 
                       na.value = "#f0f0f0",
                       guide = guide_legend(keyheight = unit(3, units = "mm"), 
                                            keywidth=unit(12, units = "mm"), 
                                            label.position = "bottom", 
                                            title = "Rate",
                                            title.position = 'top', 
                                            nrow = 1)) +
  labs(x = NULL, y = NULL, title = "Universal Credit claims rate",
       subtitle = paste0("Trafford MSOAs, ", latest),
       caption = "Source: DWP\nContains OS data © Crown copyright and database right 2020") +
  theme(plot.margin = unit(c(0.5,0,1,0), "cm"),
        plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
        axis.text = element_blank(),
        plot.caption = element_text(size=6,margin = margin(t = 55)),
        legend.position = c(0.3,-0.05),
        legend.title = element_text(size = 9),
        legend.text = element_text(size = 8),
        plot.background = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        axis.ticks = element_blank())


```

   </div>
  </div> 

```{r UCAgeData}
#Universal Credit
# Source: DWP
# URL: https://stat-xplore.dwp.gov.uk/webapi/metadata/UC_Monthly/UC_Monthly.html
# Licence: Open Government Licence


query <- list(database = unbox("str:database:UC_Monthly"),
              measures = "str:count:UC_Monthly:V_F_UC_CASELOAD_FULL",
              dimensions = c("str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE",
                             "str:field:UC_Monthly:F_UC_DATE:DATE_NAME",
                             "str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:AGE_CODE") %>% matrix(),
              recodes = list(
                `str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE` = list(
                  map = list("str:value:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE:V_C_MASTERGEOG11_LA_TO_REGION:E08000009")),
                `str:field:UC_Monthly:F_UC_DATE:DATE_NAME` = list(
                  map = as.list(paste0("str:value:UC_Monthly:F_UC_DATE:DATE_NAME:C_UC_DATE:",c(202001,202002,202003,202004,202005,202006,202007,202008,202009,202010,202011,202012,202101,202102,202103,202104,202105,202106,202107,202108,202109,202110,202111,202112,202201,202202,202203,202204,202205,202206,202207,202208,202210,202211,202212,202301,202302,202303,202304,202305,202306)))),
                `str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:AGE_CODE` = list(
                  map = as.list(paste0("str:value:UC_Monthly:V_F_UC_CASELOAD_FULL:AGE_CODE:C_UC_AGE_BAND:",c(1,2,3,4,5,6,7,8,9,10,999))))

              )) %>% toJSON()
request <- POST(
  url = path,
  body = query,
  config = add_headers(APIKey = api_key),
  encode = "json")
response <- fromJSON(content(request, as = "text"), flatten = TRUE)
# extract list items and convert to a dataframe
tabnames <- response$fields$items %>% map(~.$labels %>% unlist)
values <- response$cubes[[1]]$values
dimnames(values) <- tabnames

traf_uc <- as.data.frame.table(values, stringsAsFactors = FALSE) %>%
  as_tibble() %>%
  set_names(c(response$fields$label,"value")) %>%
  mutate(indicator="People on Universal Credit",
         period=format(as.yearmon(`Month`, "%B %Y"),"%b %Y")) %>%
  select(area_name = `National - Regional - LA - OAs`, indicator, Age="Age (bands and single year)", period, value)


table8 <- traf_uc %>%
  select(-c(area_name,indicator)) %>%
  filter(period %in% c(latestm12,latest)) %>%
  spread(period,value) %>%
  mutate(additional_claims= get(latest)- get(latestm12),
         percentage_change = round(additional_claims/get(latestm12) *100,0)) %>%
  select(Age, `latestm12`= latestm12, `latest` = latest, additional_claims, percentage_change)
```
 
In Trafford from `r latestm12ch` to `r latestch` the `r filter(table8, percentage_change == max(percentage_change)) %>% pull(Age)` age band has the largest increment in percentage change of Universal Credit claims with `r filter(table8, percentage_change == max(percentage_change)) %>% pull(percentage_change)`%. When considering the number of additional claims from  `r latestm12ch` to `r latestch` the 5-year age band `r filter(table8, additional_claims == max(additional_claims)) %>% pull(Age)` years had the highest with `r filter(table8, additional_claims == max(additional_claims)) %>% pull(additional_claims)` more claims. All age bands had an increment of claims compared to `r latestm12ch`.
 
```{r UCAgeTable}

#The age bands `r arrange(table8, additional_claims)$Age[1]` and `r arrange(table8, additional_claims)$Age[2]` have the largest decrease in number of Universal Credit claims. Claimants  under 30 decreased or not changed in number whilst claimants 30 and over increased in number when comparing `r latestm12ch` to `r latestch`.

tbl8 <- reactable(
  table8,
  pagination = FALSE,
  defaultSorted = "Age",
  defaultSortOrder = "asc",
  defaultColDef = colDef(headerClass = "header", align = "right", maxWidth = 60, format = colFormat(separators = TRUE)),
  columns = list(
    Age = colDef(name = "Age", maxWidth = 60, align = "left"),
    latestm12 = colDef(name = latestm12, maxWidth = 60),
    latest = colDef(name = latest, maxWidth = 60),
    additional_claims = colDef(
      name = "Additional claims",
      maxWidth = 175,
      cell = function(value) {
        cell_function(value,table8$additional_claims)
      }
    ),
    percentage_change = colDef(
      name = "% Change",
      maxWidth = 175,
      cell = function(value) {
        cell_function(value,table8$percentage_change)
      }
    )
    ),
  compact = TRUE,
  class = "large-table"
)

div(class = "div",
div(class= "large-table tableTitle", "Change in the Universal Credit claims by age band for Trafford"),
tbl8
)

```


Following the start of the Covid-19 pandemic in March 2020 the number of claims for all ages rose dramatically from pre-pandemic levels at a similar proportion for all age bands with the age band 16-24 rising faster in May 2020. At the start of 2020 the age band 25-34 had the highest number of claims and also the highest rate maintaining that position after the rise from March 2020 until May 2022 when the 35-44 became the age band with the highest number of claims. The age band 16-24 has the lowest number of claims among the other age bands from July 2021. All the age bands are increasing their number of claims particularly from January 2023.

  <div class="row">
   <div class = "col-sm-6">
  
```{r UCAgeMline}

uc_plot_age_c <- traf_uc %>%
  select(-c(area_name, indicator)) %>%
  mutate(Age=fct_collapse(Age, `16-24` = c("16-19", "20-24"),
                       `25-34` = c("25-29", "30-34"),
                       `35-44` = c("35-39", "40-44"),
                       `45-54` = c("45-49", "50-54"),
                       `55+` = c("55-59", "60-65","Over 65"))) %>%
  group_by(Age,period) %>%
  summarise_all(sum) %>%
  mutate(date=as.Date(as.yearmon(period), format="%b %Y")) %>%
  mutate(label = if_else(date == max(date), Age, factor(NA)))
  

 ggplot(uc_plot_age_c, aes(x = date, y = value, colour = Age)) +
  geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +
  geom_line(size = 1, show.legend = FALSE) +
  geom_text_repel(aes(x = max(date) + 60, label = label, fontface = "bold", hjust = 1), direction = "y", show.legend = FALSE) +
  #scale_color_startrek() +
  scale_x_date(breaks = c(min(uc_plot_age_c$date), as.Date(as.yearmon("Apr 2020"),format="%b %Y"), as.Date(as.yearmon("Jan 2021"),format="%b %Y"), as.Date(as.yearmon("Jan 2022"),format="%b %Y"), max(uc_plot_age_c$date)), date_labels = "%b %y") +
  scale_y_continuous(expand = c(0.005, 0.005)) +
  labs(title = "Univeral Credit claims by age band",
       subtitle = paste("Trafford, ", base, " - ",latest),
       caption = "Source: DWP",
       x = NULL,
       y = NULL,
       colour = NULL) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
    plot.margin = unit(rep(0.5,4), "cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks.x = element_line(colour = "#333333", size = 0.5))  +
  coord_cartesian(clip = "off") 


```

   </div>
  
   <div class = "col-sm-6">
  
```{r UCAgeMlineRate}
pop_trafford_age2 <- pop_trafford_age_raw %>%
    select(age = C_AGE_NAME, n = OBS_VALUE) %>%
    mutate(age = parse_number(age),
          ageband = cut(age,
                        breaks = c(0,15,24,34,44,54,64,120),
                        labels = c("0-15","16-24","25-34","35-44","45-54","55-65","65+"),
                        right = TRUE,
                        include.lowest = TRUE)) %>%
   group_by(ageband) %>%
   summarise(population = sum(n))

uc_plot_age_r <- traf_uc %>%
  select(-c(area_name, indicator)) %>%
  mutate(Age=fct_collapse(Age, `16-24` = c("16-19", "20-24"),
                       `25-34` = c("25-29", "30-34"),
                       `35-44` = c("35-39", "40-44"),
                       `45-54` = c("45-49", "50-54"),
                       `55-65` = c("55-59", "60-65"),
                       `Over 65` = "Over 65")) %>%
  filter(!Age == "Over 65") %>%
  group_by(Age,period) %>%
  summarise_all(sum) %>%
  left_join(pop_trafford_age2, by = c("Age" = "ageband")) %>%
  mutate(rate = round((value/population)*100,1),
         date=as.Date(as.yearmon(period), format="%b %Y")) %>%
  mutate(label = if_else(date == max(date), Age, factor(NA)))
  

 ggplot(uc_plot_age_r, aes(x = date, y = rate, colour = Age)) +
  geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +
  geom_line(size = 1, show.legend = FALSE) +
  geom_text_repel(aes(x = max(date) + 60, label = label, fontface = "bold", hjust = 1), direction = "y", show.legend = FALSE) +
  #scale_color_startrek() +
  scale_x_date(breaks = c(min(uc_plot_age_r$date), as.Date(as.yearmon("Apr 2020"),format="%b %Y"), as.Date(as.yearmon("Jan 2021"),format="%b %Y"), as.Date(as.yearmon("Jan 2022"),format="%b %Y"), max(uc_plot_age_r$date)), date_labels = "%b %y") +
  scale_y_continuous(expand = c(0.005, 0.005)) +
  labs(title = "Univeral Credit claims rate by age band",
       subtitle = paste("Trafford, ", base, " - ",latest),
       caption = "Source: DWP",
       x = NULL,
       y = NULL,
       colour = NULL) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
    plot.margin = unit(rep(0.5,4), "cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks.x = element_line(colour = "#333333", size = 0.5))  +
  coord_cartesian(clip = "off") 

```

   </div>
  </div>
 
```{r UCConditionData}

#Universal Credit
# Source: DWP
# URL: https://stat-xplore.dwp.gov.uk/webapi/metadata/UC_Monthly/UC_Monthly.html
# Licence: Open Government Licence

query <- list(database = unbox("str:database:UC_Monthly"),
              measures = "str:count:UC_Monthly:V_F_UC_CASELOAD_FULL",
              dimensions = c("str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE",
                             "str:field:UC_Monthly:F_UC_DATE:DATE_NAME",
                             "str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:CCCONDITIONALITY_REGIME") %>% matrix(),
              recodes = list(
                `str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE` = list(
                  map = as.list(paste0("str:value:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE:V_C_MASTERGEOG11_MSOA_TO_LA:E0", seq(2001259, 2001286, 1)))),
                `str:field:UC_Monthly:F_UC_DATE:DATE_NAME` = list(
                  map = as.list(paste0("str:value:UC_Monthly:F_UC_DATE:DATE_NAME:C_UC_DATE:",c(202306)))),
                `str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:CCCONDITIONALITY_REGIME` = list(
                  map = as.list(paste0("str:value:UC_Monthly:V_F_UC_CASELOAD_FULL:CCCONDITIONALITY_REGIME:C_UC_CONDITIONALITY_REGIME:",c("AA","AB","BC","BD","DF","CE"))))

              )) %>% toJSON()
request <- POST(
  url = path,
  body = query,
  config = add_headers(APIKey = api_key),
  encode = "json")
response <- fromJSON(content(request, as = "text"), flatten = TRUE)
# extract list items and convert to a dataframe
tabnames <- response$fields$items %>% map(~.$labels %>% unlist)
values <- response$cubes[[1]]$values
dimnames(values) <- tabnames

uc_condition <- as.data.frame.table(values, stringsAsFactors = FALSE) %>%
  as_tibble() %>%
  set_names(c(response$fields$label,"value")) %>%
  left_join(lookup%>%select(msoa11cd,msoa11nm,msoa11hclnm), by = c("National - Regional - LA - OAs" = "msoa11nm")) %>%
  mutate(period=format(as.yearmon(`Month`, "%B %Y"),"%b %Y")) %>%
  select(area_code=msoa11cd, area_name = msoa11hclnm, condition = `Conditionality Regime`, period, value)

```

The [Conditionality](https://stat-xplore.dwp.gov.uk/webapi/metadata/UC_Monthly/Conditionality%20Regime.html) regimen for entitlement to Universal Credit is associated to work-related things that claimants will have to do to maintain eligibility. In Trafford the largest proportion of Universal Credit claims are in the "No work requirements" category however there is also a significant proportion of claims under "Searching for work". In Old Trafford, 33% of Universal Credit claimants were under "No work requirements" whilst 31% where "Searching for work". In Partington, 43% of claims were under "No work requirements" whilst 20% of claims were under "Searching for work" and "Working - no requirements" each.

```{r UCConditionMMK, fig.height=5.5,fig.width=7}
uc_msoa_cond<- uc_condition %>%
  group_by(period, area_code, area_name) %>%
  mutate(condition = fct_rev(condition),
         total = sum(value), percent = value/total) 
# plot data ---------------------------

ggplot(uc_msoa_cond, aes(x = area_name, y = percent, width = total, fill = condition)) +
  geom_col(position = "stack", colour = NA) +
  facet_grid(~fct_reorder(area_name,total, .desc = T), scales = "free_x", space = "free_x") +
  scale_y_continuous(expand = c(0.005, 0.005), breaks = c(0, 0.5, 1), labels = percent) +
  scale_fill_brewer(palette = "Set3") +
  labs(x = NULL, y = NULL, 
       title = "Universal Credit Claims by Trafford's MSOAs and Conditionality regimen",
       subtitle = unique(uc_msoa_cond$period),
       caption = "The width of the columns is proportional to the number of claims. Source: DWP.",
       fill = NULL) +
  theme(plot.title = element_text(size=12,face = "bold"),
        plot.subtitle = element_text(size=10),
    panel.spacing.x = unit(0.007, "npc"),
    panel.grid.major = element_blank(),
    strip.text.x = element_blank(),
    legend.position = "right",
    axis.text.x = element_text(size=8, angle = 90, hjust = 1, vjust=0.5),
    panel.background = element_blank(),
    axis.ticks = element_blank())

```


 
```{r UCConditionDataTrend}

#Universal Credit
# Source: DWP
# URL: https://stat-xplore.dwp.gov.uk/webapi/metadata/UC_Monthly/UC_Monthly.html
# Licence: Open Government Licence

query <- list(database = unbox("str:database:UC_Monthly"),
              measures = "str:count:UC_Monthly:V_F_UC_CASELOAD_FULL",
              dimensions = c("str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE",
                             "str:field:UC_Monthly:F_UC_DATE:DATE_NAME",
                             "str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:CCCONDITIONALITY_REGIME") %>% matrix(),
              recodes = list(
                `str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE` = list(
                  map = list("str:value:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE:V_C_MASTERGEOG11_LA_TO_REGION:E08000009")),
                `str:field:UC_Monthly:F_UC_DATE:DATE_NAME` = list(
                  map = as.list(paste0("str:value:UC_Monthly:F_UC_DATE:DATE_NAME:C_UC_DATE:",c(202001,202002,202003,202004,202005,202006,202007,202008,202009,202010,202011,202012,202101,202102,202103,202104,202105,202106,202107,202108,202109,202110,202111,202112,202201,202202,202203,202204,202205,202206,202207,202208,202210,202211,202212,202301,202302,202303,202304,202305,202306)))),
                `str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:CCCONDITIONALITY_REGIME` = list(
                  map = as.list(paste0("str:value:UC_Monthly:V_F_UC_CASELOAD_FULL:CCCONDITIONALITY_REGIME:C_UC_CONDITIONALITY_REGIME:",c("AA","AB","BC","BD","DF","CE"))))

              )) %>% toJSON()
request <- POST(
  url = path,
  body = query,
  config = add_headers(APIKey = api_key),
  encode = "json")

response <- fromJSON(content(request, as = "text"), flatten = TRUE)
# extract list items and convert to a dataframe
tabnames <- response$fields$items %>% map(~.$labels %>% unlist)
values <- response$cubes[[1]]$values
dimnames(values) <- tabnames

uc_condition_trend <- as.data.frame.table(values, stringsAsFactors = FALSE) %>%
  as_tibble() %>%
  set_names(c(response$fields$label,"value")) %>%
  mutate(period=format(as.yearmon(`Month`, "%B %Y"),"%b %Y")) %>%
  select(area_name = "National - Regional - LA - OAs", condition = `Conditionality Regime`, period, value) %>%
  group_by(period) %>%
  mutate(Percent = round(value *100/sum(value),1)) %>%
  ungroup()

```

Claimants continue to move across from legacy benefits to Universal credit therefore, the number of claimants with no work requirements continue to grow. In `r latestch`, `r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$Percent[1]`% (`r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$value[1]`) of claims in Trafford were under "`r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$condition[1]`", `r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$Percent[2]`% (`r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$value[2]`) were under "`r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$condition[2]`", `r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$Percent[3]`% (`r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$value[3]`) were under "`r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$condition[3]`", `r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$Percent[4]`% (`r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$value[4]`) were under "`r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$condition[4]`", `r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$Percent[5]`% (`r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$value[5]`) were under "`r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$condition[5]`", and `r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$Percent[6]`% (`r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$value[6]`) where under "`r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$condition[6]`".
<div class="row">
<div class = "col-sm-12">
   
```{r UCConditionalityLine, fig.height=3.5, fig.width=7}

uc_plot_conditionality_c <- uc_condition_trend %>%
  mutate(date=as.Date(as.yearmon(period), format="%b %Y")) %>%
  mutate(condition = fct_rev(condition)) 
  

 ggplot(uc_plot_conditionality_c , aes(x = date, y = value, colour = condition)) +
  geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +
  geom_line(size = 1) +
  #geom_text_repel(aes(x = max(date) + 60, label = label, fontface = "bold", hjust = 1), direction = "y", show.legend = FALSE) +
  #scale_color_startrek() +
   scale_color_brewer(palette = "Set3") +
  scale_x_date(breaks = c(min(uc_plot_conditionality_c$date), as.Date(as.yearmon("Apr 2020"),format="%b %Y"), as.Date(as.yearmon("Jan 2021"),format="%b %Y"), as.Date(as.yearmon("Jan 2022"),format="%b %Y"), max(uc_plot_conditionality_c$date)), date_labels = "%b %y") +
  scale_y_continuous(expand = c(0.005, 0.005)) +
  labs(title = "Univeral Credit claims by conditionality",
       subtitle = paste("Trafford, ", base, " - ",latest),
       caption = "Source: DWP",
       x = NULL,
       y = NULL,
       colour = NULL) +
  theme_minimal(base_size = 10) +
  theme(plot.title = element_text(size=12,face = "bold"),
        plot.subtitle = element_text(size=10),
    plot.margin = unit(rep(0.5,4), "cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks.x = element_line(colour = "#333333", size = 0.5))  +
  coord_cartesian(clip = "off") 

```

</div>
</div>

</section>

</main>



```{css}

@import url('https://fonts.googleapis.com/css?family=Roboto');

html, body
{
    padding: 0;
    margin: 0;
    height: 100%;
    font-family: 'Open Sans', sans-serif;
    color: #212121;
}

h1, h2, h3, h4, h5, h6, header
{
    font-family: 'Roboto', sans-serif;
    color: #707070;
}

a
{
    color: #046dc3;
    text-decoration: none;
}

a:hover
{
    text-decoration: underline;
}

h1, h2, h3, h4, h5, h6, .header
{
    font-family: 'Roboto', sans-serif;
    color: #757575;
}

p {
  line-height: 1.8em;
}

.main-container {
  max-width: 1000px;
}
.header {
  border-bottom: 2px solid #555;
  font-weight: 400;
  text-transform: uppercase;
}

.header:hover {
  background-color: #eee;
}

.bar-cell {
  display: flex;
  align-items: center;
}

.number {
  width: 35px;
  text-align: right;
}

.bar-chart {
  display: flex;
  flex-grow: 1;
  height: 100%;
}

.bar-neg-container {
  display: flex;
  justify-content: flex-end;
}

.bar-pos-container {
  display: flex;
  justify-content: flex-start;
}

.bar {
  height: 12px;
}

.large-table {
  font-size: 13px;
}

.table {
  font-size: 12px;
}

.small-table,  .xsmall-table {
  font-size: 11px;
}

.small-table .rt-td {
  padding: 3px 3px;
}
.xsmall-table .rt-td .rt-td-inner{  
  padding: 0px 0px;
  font-size: 11px;
}

.large-table .rt-td {
  white-space: nowrap;
  text-overflow: ellipsis;
}

.tableTitle {
  font-weight: bold;
  margin: 0px;
}

```
