---
title: "Claimant Count and Universal Credit in Trafford"
lang: "en-GB"
output: html_document
---

```{r setup, include=FALSE, results="hide"}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse) ; library(httr) ; library(readxl) ; library(lubridate)
library(zoo) ;library(sf) ; library(jsonlite) ; library(scales)
library(reactable) ; library(htmltools) ; library(ggrepel)

```


```{r CCdata}

base <- "Jan 2020"
basech <- "January 2020"

#Claimant Count
# Source: ONS
# URL: https://www.nomisweb.co.uk/sources/cc
# Licence: Open Government Licence

cc_trafford <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_162_1.data.csv?geography=1811939363&date=latestMINUS57-latest&gender=0&age=0&measure=1,2&measures=20100") %>%
    select(period = DATE_NAME,
         area_code = GEOGRAPHY_CODE,
         area_name = GEOGRAPHY_NAME,
         measure = MEASURE_NAME,
         value = OBS_VALUE)

latest <- as.character(max(as.yearmon(cc_trafford$period)))
latestm12 <- as.character(max(as.yearmon(cc_trafford$period))-1)

latestch <- "Oct 2024"
latestm12ch <- "Oct 2023"

cc_traf <- cc_trafford %>%
  mutate(period=as.yearmon(period, "%b %Y"),
         measure=ifelse(measure=="Claimant count","count","rate")) %>%
  select(-area_code) %>%
  spread(measure,value) %>%
  mutate(change=round((count*100/count[1])-100,1))

```

<main>
<section>

# Claimant Count in Trafford

The [Claimant Count](https://www.nomisweb.co.uk/sources/cc) indicates the number of people claiming benefits principally for the reason of being unemployed. The Claimant Count in `r latestch` was `r filter(cc_traf,period==as.yearmon(latest, "%b %Y"))$change`% higher than in `r basech`. The rate of claimants to residents aged 16 to 64 was `r filter(cc_traf,period==as.yearmon(latest, "%b %Y"))$rate`% in `r latestch` representing an increment of `r filter(cc_traf,period==as.yearmon(latest, "%b %Y"))$rate-cc_traf$rate[1]`% points from `r basech`. The number of claimants decreased from April 2021. In July 2024 the claims showed and increase.


 <div class="container-fluid">
  <div class="row">
   <div class = "col-sm-4">

```{r CCTraffPCHTable}

table1 <- cc_traf %>% 
  mutate(period = as.Date(period,frac=0)) %>%
  select(-area_name)

tbl1 <- reactable(
  table1,
  pagination = TRUE,
  defaultSorted = "period",
  defaultSortOrder = "desc",
  fullWidth = FALSE,
  defaultColDef = colDef(headerClass = "header", align = "right", minWidth = 60, format = colFormat(separators = TRUE)),
  columns = list(
    period = colDef(name = "Date", align = "left", cell = function(value) date_format("%b %Y")(value)),
   # area_name = colDef(name = "Area", align = "left"),
    change = colDef(name = "Change %",
                    format = colFormat(digits = 1))
    ),
  compact = TRUE,
  class = "table"
)

div(class = "div",
div(class= "tableTitle", "Change in the Claimant Count for Trafford compared to January 2020"),
tbl1
)

```

   </div>
  
   <div class = "col-sm-8">
```{r CCTrafPCHPLineP}

cc_traffordp <- cc_traf %>% mutate(date=as.Date(period,frac=0)) %>%
  mutate(change = round(change,0)) %>%
  mutate(index = case_when(change == 0 ~ "Label",
                           period == latest ~ "Label",
                           change -lag(change)<=0 & change - lead(change)<0 ~ "Label",
                           change -lag(change)>=0 & change - lead(change)>0 ~ "Label",
                           change -lead(change)<(-20) | change - lag(change)>20 ~ "Label",
                           TRUE ~ "No label"))

ggplot(cc_traffordp,aes(x = date, y = round(change,0))) + 
  geom_line(aes(group=1), linewidth = 1, alpha = 1, color="#006d2c") + #"#66c2a4", "#006d2c" dark
  geom_point(size = 3, alpha = 1, color="#006d2c") + 
    geom_text(data=cc_traffordp,
            aes(label = paste0(change, "%"), color = index),
            fontface = 1.5, size = 2.5, hjust = 0.3, vjust = -1.3)+
  scale_color_manual(values = c("#006d2c", "transparent")) +
  scale_x_date(breaks = seq(as.Date("2020-03-01"), as.Date(max(cc_traffordp$period)), by="3 months"), date_labels="%b %y", expand = c(0.03,0.03)) +
  scale_y_continuous(limits = c(0,135)) +
  labs(title = "Claimant Count percentage change",
       subtitle = paste("Trafford, ", base, " - ", latest),
       caption = "Source: ONS",
       x = NULL,
       y = NULL,
       colour = NULL) +
  #theme_lab() +
  theme_minimal() +
  theme(plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        panel.grid.minor = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.text.y = element_blank(),
        legend.position = "none")

```

   </div>
  </div>
 
```{r CCMSOAData, results="hide"}
#Claimant Count
# Source: ONS
# URL: https://www.nomisweb.co.uk/sources/cc
# Licence: Open Government Licence

 #House of Commons Library MSOA Names
# URL: https://visual.parliament.uk/msoanames
lookup <- read_csv("https://houseofcommonslibrary.github.io/msoanames/MSOA-Names-Latest.csv") %>%
  filter(Laname=="Trafford")

# Mid-2020 population estimates
# Source: Nomis / ONS
# URL: https://www.nomisweb.co.uk/datasets/pestsyoaoa
# Licence: Open Government Licence 3.0

# msoa_pop <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_2010_1.data.csv?geography=1245709510...1245709537&date=latest&gender=0&c_age=203&measures=20100") %>%
#     select(area_code = GEOGRAPHY_CODE,
#          pop16_64 = OBS_VALUE)


# Mid-2022 population estimates
# URL: https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/middlesuperoutputareamidyearpopulationestimatesnationalstatistics

tmp <- tempfile(fileext = ".xlsx")


GET(url = "https://www.ons.gov.uk/file?uri=/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/middlesuperoutputareamidyearpopulationestimates/mid2021andmid2022/sapemsoasyoatablefinal.xlsx",
    write_disk(tmp))

msoa_pop <- read_xlsx(tmp, sheet = 6, skip = 3) %>%
  filter(`LAD 2021 Code` == "E08000009") %>%
  mutate(F16to64 = rowSums(pick(F16:F64), na.rm = TRUE), M16to64 = rowSums(pick(M16:M64), na.rm = TRUE)) %>%
  mutate(Count = F16to64+M16to64) %>%
  select(area_code = `MSOA 2021 Code`, pop16_64 = Count) 

# claimant_ward <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_162_1.data.csv?geography=1660945005...1660945019,1660945021,1660945020,1660945022...1660945025&date=latestMINUS12,latest&gender=0&age=0&measure=1,2&measures=20100")
#https://www.nomisweb.co.uk/api/v01/dataset/NM_162_1.data.csv?geography=1245709510...1245709537&date=latest&gender=0&age=0&measure=1&measures=20100

claimant_msoa <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_162_1.data.csv?geography=1245709510...1245709537&date=latestMINUS12,latest&gender=0&age=0&measure=1&measures=20100")

claimant_area <- claimant_msoa %>%
  select(period = DATE_NAME,
         area_code = GEOGRAPHY_CODE,
         area_code2 = GEOGRAPHY_NAME,
         measure = MEASURE_NAME,
         value = OBS_VALUE) %>%
  mutate(period = as.yearmon(period, "%b %Y")) %>%
  left_join(lookup%>%select(msoa11cd,msoa11nm,msoa11hclnm), by = c("area_code2" = "msoa11nm")) %>%
  rename(area_name = msoa11hclnm)

c_count <- claimant_area %>%
  spread(period,value) %>%
  mutate(additional_claims= get(latest) - get(latestm12),
         `Share` = round(get(latest)/sum(get(latest)) *100,1),
         percentage_change = round(additional_claims/get(latestm12) *100,0))

c_rate <- c_count %>%
    left_join(msoa_pop, by="area_code") %>%
    mutate(latestm12_r = round(get(latestm12)*100/pop16_64,1), latest_r=round(get(latest)*100/pop16_64,1),Change=round(latest_r-latestm12_r,1))
  

```
 
The Trafford MSOA with the largest share of claims in `r latestch` was `r filter(c_count, Share == max(Share)) %>% pull(area_name)` with `r filter(c_count, Share == max(Share))$Share`% of all claimants within Trafford, almost the double of claims than Gorse Hill, the MSOA with the second largest number of claims. 20 MSOAs out of 28 have more claims in `r latestch`  compared to `r latestm12ch`. `r arrange(c_count, desc(additional_claims))$area_name[1]` have the largest increase in number of claims with `r abs(arrange(c_count, desc(additional_claims))$additional_claims[1])` more, followed by `r arrange(c_count, desc(additional_claims))$area_name[2]` (`r arrange(c_count, desc(additional_claims))$additional_claims[2]`). `r arrange(c_count, additional_claims)$area_name[1]` and `r arrange(c_count, additional_claims)$area_name[2]` had a reduction in number of claims with `r abs(arrange(c_count, additional_claims)$additional_claims[1])` less, along with `r arrange(c_count, additional_claims)$area_name[3]`, `r arrange(c_count, additional_claims)$area_name[4]` and `r arrange(c_count, additional_claims)$area_name[4]`  with `r arrange(c_count, additional_claims)$additional_claims[3]`.


```{r CCMSOATable}

table2 <- c_rate %>% rename(latestm12_c=latestm12, latest_c=latest) %>%
  select(-area_code2, -measure, -msoa11cd, -pop16_64, -area_code)

cell_function <- function(value, colTable){
        width_neg_tot <- ifelse(min(colTable, na.rm = TRUE) < 0,abs(min(colTable, na.rm = TRUE)),0)
        width_pos_tot <- ifelse(max(colTable, na.rm = TRUE) >= 0,max(colTable, na.rm = TRUE),0)
        width_bar_tot <- width_neg_tot + width_pos_tot
        width_neg_percent <- paste0(width_neg_tot * 100 /width_bar_tot, "%")
        width_pos_percent <- paste0(width_pos_tot * 100 /width_bar_tot, "%")
        width_neg <- ifelse(value < 0, paste0(abs(value) * 100 /width_neg_tot, "%"),0) 
        width_pos <- ifelse(value >= 0, paste0(value * 100 / width_pos_tot, "%"),0)
        value <- format(value, big.mark = ",")
        value <- format(value, width = 9, justify = "right")
        bar <- div(
          class = "bar-chart",
          style = list(marginRight = "6px", marginLeft = "6px"),
          div(class = "bar-neg-container",style = list(width = width_neg_percent, borderRight = "1px solid rgba(0, 0, 0, 0.2)"),
            div(class = "bar", style = list(width = width_neg, backgroundColor = "#66c2a4"))),
          div(class = "bar-pos-container",style = list(width = width_pos_percent, borderLeft = "1px solid rgba(0, 0, 0, 0.2)"),
            div(class = "bar", style = list(width = width_pos, backgroundColor = "#006d2c")))
        )
        div(class = "bar-cell", span(class = "number", value), bar)
}

tbl2 <- reactable(
  table2,
  pagination = FALSE,
  defaultSorted = "area_name",
  defaultSortOrder = "asc",
  defaultColDef = colDef(headerClass = "header", align = "left", format = colFormat(separators = TRUE)),
  columns = list(
    area_name = colDef(name = "MSOA", align = "left"),
    latestm12_c = colDef(name = latestm12, maxWidth = 60, align = "right"),
    latest_c = colDef(name = latest, maxWidth = 60, align = "right"),
    additional_claims = colDef(
      name = "Additional claims",
      cell = function(value) {
        cell_function(value,table2$additional_claims)
      }
    ),
    Share = colDef(name = "Share Oct 24", maxWidth = 80, align = "left"),
    percentage_change = colDef(name = "% Change", maxWidth = 80, align = "right"),
    latestm12_r = colDef(name = latestm12, maxWidth = 60, align = "left"),
    latest_r = colDef(name = latest, maxWidth = 60, align = "left"),
    `Change` = colDef(
        cell = function(value) {
          cell_function(value,table2$`Change`)
      }
    )
    ),
  columnGroups = list(
    colGroup(name = "Count", columns = c("latestm12_c", "latest_c", "additional_claims", "Share", "percentage_change")),
    colGroup(name = "Rate", columns = c("latestm12_r", "latest_r", "Change"))
  ),
  compact = TRUE,
  class = "large-table"
)

div(class = "div",
div(class= "large-table tableTitle", "Change in the Claimant Count for Trafford's MSOAs"),
tbl2
)


```

The MSOAs with higher claimant rate in `r latestch` were `r arrange(table2, desc(latest_r))$area_name[1]` with `r arrange(table2, desc(latest_r))$latest_r[1]`%,  `r arrange(table2, desc(latest_r))$area_name[2]` with `r arrange(table2, desc(latest_r))$latest_r[2]`% and  `r arrange(table2, desc(latest_r))$area_name[3]` with `r arrange(table2, desc(latest_r))$latest_r[3]`%. The MSOAs where the claimant rate have decreased more from `r latestm12ch` to `r latestch` were `r arrange(table2, Change)$area_name[1]`, `r arrange(table2, Change)$area_name[2]` and `r arrange(table2, Change)$area_name[2]` with `r arrange(table2, Change)$Change[1]`% points decrease, followed by `r arrange(table2, Change)$area_name[4]` and `r arrange(table2, Change)$area_name[5]` with `r arrange(table2, Change)$Change[5]`%.


  <div class="row">
   <div class = "col-sm-6">

```{r CCMSOADot, fig.height=7}

#, `r arrange(table2, Change)$area_name[2]` with `r arrange(table2, Change)$Change[2]`% followed by `r arrange(table2, Change)$area_name[3]` with `r arrange(table2, Change)$Change[3]`%, `r arrange(table2, Change)$area_name[4]` with `r arrange(table2, Change)$Change[4]`% and `r arrange(table2, Change)$area_name[5]` with `r arrange(table2, Change)$Change[5]`%

#From the top 5 MSOAs with the highest rates, only Partington had a lower rate when comparing `r latestm12ch` to `r latestch`.

df_traff <-  cc_trafford %>%
  mutate(period = as.yearmon(period, format="%B %Y")) %>%
  filter(period %in% c(as.yearmon(latest),as.yearmon(latestm12))) 

# df_traff_plot<- claimant_area %>%
#    mutate(period = as.yearmon(period, format="%b %Y")) %>%
#    bind_rows(df_traff %>% filter(measure == "")) %>%
#   filter(measure=="Claimants as a proportion of residents aged 16-64") %>%
#    mutate(period = factor(period, levels = c(latestm12,latest)),
#          area_name = fct_reorder(factor(area_name), ifelse(period == latest, value, NA), na.rm = TRUE),
#          area_name = fct_relevel(factor(area_name), "Trafford"))

cc_plot_dot <- table2 %>% 
  select (area_name,{{latestm12}}:=latestm12_r, {{latest}}:=latest_r) %>%
  gather(period,value,-one_of("area_name")) %>%
  mutate(period=as.yearmon(period, format="%B %Y")) %>%
  bind_rows(df_traff %>% filter(measure == "Claimants as a proportion of residents aged 16-64") %>% select(area_name,period,value) %>% filter(period %in% c(as.yearmon(latest),as.yearmon(latestm12))) )%>%
   mutate(period = factor(period, levels = c(latestm12,latest)),
         area_name = fct_reorder(factor(area_name), ifelse(period == latest, value, NA), na.rm = TRUE),
         area_name = fct_relevel(factor(area_name), "Trafford"))


ggplot(cc_plot_dot, aes(value,area_name)) +
  geom_line(color = "grey", size=1) +
  geom_point(aes(color = period), size = 4) +
  scale_x_continuous(labels = function(x){ paste0(x, "%") }, limits = c(0, 8.5)) +
  scale_colour_manual(values = c( "#66c2a4", "#006d2c")) +
  labs(x = "percentage of residents aged 16 to 64", y = NULL,
       title = "Claimant Rate", 
       subtitle = paste0("Trafford MSOAs, ", latestm12, " and ", latest), 
       caption = "Source: ONS")+
  theme_minimal() +
  theme(plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
        panel.grid.major.y = element_blank(),
        axis.text.y = element_text(hjust = 1),
        legend.position = "bottom", 
        legend.title = element_blank())

```

   </div>
  
   <div class = "col-sm-6">

```{r CCmsoaMap, fig.height=6}

c_rate_msoa <- st_read("msoa_boundaries.geojson", quiet = TRUE) %>% 
  left_join(., select(c_rate,latest_r,area_code)  , by = "area_code") 


ggplot() +
  geom_sf(data = c_rate_msoa, aes(fill = latest_r), alpha = 1, colour = "#FFFFFF",  size = 0.5) +
  scale_fill_gradientn(colours = c("#edf8fb","#b2e2e2","#66c2a4","#2ca25f","#006d2c"), 
                       na.value = "#f0f0f0",
                       guide = guide_legend(keyheight = unit(3, units = "mm"), 
                                            keywidth=unit(12, units = "mm"), 
                                            label.position = "bottom", 
                                            title = "Rate",
                                            title.position = 'top', 
                                            nrow = 1)) +
  labs(x = NULL, y = NULL, title = "Claimant Rate",
       subtitle = paste("Trafford MSOAs, ", latest),
       caption = "Source: ONS\nContains OS data © Crown copyright and database right 2020") +
  theme(plot.margin = unit(c(0.5,0,1,0), "cm"),
        plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
        axis.text = element_blank(),
        plot.caption = element_text(size=6,margin = margin(t = 55)),
        legend.position = c(0.3,-0.05),
        legend.title = element_text(size = 9),
        legend.text = element_text(size = 8),
        plot.background = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        axis.ticks = element_blank())

``` 

   </div>
  </div> 
 
```{r CCbyAgeTableData}

#Claimant Count
# Source: ONS
# URL: https://www.nomisweb.co.uk/sources/cc
# Licence: Open Government Licence

raw_cc_age <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_162_1.data.csv?geography=1811939363&date=latestMINUS57-latest&gender=0&age=0,2,10,11,3,12...16,4,17...20&measure=1&measures=20100") 


cc_age <- raw_cc_age %>%
  select(period = DATE_NAME,
         area_code = GEOGRAPHY_CODE,
         area_name = GEOGRAPHY_NAME,
         age = AGE_NAME,
         age_type = AGE_TYPE,
         value = OBS_VALUE) %>%
  mutate(period = as.yearmon(period, "%b %Y"),
         Age =  gsub("Aged |All categories: Age ", "", age)) %>%
    select(period, age_type, Age, value) %>%
  spread(period,value) %>%
  mutate(additional_claims= get(latest)-get(latestm12),
         percentage_change = round(additional_claims/get(latestm12) *100,0))

age_bands <- cc_age %>%
  filter(age_type=="Ageband") %>%
  mutate(Age=factor(Age, levels=c("16-24", "25-49", "50+", "16+"))) %>%
  arrange(Age) %>%
  select(-age_type)

quinary <- cc_age %>%
  filter(age_type=="Quinary") %>%
  mutate_if(is.numeric, list(~na_if(., Inf)))

table3 <- quinary %>% select(Age, `latestm12`= latestm12, `latest` = latest, additional_claims, percentage_change)
table4 <- age_bands %>% select(Age, `latestm12`= latestm12, `latest` = latest, additional_claims, percentage_change)

```


In Trafford, from  `r latestm12ch` to `r latestch` 4 out of 11 of the age bands has decreased the number of claimants. The age band `r filter(table3, additional_claims == max(additional_claims)) %>% pull(Age)`  had the most increase with `r filter(table3, additional_claims == max(additional_claims)) %>% pull(additional_claims)` more claimants. When considering the reduction of claims the age band `r filter(table3, additional_claims == min(additional_claims)) %>% pull(Age)` years had the largest reduction with `r filter(table3, additional_claims == min(additional_claims)) %>% pull(additional_claims)` less claims.

  <div class="row">
   <div class = "col-sm-6">

```{r CCbyAgeTable}


tbl3 <- reactable(
  table3,
  pagination = FALSE,
  defaultSorted = "Age",
  defaultSortOrder = "asc",
  defaultColDef = colDef(headerClass = "header", align = "right", format = colFormat(separators = TRUE)),
  columns = list(
    Age = colDef(maxWidth = 50, align = "left"),
    latestm12 = colDef(name = latestm12, maxWidth = 60),
    latest = colDef(name = latest, maxWidth = 60),
    additional_claims = colDef(
      name = "Additional claims",
      cell = function(value) {
        cell_function(value,table3$additional_claims)
      }
    ),
    percentage_change = colDef(
      name = "% Change",
      cell = function(value) {
        cell_function(value,table3$percentage_change)
      }
    )
    ),
  compact = TRUE,
  class = "table"
)

div(class = "div",
div(class= "large-table tableTitle", "Change in the Claimant Count by age bands in Trafford"),
tbl3
)

```

   </div>
  
   <div class = "col-sm-6">

```{r CCbyAgeTable2}

tbl4 <- reactable(
  table4,
  pagination = FALSE,
  defaultSorted = "Age",
  defaultSortOrder = "asc",
  defaultColDef = colDef(headerClass = "header", align = "right", format = colFormat(separators = TRUE)),
  columns = list(
    Age = colDef(name = "Age", maxWidth = 60, align = "left"),
    latestm12 = colDef(name = latestm12, maxWidth = 60),
    latest = colDef(name = latest, maxWidth = 60),
    additional_claims = colDef(
      name = "Additional claims",
      cell = function(value) {
        cell_function(value,table4$additional_claims)
      }
    ),
    percentage_change = colDef(
      name = "% Change",
      cell = function(value) {
        cell_function(value, table4$percentage_change)
      }
    )
    ),
  compact = TRUE,
  class = "table"
)

div(class = "div",
div(class= "large-table tableTitle", "Change in the Claimant Count by broad age bands in Trafford"),
tbl4
)


```

   </div>
  </div>
 
At the start of 2020 the age band 25-34 already had the highest number of claims, however after the effects of the Covid-19 pandemic which started in March 2020 this figure increased more sharply compared to the other age bands. The claimant rate as a proportion of Trafford's population within each age band showed then a higher increase in the age bands of 16-24 and 25-34 compared to the others. From January 2023 the count and rate of all age bands started to increase but the figures are fluctuating from April 2023. The age groups 35 to 44 and 45 to 54 had an increase on July 2024.

  <div class="row">
   <div class = "col-sm-6">

```{r CCAgeMLine}

plotdf <- quinary %>%
  select(-c(age_type, additional_claims, percentage_change)) %>%
  mutate(Age=fct_collapse(Age, `16-24` = c("16-17", "18-24"),
                       `25-34` = c("25-29", "30-34"),
                       `35-44` = c("35-39", "40-44"),
                       `45-54` = c("45-49", "50-54"),
                       `55+` = c("55-59", "60-64","65+"))) %>%
  group_by(Age) %>%
  summarise_all(sum) %>%
  gather(date,value, base:latest) %>%
  mutate(date=as.Date(as.yearmon(date), format="%b %Y")) %>%
  mutate(label = if_else(date == max(date), Age, factor(NA)))
  

 ggplot(plotdf, aes(x = date, y = value, colour = Age)) +
  geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +
  geom_line(size = 1, show.legend = FALSE) +
  geom_text_repel(aes(x = max(date) + 30, label = label, fontface = "bold", hjust = 0), direction = "y", show.legend = FALSE) +
  scale_x_date(breaks = c(min(plotdf$date), as.Date(as.yearmon("Jan 2021"),format="%b %Y"), as.Date(as.yearmon("Jan 2022"),format="%b %Y"),as.Date(as.yearmon("Jan 2023"),format="%b %Y"), as.Date(as.yearmon("Jan 2024"),format="%b %Y")), date_labels = "%b %y") +
  scale_y_continuous(expand = c(0.005, 0.005)) +
  labs(title = "Claimant Count by age band",
       subtitle = paste("Trafford, ", base, " - ",latest),
       caption = "Source: ONS",
       x = NULL,
       y = NULL,
       colour = NULL) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
    plot.margin = unit(rep(0.5,4), "cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks.x = element_line(colour = "#333333", size = 0.5))  +
  coord_cartesian(clip = "off") 


```

   </div>
  
   <div class = "col-sm-6">

```{r CCAgeRate}

# Source:Mid-2019 population estimates for local authorities in England
# URL: https://www.nomisweb.co.uk/datasets/pestsyoala
# Licence: Open Government Licence v3.0


#pop_trafford_age_raw <- read_csv("http://www.nomisweb.co.uk/api/v01/dataset/NM_2002_1.data.csv?geography=1811939363&date=latest&gender=0&c_age=101...191&measures=20100&select=date_name,geography_name,geography_code,gender_name,c_age_name,measures_name,obs_value,obs_status_name")

pop_trafford_age_raw <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_2002_1.data.csv?geography=1778385132&date=latestMINUS3-latest&gender=0&c_age=101...191&measures=20100")


pop_trafford_age <- pop_trafford_age_raw %>%
    select(year = DATE_NAME, age = C_AGE_NAME, n = OBS_VALUE) %>%
    mutate(age = parse_number(age),
          ageband = cut(age,
                        breaks = c(0,15,24,34,44,54,64,120),
                        labels = c("0-15","16-24","25-34","35-44","45-54","55-64","65+"),
                        right = TRUE,
                        include.lowest = TRUE)) %>%
   group_by(ageband,year) %>%
   summarise(population = sum(n))

#write_csv(pop_trafford_age_raw, "pop_trafford_age_raw.csv")

plotdf2 <- quinary %>%
  select(-c(age_type, additional_claims, percentage_change)) %>%
  mutate(Age=fct_collapse(Age, `16-24` = c("16-17", "18-24"),
                       `25-34` = c("25-29", "30-34"),
                       `35-44` = c("35-39", "40-44"),
                       `45-54` = c("45-49", "50-54"),
                       `55-64` = c("55-59", "60-64"),
                       `65+` = "65+")) %>%
  filter(!Age == "65+") %>%
  group_by(Age) %>%
  summarise_all(sum) %>%
  gather(date,value, base:latest) %>%
  mutate(year = as.integer(word(date,-1,-1))) %>%
  left_join(pop_trafford_age, by = c("Age" = "ageband", "year")) %>%
  arrange(Age) %>%
  fill(population) %>%
  mutate(rate = round((value/population)*100,1),
         date=as.Date(as.yearmon(date), format="%b %Y")) %>%
  mutate(label = if_else(date == max(date), Age, factor(NA)))

 ggplot(plotdf2, aes(x = date, y = rate, colour = Age)) +
  geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +
  geom_line(size = 1, show.legend = FALSE) +
  geom_text_repel(aes(x = max(date) + 30, label = label, fontface = "bold", hjust = 0), direction = "y", show.legend = FALSE) +
  #scale_color_startrek() +
  scale_x_date(breaks = c(min(plotdf2$date), as.Date(as.yearmon("Jan 2021"),format="%b %Y"), as.Date(as.yearmon("Jan 2022"),format="%b %Y"),as.Date(as.yearmon("Jan 2023"),format="%b %Y"), as.Date(as.yearmon("Jan 2024"),format="%b %Y")), date_labels = "%b %y") +
  scale_y_continuous(expand = c(0.005, 0.005)) +
  labs(title = "Claimant rate by age band",
       subtitle = paste("Trafford, ", base, " - ",latest),
       caption = "Source: ONS",
       x = NULL,
       y = NULL,
       colour = NULL) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
    plot.margin = unit(rep(0.5,4), "cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks.x = element_line(colour = "#333333", size = 0.5))  +
  coord_cartesian(clip = "off") 

```

   </div>
  </div>

More than half of Trafford's claimants are residents of 1 quarter of the MSOAs. Around half of the claimants are between 25 and 44 years.

  <div class="row">
   <div class = "col-sm-6">

```{r CCMSOAAgeTable}
#Claimant Count
# Source: ONS
# URL: https://www.nomisweb.co.uk/sources/cc
# Licence: Open Government Licence


cc_msoa_age <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_162_1.data.csv?geography=1245709510...1245709537&date=latest&gender=0&age=10...20&measure=1&measures=20100") %>%
  select(period = DATE_NAME, area_code = GEOGRAPHY_CODE, area_code2 = GEOGRAPHY_NAME, ageband = AGE_NAME, n = OBS_VALUE) %>%
    left_join(lookup%>%select(msoa11cd,msoa11nm,msoa11hclnm), by = c("area_code2" = "msoa11nm")) %>%
  rename(area_name = msoa11hclnm) %>%
  mutate(period = as.Date(paste("01", period), format = "%d %b %Y"),
         ageband=gsub("Aged ", "", ageband)) %>%
  mutate(ageband = case_when(
    ageband == "16-17" ~ "16-17",
    ageband == "18-24" ~ "18-24",
    ageband %in% c("25-29","30-34") ~ "25-34",
    ageband %in% c("35-39", "40-44") ~ "35-44",
    ageband %in% c("45-49", "50-54") ~ "45-54",
    ageband %in% c("55-59", "60-64") ~ "55-64",
    TRUE ~ "65+"
  )) %>%
  group_by(period, ageband, area_code, area_name) %>%
  summarise(n = sum(n)) %>%
  ungroup()


cc_msoa_aget<- cc_msoa_age %>%
  select(MSOA=area_name,ageband,n) %>%
  spread(ageband,n)

tbl5 <- reactable(
  cc_msoa_aget,
  pagination = FALSE,
  defaultSorted = "MSOA",
  defaultSortOrder = "asc",
  defaultColDef = colDef(headerClass = "header", align = "left", maxWidth = 45, format = colFormat(separators = TRUE)),
  columns = list(
    MSOA = colDef(maxWidth = 200, align = "left")
    ),
  class = "xsmall-table"
)

div(class = "div",
div(class= "large-table tableTitle", "Claimant Count for Trafford's MSOAs by age band"),
tbl5
)

```

   </div>
  
   <div class = "col-sm-6">

```{r CCMSOAAgeMMK, fig.height=6}

cc_msoa_agep<- cc_msoa_age %>%
  spread(ageband,n) %>%
  mutate("16-24" = `16-17`+`18-24`) %>%
  select(period:area_name,`16-24`,`25-34`:`65+`) %>%
  gather(ageband,n,`16-24`:`65+`) %>%
  group_by(period, area_code, area_name) %>%
  mutate(ageband = fct_rev(ageband),
         total = sum(n), percent = n/total) 

stats <- cc_msoa_agep %>% 
  summarise(totalm =sum(n)) %>%
  ungroup() %>%
  mutate(perM = totalm*100/sum(totalm)) %>%
  arrange(desc(perM)) %>%
  mutate(incre = cumsum(perM))

stats2 <- cc_msoa_agep %>% 
  ungroup() %>%
  group_by(period, ageband) %>%
    summarise(totalm =sum(n)) %>%
  mutate(perM = totalm*100/sum(totalm)) %>%
  arrange(desc(ageband)) %>%
  mutate(incre = cumsum(perM))

ggplot(cc_msoa_agep, aes(x = area_name, y = percent, width = total, fill = ageband)) +
  geom_col(position = "stack", colour = NA) +
  facet_grid(~fct_reorder(area_name,total, .desc = T), scales = "free_x", space = "free_x") +
  scale_y_continuous(expand = c(0.005, 0.005), breaks = c(0, 0.5, 1), labels = percent) +
  scale_fill_brewer(palette = "Set3") +
  labs(x = NULL, y = NULL, 
       title = "Claimant Count by Trafford's MSOAs and age band",
       subtitle = as.yearmon(unique(cc_msoa_agep$period)),
       caption = "The width of the columns is proportional to the number of claimants. Source: DWP.",
       fill = NULL) +
  theme(plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
    panel.spacing.x = unit(0.01, "npc"),
    panel.grid.major = element_blank(),
    strip.text.x = element_blank(),
    legend.position = "right",
    axis.text.x = element_text(size=8, angle = 90, hjust = 1, vjust=0.5),
    panel.background = element_blank(),
    axis.ticks = element_blank())

```

   </div>
  </div>
 </div>
</section>

<section>

 <div class="container-fluid">
 
## Universal Credit claims in Trafford

```{r UCTraffPCHData}

# add your API key
api_key <- ""
# API endpoint
path <- "https://stat-xplore.dwp.gov.uk/webapi/rest/v1/table"

# Source:Mid-2019 population estimates for local authorities in England
# URL: https://www.nomisweb.co.uk/datasets/pestsyoala
# Licence: Open Government Licence v3.0

pop_trafford <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_2002_1.data.csv?geography=1778385132&date=latestMINUS3-latest&gender=0&c_age=203&measures=20100") %>%
    select(area_code = GEOGRAPHY_CODE,
           year = DATE_NAME,
         pop16_64 = OBS_VALUE)

#Universal Credit
# Source: DWP
# URL: https://stat-xplore.dwp.gov.uk/webapi/metadata/UC_Monthly/UC_Monthly.html
# Licence: Open Government Licence

query <- list(database = unbox("str:database:UC_Monthly"),
              measures = "str:count:UC_Monthly:V_F_UC_CASELOAD_FULL",
              dimensions = c("str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE",
                             "str:field:UC_Monthly:F_UC_DATE:DATE_NAME") %>% matrix(),
              recodes = list(
                `str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE` = list(
                  map = list("str:value:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE:V_C_MASTERGEOG11_LA_TO_REGION:E08000009")),
                `str:field:UC_Monthly:F_UC_DATE:DATE_NAME` = list(
                  map = as.list(paste0("str:value:UC_Monthly:F_UC_DATE:DATE_NAME:C_UC_DATE:",c(202001,202002,202003,202004,202005,202006,202007,202008,202009,202010,202011,202012,202101,202102,202103,202104,202105,202106,202107,202108,202109,202110,202111,202112,202201,202202,202203,202204,202205,202206,202207,202208,202209,202210,202211,202212,202301,202302,202303,202304,202305,202306,202307,202308,202309,202310,202311,202312,202401,202402,202403,202404,202405,202406,202407,202408,202409,202410))))
              )) %>% toJSON()
request <- POST(
  url = path,
  body = query,
  config = add_headers(APIKey = api_key),
  encode = "json")
response <- fromJSON(content(request, as = "text"), flatten = TRUE)
# extract list items and convert to a dataframe
tabnames <- response$fields$items %>% map(~.$labels %>% unlist)
values <- response$cubes[[1]]$values
dimnames(values) <- tabnames

trafford_uc <- values %>%
  as.data.frame.table(values, stringsAsFactors = FALSE) %>%
  as_tibble() %>%
  set_names(c(response$fields$label,"value")) %>%
mutate(period=as.yearmon(`Month`, "%B %Y")) %>%
  select(area_name = "National - Regional - LA - OAs",period,value) %>%
  mutate(year = as.integer(word(period,-1,-1))) %>%
  left_join(pop_trafford, by = "year") %>%
  fill(pop16_64) %>%
  mutate(rate= round(value*100/pop_trafford$pop16_64,1),
         change=round((value*100/value[1])-100,1))


```

The [Universal Credit](https://stat-xplore.dwp.gov.uk/webapi/metadata/UC_Monthly/UC_Monthly.html) claims in Trafford have increased `r filter(trafford_uc, period == latest) %>% pull(change)`% from `r basech` to `r latestch`. The rate of claims as a proportion of people age 16 to 64 increased form `r filter(trafford_uc, period == base) %>% pull(rate)`% to `r filter(trafford_uc, period == latest) %>% pull(rate)`% from `r basech` to `r latestch`. Claimants are migrating from legacy benefits to Universal Credit which replace six means-tested benefits for working-age people: Working Tax Credit, Child Tax Credit, Housing Benefit, Income Support, income-based Jobseeker’s Allowance, and income-related Employment band Support Allowance (ESA). 


  <div class="row">
   <div class = "col-sm-4">

```{r UCTrafPCHTable}

table6 <-  trafford_uc %>% 
  mutate(period = as.Date(period,frac=0)) %>%
  select(-area_name, -area_code, -year, -pop16_64)

 tbl6 <- reactable(
  table6,
  pagination = TRUE,
  defaultSorted = "period",
  defaultSortOrder = "desc",
  defaultColDef = colDef(headerClass = "header", align = "right", minWidth = 60, format = colFormat(separators = TRUE)),
  columns = list(
    period = colDef(name = "Date", align = "left", cell = function(value) date_format("%b %Y")(value))
    #area_name = colDef(name = "Area", maxWidth = 100, align = "left")
    ),
  compact = TRUE,
  class = "table"
)

div(class = "div",
div(class= "tableTitle", "Change in the Universal Credit claims for Trafford"),
tbl6
)

```

   </div>
  
   <div class = "col-sm-8">

```{r UCTrafPCHPlot}

uc_plot_ch <- trafford_uc %>% mutate(date=as.Date(period,frac=0)) %>%
  mutate(change = round(change,0)) %>%
  mutate(index = case_when(change == 0 ~ "Label",
                           period == latest ~ "Label",
                           change -lag(change)<=0 & change - lead(change)<0 ~ "Label",
                           change -lag(change)>=0 & change - lead(change)>0 ~ "Label",
                           change -lead(change)<(-20) | change - lag(change)>20 ~ "Label",
                           TRUE ~ "No label"))

ggplot(uc_plot_ch,aes(x = date, y = change)) + 
  geom_line(aes(group=1), size = 1, alpha = 1, color="#006d2c") + 
  geom_point(size = 2, alpha = 1, color="#006d2c") + 
    geom_text(data=uc_plot_ch %>% filter(!period==base),
            aes(label = paste0(change, "%"), color = index),
            fontface = 1.5, size = 2.5, hjust = 0.3, vjust = -1.3)+
  scale_color_manual(values = c("#006d2c", "transparent")) +
  scale_x_date(breaks = seq(as.Date("2020-03-01"), as.Date(max(uc_plot_ch$period)), by="4 months"), date_labels="%b %y", expand = c(0.03,0.03)) +
  scale_y_continuous(limits = c(0,115)) +
  labs(title = "Universal Credit claims percentage change",
       subtitle = paste0("Trafford, ", base, " - ", latest),
       caption = "Source: DWP",
       x = NULL,
       y = NULL,
       colour = NULL) +
  #theme_lab() +
  theme_minimal() +
  theme(plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
        panel.grid.minor = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        axis.text.y = element_blank(),
        legend.position = "none")

```

   </div>
  </div>
 
```{r UCMSOATableData}
 



#Universal Credit
# Source: DWP
# URL: https://stat-xplore.dwp.gov.uk/webapi/metadata/UC_Monthly/UC_Monthly.html
# Licence: Open Government Licence


query <- list(database = unbox("str:database:UC_Monthly"),
              measures = "str:count:UC_Monthly:V_F_UC_CASELOAD_FULL",
              dimensions = c("str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE",
                             "str:field:UC_Monthly:F_UC_DATE:DATE_NAME") %>% matrix(),
              recodes = list(
                `str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE` = list(
                  map = as.list(paste0("str:value:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE:V_C_MASTERGEOG11_MSOA_TO_LA:E0", seq(2001259, 2001286, 1)))),
                `str:field:UC_Monthly:F_UC_DATE:DATE_NAME` = list(
                  map = as.list(paste0("str:value:UC_Monthly:F_UC_DATE:DATE_NAME:C_UC_DATE:",c(202310,202410))))
              )) %>% toJSON()
request <- POST(
  url = path,
  body = query,
  config = add_headers(APIKey = api_key),
  encode = "json")
response <- fromJSON(content(request, as = "text"), flatten = TRUE)
# extract list items and convert to a dataframe
tabnames <- response$fields$items %>% map(~.$labels %>% unlist)
values <- response$cubes[[1]]$values
dimnames(values) <- tabnames

uc <- as.data.frame.table(values, stringsAsFactors = FALSE) %>%
  as_tibble() %>%
  set_names(c(response$fields$label,"value")) %>%
  left_join(lookup%>%select(msoa11cd,msoa11nm,msoa11hclnm), by = c("National - Regional - LA - OAs" = "msoa11nm")) %>%
  mutate(period=format(as.yearmon(`Month`, "%B %Y"),"%b %Y")) %>%
  select(area_code=msoa11cd, area_name = msoa11hclnm, period, value) %>%
  filter(period %in% c(latestm12,latest)) %>%
  spread(period,value) %>%
  mutate(additional_claims= get(latest)-get(latestm12),
        percentage_change = round(additional_claims/get(latestm12) *100,0),
        `Share` = round(get(latest)/sum(get(latest)) *100,1)) %>%
  left_join(msoa_pop, by="area_code") %>%
  mutate(latestm12_rate = round(get(latestm12)*100/pop16_64,1), latest_rate=round(get(latest)*100/pop16_64,1),Change=round(latest_rate-latestm12_rate,1))

table7 <- uc %>%
  select(-c(area_code,pop16_64))
 
```
 
The Trafford MSOA with the largest share of claims in `r latestch` was `r filter(table7, Share == max(Share)) %>% pull(area_name)` with `r filter(table7, Share == max(Share))$Share`% of all Universal Credit claims in Trafford. 27 out of 28 MSOAs have more claims in `r latestch`  compared to `r latestm12ch`. The MSOAs with more additional claims when comparing `r latestm12ch` to `r latestch` were `r arrange(table7,desc(additional_claims))$area_name[1]` (`r arrange(table7,desc(additional_claims))$additional_claims[1]`), `r arrange(table7,desc(additional_claims))$area_name[2]` (`r arrange(table7, desc(additional_claims))$additional_claims[2]`) and `r arrange(table7, desc(additional_claims))$area_name[3]` (`r arrange(table7,desc(additional_claims))$additional_claims[3]`). The MSOAs with more increase in percentage change from `r latestm12ch` to `r latestch` were `r arrange(table7, desc(percentage_change))$area_name[1]` with `r arrange(table7, desc(percentage_change))$percentage_change[1]`% change and `r arrange(table7, desc(percentage_change))$area_name[2]` with `r arrange(table7, desc(percentage_change))$percentage_change[2]`% change.


```{r UCMSOATable}

#The MSOAs that increased the number of claims from `r latestm12ch` to `r latestch` were `r arrange(table7, desc(additional_claims))$area_name[1]` with `r arrange(table7, desc(additional_claims))$additional_claims[1]`  and `r arrange(table7, desc(additional_claims))$area_name[2]` with `r arrange(table7, desc(additional_claims))$additional_claims[2]`.

#The MSOA with reductions in claims when comparing `r latestm12ch` to `r latestch` is `r arrange(table7,additional_claims)$area_name[1]` (`r arrange(table7,additional_claims)$additional_claims[1]`). The MSOA with reduction in percentage change from `r latestm12ch` to `r latestch` was `r arrange(table7, percentage_change)$area_name[1]` with `r arrange(table7, percentage_change)$percentage_change[1]`% change.

tbl7 <- reactable(
  table7,
  pagination = FALSE,
  defaultSorted = "area_name",
  defaultSortOrder = "asc",
  defaultColDef = colDef(headerClass = "header", align = "right", maxWidth = 60, format = colFormat(separators = TRUE)),
  columns = list(
    area_name = colDef(name = "MSOA", maxWidth = 220, align = "left"),
    additional_claims = colDef(
      name = "Additional claims",
      maxWidth = 175,
      align = "left",
      cell = function(value) {
        cell_function(value,table7$additional_claims)
      }
    ),
    percentage_change = colDef(name = "% Change", maxWidth = 80),
    Share = colDef(name = "Share Oct 24", maxWidth = 80, align = "left"),
    latestm12_rate = colDef(name = latestm12, maxWidth = 60, align = "left"),
    latest_rate = colDef(name = latest, maxWidth = 60, align = "left"),
    `Change` = colDef(
      maxWidth = 175,
      cell = function(value) {
        cell_function(value,table7$Change)
      }
    )
    ),
  columnGroups = list(
    colGroup(name = "Count", columns = c(latestm12, latest, "additional_claims", "percentage_change", "Share")),
    colGroup(name = "Rate", columns = c("latestm12_rate", "latest_rate", "Change"))
  ),
  compact = TRUE,
  class = "large-table"
)

div(class = "div",
div(class= "large-table tableTitle", "Change in the Universal Credit claims for Trafford's MSOAs"),
tbl7
)

```

The MSOAs with higher rates of Universal Credit claims in `r latestch` were `r filter(table7, latest_rate == max(latest_rate)) %>% pull(area_name)` with `r filter(table7, latest_rate == max(latest_rate))$latest_rate`% and `r arrange(table7, desc(latest_rate))$area_name[2]` with `r arrange(table7, desc(latest_rate))$latest_rate[2]`%. The MSOAs where the rate of Universal Credit claims has increased more from `r latestm12ch` to `r latestch` were `r arrange(table7, desc(Change))$area_name[1]` with `r arrange(table7, desc(Change))$Change[1]`% points increase and `r arrange(table7, desc(Change))$area_name[2]` with `r arrange(table7, desc(Change))$Change[2]`% points increase.

  <div class="row">
  <div class = "col-sm-6">

```{r UCMSOADotPlot, fig.height=7}

# The MSOA where the rate of Universal Credit claims has decreased from `r latestm12ch` to `r latestch` was `r arrange(table7, Change)$area_name[1]` with `r arrange(table7, Change)$Change[1]`%.

uc_plot_dot <- table7 %>% 
  select (area_name,{{latestm12}}:=latestm12_rate, {{latest}}:=latest_rate) %>%
  gather(period,value,-one_of("area_name")) %>%
  mutate(period=as.yearmon(period, format="%B %Y")) %>%
  bind_rows(trafford_uc %>% select(area_name,period,value=rate) %>% filter(period %in% c(as.yearmon(latest),as.yearmon(latestm12))) )%>%
   mutate(period = factor(period, levels = c(latestm12,latest)),
         area_name = fct_reorder(factor(area_name), ifelse(period == latest, value, NA), na.rm = TRUE),
         area_name = fct_relevel(factor(area_name), "Trafford"))

ggplot(uc_plot_dot, aes(value,area_name)) +
  geom_line(color = "grey", size=0.5) +
  geom_point(aes(color = period), size = 3) +
  scale_x_continuous(labels = function(x){ paste0(x, "%") }, limits = c(0, 35)) +
  scale_colour_manual(values = c( "#66c2a4", "#006d2c")) +
  labs(x = "percentage of residents aged 16 to 64", y = NULL,
       title = "Universal Credit claims rate", 
       subtitle = paste0("Trafford MSOAs, ", latestm12, " and ", latest), 
       caption = "Source: DWP")+
  theme_minimal() +
  theme(plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
        panel.grid.major.y = element_blank(),
        axis.text.y = element_text(hjust = 1),
        legend.position = "bottom", 
        legend.title = element_blank())

```

   </div>
  
   <div class = "col-sm-6">

```{r UCMSOAMap, fig.height=6}

msoa_uc <- st_read("msoa_boundaries.geojson", quiet = TRUE) %>% 
  left_join(., select(uc,latest_rate,area_code), by = "area_code")

ggplot() +
  geom_sf(data = msoa_uc, aes(fill = latest_rate), alpha = 1, colour = "#FFFFFF",  size = 0.5) +
  scale_fill_gradientn(colours = c("#edf8fb","#b2e2e2","#66c2a4","#2ca25f","#006d2c"), 
                       na.value = "#f0f0f0",
                       guide = guide_legend(keyheight = unit(3, units = "mm"), 
                                            keywidth=unit(12, units = "mm"), 
                                            label.position = "bottom", 
                                            title = "Rate",
                                            title.position = 'top', 
                                            nrow = 1)) +
  labs(x = NULL, y = NULL, title = "Universal Credit claims rate",
       subtitle = paste0("Trafford MSOAs, ", latest),
       caption = "Source: DWP\nContains OS data © Crown copyright and database right 2020") +
  theme(plot.margin = unit(c(0.5,0,1,0), "cm"),
        plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
        axis.text = element_blank(),
        plot.caption = element_text(size=6,margin = margin(t = 55)),
        legend.position = c(0.3,-0.05),
        legend.title = element_text(size = 9),
        legend.text = element_text(size = 8),
        plot.background = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        axis.ticks = element_blank())


```

   </div>
  </div> 

```{r UCAgeData}
#Universal Credit
# Source: DWP
# URL: https://stat-xplore.dwp.gov.uk/webapi/metadata/UC_Monthly/UC_Monthly.html
# Licence: Open Government Licence


query <- list(database = unbox("str:database:UC_Monthly"),
              measures = "str:count:UC_Monthly:V_F_UC_CASELOAD_FULL",
              dimensions = c("str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE",
                             "str:field:UC_Monthly:F_UC_DATE:DATE_NAME",
                             "str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:AGE_CODE") %>% matrix(),
              recodes = list(
                `str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE` = list(
                  map = list("str:value:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE:V_C_MASTERGEOG11_LA_TO_REGION:E08000009")),
                `str:field:UC_Monthly:F_UC_DATE:DATE_NAME` = list(
                  map = as.list(paste0("str:value:UC_Monthly:F_UC_DATE:DATE_NAME:C_UC_DATE:",c(202001,202002,202003,202004,202005,202006,202007,202008,202009,202010,202011,202012,202101,202102,202103,202104,202105,202106,202107,202108,202109,202110,202111,202112,202201,202202,202203,202204,202205,202206,202207,202208,202210,202211,202212,202301,202302,202303,202304,202305,202306,202307,202308,202309,202310,202311,202312,202401,202402,202403,202404,202405,202406,202407,202408,202409,202410)))),
                `str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:AGE_CODE` = list(
                  map = as.list(paste0("str:value:UC_Monthly:V_F_UC_CASELOAD_FULL:AGE_CODE:C_UC_AGE_BAND:",c(1,2,3,4,5,6,7,8,9,10,999))))

              )) %>% toJSON()
request <- POST(
  url = path,
  body = query,
  config = add_headers(APIKey = api_key),
  encode = "json")
response <- fromJSON(content(request, as = "text"), flatten = TRUE)
# extract list items and convert to a dataframe
tabnames <- response$fields$items %>% map(~.$labels %>% unlist)
values <- response$cubes[[1]]$values
dimnames(values) <- tabnames

traf_uc <- as.data.frame.table(values, stringsAsFactors = FALSE) %>%
  as_tibble() %>%
  set_names(c(response$fields$label,"value")) %>%
  mutate(indicator="People on Universal Credit",
         period=format(as.yearmon(`Month`, "%B %Y"),"%b %Y")) %>%
  select(area_name = `National - Regional - LA - OAs`, indicator, Age="Age (bands and single year)", period, value)


table8 <- traf_uc %>%
  select(-c(area_name,indicator)) %>%
  filter(period %in% c(latestm12,latest)) %>%
  spread(period,value) %>%
  mutate(additional_claims= get(latest)- get(latestm12),
         percentage_change = round(additional_claims/get(latestm12) *100,0)) %>%
  select(Age, `latestm12`= latestm12, `latest` = latest, additional_claims, percentage_change)
```
 
In Trafford from `r latestm12ch` to `r latestch` the `r filter(table8, percentage_change == max(percentage_change)) %>% pull(Age)` age band has the largest increment in percentage change of Universal Credit claims with `r filter(table8, percentage_change == max(percentage_change)) %>% pull(percentage_change)`%. When considering the number of additional claims from  `r latestm12ch` to `r latestch` the 5-year age band `r filter(table8, additional_claims == max(additional_claims)) %>% pull(Age)` years had the highest with `r filter(table8, additional_claims == max(additional_claims)) %>% pull(additional_claims)` more claims. All age bands but the over 65 had an increment of claims compared to `r latestm12ch`.
 
```{r UCAgeTable}

#The age bands `r arrange(table8, additional_claims)$Age[1]` and `r arrange(table8, additional_claims)$Age[2]` have the largest decrease in number of Universal Credit claims. Claimants  under 30 decreased or not changed in number whilst claimants 30 and over increased in number when comparing `r latestm12ch` to `r latestch`.

#The age bands `r arrange(table8, additional_claims)$Age[1]` had a decrease in number of Universal Credit claims.

tbl8 <- reactable(
  table8,
  pagination = FALSE,
  defaultSorted = "Age",
  defaultSortOrder = "asc",
  defaultColDef = colDef(headerClass = "header", align = "right", maxWidth = 60, format = colFormat(separators = TRUE)),
  columns = list(
    Age = colDef(name = "Age", maxWidth = 60, align = "left"),
    latestm12 = colDef(name = latestm12, maxWidth = 60),
    latest = colDef(name = latest, maxWidth = 60),
    additional_claims = colDef(
      name = "Additional claims",
      maxWidth = 175,
      cell = function(value) {
        cell_function(value,table8$additional_claims)
      }
    ),
    percentage_change = colDef(
      name = "% Change",
      maxWidth = 175,
      cell = function(value) {
        cell_function(value,table8$percentage_change)
      }
    )
    ),
  compact = TRUE,
  class = "large-table"
)

div(class = "div",
div(class= "large-table tableTitle", "Change in the Universal Credit claims by age band for Trafford"),
tbl8
)

```


Following the start of the Covid-19 pandemic in March 2020 the number of claims for all ages rose dramatically from pre-pandemic levels at a similar proportion for all age bands with the age band 16-24 rising faster in May 2020. At the start of 2020 the age band 25-34 had the highest number of claims and also the highest rate maintaining that position after the rise from March 2020 until May 2022 when the 35-44 became the age band with the highest number of claims. The age band 16-24 has the lowest number of claims among the other age bands from July 2021. All the age bands are increasing their number of claims particularly from January 2023.

  <div class="row">
   <div class = "col-sm-6">
  
```{r UCAgeMline}

uc_plot_age_c <- traf_uc %>%
  select(-c(area_name, indicator)) %>%
  mutate(Age=fct_collapse(Age, `16-24` = c("16-19", "20-24"),
                       `25-34` = c("25-29", "30-34"),
                       `35-44` = c("35-39", "40-44"),
                       `45-54` = c("45-49", "50-54"),
                       `55+` = c("55-59", "60-65","Over 65"))) %>%
  group_by(Age,period) %>%
  summarise_all(sum) %>%
  mutate(date=as.Date(as.yearmon(period), format="%b %Y")) %>%
  mutate(label = if_else(date == max(date), Age, factor(NA)))
  

 ggplot(uc_plot_age_c, aes(x = date, y = value, colour = Age)) +
  geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +
  geom_line(size = 1, show.legend = FALSE) +
  geom_text_repel(aes(x = max(date) + 30, label = label, fontface = "bold", hjust = 0), direction = "y", show.legend = FALSE) +
  #scale_color_startrek() +
  scale_x_date(breaks = c(min(uc_plot_age_c$date), as.Date(as.yearmon("Jan 2021"),format="%b %Y"), as.Date(as.yearmon("Jan 2022"),format="%b %Y"), as.Date(as.yearmon("Jan 2023"),format="%b %Y"),as.Date(as.yearmon("Jan 2024"),format="%b %Y")), date_labels = "%b %y") +
  scale_y_continuous(expand = c(0.005, 0.005)) +
  labs(title = "Universal Credit claims by age band",
       subtitle = paste("Trafford, ", base, " - ",latest),
       caption = "Source: DWP",
       x = NULL,
       y = NULL,
       colour = NULL) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
    plot.margin = unit(rep(0.5,4), "cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks.x = element_line(colour = "#333333", size = 0.5))  +
  coord_cartesian(clip = "off") 


```

   </div>
  
   <div class = "col-sm-6">
  
```{r UCAgeMlineRate}
pop_trafford_age2 <- pop_trafford_age_raw %>%
    select(year = DATE_NAME, age = C_AGE_NAME, n = OBS_VALUE) %>%
    mutate(age = parse_number(age),
          ageband = cut(age,
                        breaks = c(0,15,24,34,44,54,65,120),
                        labels = c("0-15","16-24","25-34","35-44","45-54","55-65","Over 65"),
                        right = TRUE,
                        include.lowest = TRUE)) %>%
   group_by(ageband,year) %>%
   summarise(population = sum(n))

pop_trafford_age <- pop_trafford_age_raw %>%
    select(year = DATE_NAME, age = C_AGE_NAME, n = OBS_VALUE) %>%
    mutate(age = parse_number(age),
          ageband = cut(age,
                        breaks = c(0,15,24,34,44,54,64,120),
                        labels = c("0-15","16-24","25-34","35-44","45-54","55-64","65+"),
                        right = TRUE,
                        include.lowest = TRUE))%>%
   group_by(ageband,year) %>%
   summarise(population = sum(n))


uc_plot_age_r <- traf_uc %>%
  select(-c(area_name, indicator)) %>%
  mutate(Age=fct_collapse(Age, `16-24` = c("16-19", "20-24"),
                       `25-34` = c("25-29", "30-34"),
                       `35-44` = c("35-39", "40-44"),
                       `45-54` = c("45-49", "50-54"),
                       `55-65` = c("55-59", "60-65"),
                       `Over 65` = "Over 65")) %>%
  filter(!Age == "Over 65") %>%
  group_by(Age,period) %>%
  summarise_all(sum) %>%
  mutate(year = as.integer(word(period,-1,-1))) %>%
  left_join(pop_trafford_age2, by = c("Age" = "ageband", "year")) %>%
  arrange(Age) %>%
  fill(population) %>%
  #left_join(pop_trafford_age2, by = c("Age" = "ageband")) %>%
  mutate(rate = round((value/population)*100,1),
         date=as.Date(as.yearmon(period), format="%b %Y")) %>%
  mutate(label = if_else(date == max(date), Age, factor(NA)))
  

 ggplot(uc_plot_age_r, aes(x = date, y = rate, colour = Age)) +
  geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +
  geom_line(size = 1, show.legend = FALSE) +
  geom_text_repel(aes(x = max(date) + 30, label = label, fontface = "bold", hjust = 0), direction = "y", show.legend = FALSE) +
  #scale_color_startrek() +
  scale_x_date(breaks = c(min(uc_plot_age_r$date), as.Date(as.yearmon("Jan 2021"),format="%b %Y"), as.Date(as.yearmon("Jan 2022"),format="%b %Y"), as.Date(as.yearmon("Jan 2023"),format="%b %Y"),as.Date(as.yearmon("Jan 2024"),format="%b %Y")), date_labels = "%b %y") +
  scale_y_continuous(expand = c(0.005, 0.005)) +
  labs(title = "Universal Credit claims rate by age band",
       subtitle = paste("Trafford, ", base, " - ",latest),
       caption = "Source: DWP",
       x = NULL,
       y = NULL,
       colour = NULL) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(size=16,face = "bold"),
        plot.subtitle = element_text(size=14),
    plot.margin = unit(rep(0.5,4), "cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks.x = element_line(colour = "#333333", size = 0.5))  +
  coord_cartesian(clip = "off") 

```

   </div>
  </div>
 
```{r UCConditionData}

#Universal Credit
# Source: DWP
# URL: https://stat-xplore.dwp.gov.uk/webapi/metadata/UC_Monthly/UC_Monthly.html
# Licence: Open Government Licence

query <- list(database = unbox("str:database:UC_Monthly"),
              measures = "str:count:UC_Monthly:V_F_UC_CASELOAD_FULL",
              dimensions = c("str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE",
                             "str:field:UC_Monthly:F_UC_DATE:DATE_NAME",
                             "str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:CCCONDITIONALITY_REGIME") %>% matrix(),
              recodes = list(
                `str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE` = list(
                  map = as.list(paste0("str:value:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE:V_C_MASTERGEOG11_MSOA_TO_LA:E0", seq(2001259, 2001286, 1)))),
                `str:field:UC_Monthly:F_UC_DATE:DATE_NAME` = list(
                  map = as.list(paste0("str:value:UC_Monthly:F_UC_DATE:DATE_NAME:C_UC_DATE:",c(202410)))),
                `str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:CCCONDITIONALITY_REGIME` = list(
                  map = as.list(paste0("str:value:UC_Monthly:V_F_UC_CASELOAD_FULL:CCCONDITIONALITY_REGIME:C_UC_CONDITIONALITY_REGIME:",c("AA","AB","BC","BD","DF","CE"))))

              )) %>% toJSON()
request <- POST(
  url = path,
  body = query,
  config = add_headers(APIKey = api_key),
  encode = "json")
response <- fromJSON(content(request, as = "text"), flatten = TRUE)
# extract list items and convert to a dataframe
tabnames <- response$fields$items %>% map(~.$labels %>% unlist)
values <- response$cubes[[1]]$values
dimnames(values) <- tabnames

uc_condition <- as.data.frame.table(values, stringsAsFactors = FALSE) %>%
  as_tibble() %>%
  set_names(c(response$fields$label,"value")) %>%
  left_join(lookup%>%select(msoa11cd,msoa11nm,msoa11hclnm), by = c("National - Regional - LA - OAs" = "msoa11nm")) %>%
  mutate(period=format(as.yearmon(`Month`, "%B %Y"),"%b %Y")) %>%
  select(area_code=msoa11cd, area_name = msoa11hclnm, condition = `Conditionality Regime`, period, value)

uc_msoa_cond<- uc_condition %>%
  group_by(period, area_code, area_name) %>%
  mutate(condition = fct_rev(condition),
         total = sum(value), percent = value/total) 

```

The [Conditionality](https://stat-xplore.dwp.gov.uk/webapi/metadata/UC_Monthly/Conditionality%20Regime.html) regimen for entitlement to Universal Credit is associated to work-related things that claimants will have to do to maintain eligibility. In Trafford the largest proportion of Universal Credit claims are in the "No work requirements" category however there is also a large proportion of claims under "Searching for work". In Old Trafford, 38% of Universal Credit claimants were under "No work requirements" whilst 30% where "Searching for work". In Partington, 49% of claims were under "No work requirements" whilst 16% of claims were under "Working - no requirements", and 17% under "Searching for work".

```{r UCConditionMMK, fig.height=5.5,fig.width=7}

# plot data ---------------------------

ggplot(uc_msoa_cond, aes(x = area_name, y = percent, width = total, fill = condition)) +
  geom_col(position = "stack", colour = NA) +
  facet_grid(~fct_reorder(area_name,total, .desc = T), scales = "free_x", space = "free_x") +
  scale_y_continuous(expand = c(0.005, 0.005), breaks = c(0, 0.5, 1), labels = percent) +
  scale_fill_brewer(palette = "Set3") +
  labs(x = NULL, y = NULL, 
       title = "Universal Credit Claims by Trafford's MSOAs and Conditionality regimen",
       subtitle = unique(uc_msoa_cond$period),
       caption = "The width of the columns is proportional to the number of claims. Source: DWP.",
       fill = NULL) +
  theme(plot.title = element_text(size=12,face = "bold"),
        plot.subtitle = element_text(size=10),
    panel.spacing.x = unit(0.007, "npc"),
    panel.grid.major = element_blank(),
    strip.text.x = element_blank(),
    legend.position = "right",
    axis.text.x = element_text(size=8, angle = 90, hjust = 1, vjust=0.5),
    panel.background = element_blank(),
    axis.ticks = element_blank())

```


 
```{r UCConditionDataTrend}

#Universal Credit
# Source: DWP
# URL: https://stat-xplore.dwp.gov.uk/webapi/metadata/UC_Monthly/UC_Monthly.html
# Licence: Open Government Licence

query <- list(database = unbox("str:database:UC_Monthly"),
              measures = "str:count:UC_Monthly:V_F_UC_CASELOAD_FULL",
              dimensions = c("str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE",
                             "str:field:UC_Monthly:F_UC_DATE:DATE_NAME",
                             "str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:CCCONDITIONALITY_REGIME") %>% matrix(),
              recodes = list(
                `str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE` = list(
                  map = list("str:value:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE:V_C_MASTERGEOG11_LA_TO_REGION:E08000009")),
                `str:field:UC_Monthly:F_UC_DATE:DATE_NAME` = list(
                  map = as.list(paste0("str:value:UC_Monthly:F_UC_DATE:DATE_NAME:C_UC_DATE:",c(202001,202002,202003,202004,202005,202006,202007,202008,202009,202010,202011,202012,202101,202102,202103,202104,202105,202106,202107,202108,202109,202110,202111,202112,202201,202202,202203,202204,202205,202206,202207,202208,202210,202211,202212,202301,202302,202303,202304,202305,202306,202307,202308,202309,202310,202311,202312,202401,202402,202403,202404,202405,202406,202407,202408,202409,202410)))),
                `str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:CCCONDITIONALITY_REGIME` = list(
                  map = as.list(paste0("str:value:UC_Monthly:V_F_UC_CASELOAD_FULL:CCCONDITIONALITY_REGIME:C_UC_CONDITIONALITY_REGIME:",c("AA","AB","BC","BD","DF","CE"))))

              )) %>% toJSON()
request <- POST(
  url = path,
  body = query,
  config = add_headers(APIKey = api_key),
  encode = "json")

response <- fromJSON(content(request, as = "text"), flatten = TRUE)
# extract list items and convert to a dataframe
tabnames <- response$fields$items %>% map(~.$labels %>% unlist)
values <- response$cubes[[1]]$values
dimnames(values) <- tabnames

uc_condition_trend <- as.data.frame.table(values, stringsAsFactors = FALSE) %>%
  as_tibble() %>%
  set_names(c(response$fields$label,"value")) %>%
  mutate(period=format(as.yearmon(`Month`, "%B %Y"),"%b %Y")) %>%
  select(area_name = "National - Regional - LA - OAs", condition = `Conditionality Regime`, period, value) %>%
  group_by(period) %>%
  mutate(Percent = round(value *100/sum(value),1)) %>%
  ungroup()

```

Claimants moving across from legacy benefits to Universal credit and new applications are increasing the number of claimants with no work requirements. Claimants with a disability or health condition which prevents them from working or if they provide care for at least 35 hours a week are not be asked to work. In `r latestch`, `r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$Percent[1]`% (`r format((uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$value[1], big.mark = ",")`) of claims in Trafford were under "`r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$condition[1]`", `r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$Percent[2]`% (`r format((uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$value[2], big.mark=",")`) were under "`r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$condition[2]`", `r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$Percent[3]`% (`r format((uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$value[3], big.mark=",")`) were under "`r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$condition[3]`", `r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$Percent[4]`% (`r format((uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$value[4], big.mark = ",")`) were under "`r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$condition[4]`", `r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$Percent[5]`% (`r format((uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$value[5],big.mark=",")`) were under "`r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$condition[5]`", and `r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$Percent[6]`% (`r format((uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$value[6], big.mark = ",")`) where under "`r (uc_condition_trend %>%filter(period == latest) %>% arrange(desc(Percent)))$condition[6]`".
<div class="row">
<div class = "col-sm-12">
   
```{r UCConditionalityLine, fig.height=3.5, fig.width=7}

uc_plot_conditionality_c <- uc_condition_trend %>%
  mutate(date=as.Date(as.yearmon(period), format="%b %Y")) %>%
  mutate(condition = fct_rev(condition)) 
  

 ggplot(uc_plot_conditionality_c , aes(x = date, y = value, colour = condition)) +
  geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +
  geom_line(size = 1) +
  #geom_text_repel(aes(x = max(date) + 60, label = label, fontface = "bold", hjust = 1), direction = "y", show.legend = FALSE) +
  #scale_color_startrek() +
   scale_color_brewer(palette = "Set3") +
  scale_x_date(breaks = c(min(uc_plot_conditionality_c$date), as.Date(as.yearmon("Jan 2021"),format="%b %Y"), as.Date(as.yearmon("Jan 2022"),format="%b %Y"),  as.Date(as.yearmon("Jan 2023"),format="%b %Y"), as.Date(as.yearmon("Jan 2024"),format="%b %Y")), date_labels = "%b %y") +
  scale_y_continuous(expand = c(0.005, 0.005)) +
  labs(title = "Universal Credit claims by conditionality",
       subtitle = paste("Trafford, ", base, " - ",latest),
       caption = "Source: DWP",
       x = NULL,
       y = NULL,
       colour = NULL) +
  theme_minimal(base_size = 10) +
  theme(plot.title = element_text(size=12,face = "bold"),
        plot.subtitle = element_text(size=10),
    plot.margin = unit(rep(0.5,4), "cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks.x = element_line(colour = "#333333", size = 0.5))  +
  coord_cartesian(clip = "off") 

```

</div>
</div>

```{r UCEntitlementsDataTrend}

#Universal Credit
# Source: DWP
# URL: https://stat-xplore.dwp.gov.uk/webapi/metadata/UC_Monthly/UC_Monthly.html
# Licence: Open Government Licence

query <- list(database = unbox("str:database:UC_Households"),
              measures = "str:count:UC_Households:V_F_UC_HOUSEHOLDS",
              dimensions = c("str:field:UC_Households:V_F_UC_HOUSEHOLDS:COA_CODE",
                             "str:field:UC_Households:F_UC_HH_DATE:DATE_NAME",
                             "str:field:UC_Households:V_F_UC_HOUSEHOLDS:HCLCW_ENTITLEMENT") %>% matrix(),
              recodes = list(
                `str:field:UC_Households:V_F_UC_HOUSEHOLDS:COA_CODE` = list(
                  map = list("str:value:UC_Households:V_F_UC_HOUSEHOLDS:COA_CODE:V_C_MASTERGEOG11_LA_TO_REGION:E08000009")),
                `str:field:UC_Households:F_UC_HH_DATE:DATE_NAME` = list(
                  map = as.list(paste0("str:value:UC_Households:F_UC_HH_DATE:DATE_NAME:C_UC_HH_DATE:",c(202001,202002,202003,202004,202005,202006,202007,202008,202009,202010,202011,202012,202101,202102,202103,202104,202105,202106,202107,202108,202109,202110,202111,202112,202201,202202,202203,202204,202205,202206,202207,202208,202210,202211,202212,202301,202302,202303,202304,202305,202306,202307,202308,202309,202310,202311,202312,202401,202402,202403,202404,202405)))),
                `str:field:UC_Households:V_F_UC_HOUSEHOLDS:HCLCW_ENTITLEMENT` = list(
                  map = as.list(paste0("str:value:UC_Households:V_F_UC_HOUSEHOLDS:HCLCW_ENTITLEMENT:C_UC_LCW_ENTITLEMENT:",c(0,1,2,3))))

              )) %>% toJSON()


request <- POST(
  url = path,
  body = query,
  config = add_headers(APIKey = api_key),
  encode = "json")
response <- fromJSON(content(request, as = "text"), flatten = TRUE)
# extract list items and convert to a dataframe
tabnames <- response$fields$items %>% purrr::map(~.$labels %>% unlist)
values <- response$cubes[[1]]$values
dimnames(values) <- tabnames

ent_wa <- as.data.frame.table(values, stringsAsFactors = FALSE) %>%
  as_tibble() %>%
  set_names(c(response$fields$label,"value"))  %>%
   mutate(indicator="Limited Capability for Work Entitlement") %>%
  rename(category = `Limited Capability for Work Entitlement`)

query <- list(database = unbox("str:database:UC_Households"),
              measures = "str:count:UC_Households:V_F_UC_HOUSEHOLDS",
              dimensions = c("str:field:UC_Households:V_F_UC_HOUSEHOLDS:COA_CODE",
                             "str:field:UC_Households:F_UC_HH_DATE:DATE_NAME",
                             "str:field:UC_Households:V_F_UC_HOUSEHOLDS:TENURE") %>% matrix(),
              recodes = list(
                `str:field:UC_Households:V_F_UC_HOUSEHOLDS:COA_CODE` = list(
                  map = list("str:value:UC_Households:V_F_UC_HOUSEHOLDS:COA_CODE:V_C_MASTERGEOG11_LA_TO_REGION:E08000009")),
                `str:field:UC_Households:F_UC_HH_DATE:DATE_NAME` = list(
                  map = as.list(paste0("str:value:UC_Households:F_UC_HH_DATE:DATE_NAME:C_UC_HH_DATE:",c(202001,202002,202003,202004,202005,202006,202007,202008,202009,202010,202011,202012,202101,202102,202103,202104,202105,202106,202107,202108,202109,202110,202111,202112,202201,202202,202203,202204,202205,202206,202207,202208,202210,202211,202212,202301,202302,202303,202304,202305,202306,202307,202308,202309,202310,202311,202312,202401,202402,202403,202404,202405)))),
                `str:field:UC_Households:V_F_UC_HOUSEHOLDS:TENURE` = list(
                  map = as.list(paste0("str:value:UC_Households:V_F_UC_HOUSEHOLDS:TENURE:C_UC_HOUSING_ENTITLEMENT:",c(0,1))))
                
              )) %>% toJSON()
request <- POST(
  url = path,
  body = query,
  config = add_headers(APIKey = api_key),
  encode = "json")
response <- fromJSON(content(request, as = "text"), flatten = TRUE)
# extract list items and convert to a dataframe
tabnames <- response$fields$items %>% purrr::map(~.$labels %>% unlist)
values <- response$cubes[[1]]$values
dimnames(values) <- tabnames

ent_hous <- as.data.frame.table(values, stringsAsFactors = FALSE) %>%
  as_tibble() %>%
  set_names(c(response$fields$label,"value")) %>%
  mutate(indicator = "Housing Entitlement") %>%
  rename(category = `Housing Entitlement - Tenure`) 


query <- list(database = unbox("str:database:UC_Households"),
              measures = "str:count:UC_Households:V_F_UC_HOUSEHOLDS",
              dimensions = c("str:field:UC_Households:V_F_UC_HOUSEHOLDS:COA_CODE",
                             "str:field:UC_Households:F_UC_HH_DATE:DATE_NAME",
                             "str:field:UC_Households:V_F_UC_HOUSEHOLDS:HCDISABLED_CHILD_ENTITLEMENT") %>% matrix(),
              recodes = list(
                `str:field:UC_Households:V_F_UC_HOUSEHOLDS:COA_CODE` = list(
                  map = list("str:value:UC_Households:V_F_UC_HOUSEHOLDS:COA_CODE:V_C_MASTERGEOG11_LA_TO_REGION:E08000009")),
                `str:field:UC_Households:F_UC_HH_DATE:DATE_NAME` = list(
                  map = as.list(paste0("str:value:UC_Households:F_UC_HH_DATE:DATE_NAME:C_UC_HH_DATE:",c(202001,202002,202003,202004,202005,202006,202007,202008,202009,202010,202011,202012,202101,202102,202103,202104,202105,202106,202107,202108,202109,202110,202111,202112,202201,202202,202203,202204,202205,202206,202207,202208,202210,202211,202212,202301,202302,202303,202304,202305,202306,202307,202308,202309,202310,202311,202312,202401,202402,202403,202404,202405)))),
                `str:field:UC_Households:V_F_UC_HOUSEHOLDS:HCDISABLED_CHILD_ENTITLEMENT` = list(
                  map = as.list(paste0("str:value:UC_Households:V_F_UC_HOUSEHOLDS:HCDISABLED_CHILD_ENTITLEMENT:C_UC_DISABLED_CHILD_ENTITLEMENT:",c(0,1))))

              )) %>% toJSON()
request <- POST(
  url = path,
  body = query,
  config = add_headers(APIKey = api_key),
  encode = "json")
response <- fromJSON(content(request, as = "text"), flatten = TRUE)
# extract list items and convert to a dataframe
tabnames <- response$fields$items %>% purrr::map(~.$labels %>% unlist)
values <- response$cubes[[1]]$values
dimnames(values) <- tabnames

ent_dis_child <- as.data.frame.table(values, stringsAsFactors = FALSE) %>%
  as_tibble() %>%
  set_names(c(response$fields$label,"value")) %>%
  mutate(indicator = "Disabled Child Entitlement")%>%
  rename(category = `Disabled Child Entitlement`)



query <- list(database = unbox("str:database:UC_Households"),
              measures = "str:count:UC_Households:V_F_UC_HOUSEHOLDS",
              dimensions = c("str:field:UC_Households:V_F_UC_HOUSEHOLDS:COA_CODE",
                             "str:field:UC_Households:F_UC_HH_DATE:DATE_NAME",
                             "str:field:UC_Households:V_F_UC_HOUSEHOLDS:HCCHILD_ENTITLEMENT") %>% matrix(),
              recodes = list(
                `str:field:UC_Households:V_F_UC_HOUSEHOLDS:COA_CODE` = list(
                  map = list("str:value:UC_Households:V_F_UC_HOUSEHOLDS:COA_CODE:V_C_MASTERGEOG11_LA_TO_REGION:E08000009")),
                `str:field:UC_Households:F_UC_HH_DATE:DATE_NAME` = list(
                  map = as.list(paste0("str:value:UC_Households:F_UC_HH_DATE:DATE_NAME:C_UC_HH_DATE:",c(202001,202002,202003,202004,202005,202006,202007,202008,202009,202010,202011,202012,202101,202102,202103,202104,202105,202106,202107,202108,202109,202110,202111,202112,202201,202202,202203,202204,202205,202206,202207,202208,202210,202211,202212,202301,202302,202303,202304,202305,202306,202307,202308,202309,202310,202311,202312,202401,202402,202403,202404,202405)))),
                `str:field:UC_Households:V_F_UC_HOUSEHOLDS:HCCHILD_ENTITLEMENT` = list(
                  map = as.list(paste0("str:value:UC_Households:V_F_UC_HOUSEHOLDS:HCCHILD_ENTITLEMENT:C_UC_CHILD_ENTITLEMENT:",c(0,1))))
                
              )) %>% toJSON()
request <- POST(
  url = path,
  body = query,
  config = add_headers(APIKey = api_key),
  encode = "json")
response <- fromJSON(content(request, as = "text"), flatten = TRUE)
# extract list items and convert to a dataframe
tabnames <- response$fields$items %>% purrr::map(~.$labels %>% unlist)
values <- response$cubes[[1]]$values
dimnames(values) <- tabnames

ent_child <- as.data.frame.table(values, stringsAsFactors = FALSE) %>%
  as_tibble() %>%
  set_names(c(response$fields$label,"value")) %>%
  mutate(indicator = "Child Entitlement") %>%
  rename(category = `Child Entitlement`)




query <- list(database = unbox("str:database:UC_Households"),
              measures = "str:count:UC_Households:V_F_UC_HOUSEHOLDS",
              dimensions = c("str:field:UC_Households:V_F_UC_HOUSEHOLDS:COA_CODE",
                             "str:field:UC_Households:F_UC_HH_DATE:DATE_NAME",
                             "str:field:UC_Households:V_F_UC_HOUSEHOLDS:HCCARER_ENTITLEMENT") %>% matrix(),
              recodes = list(
                `str:field:UC_Households:V_F_UC_HOUSEHOLDS:COA_CODE` = list(
                  map = list("str:value:UC_Households:V_F_UC_HOUSEHOLDS:COA_CODE:V_C_MASTERGEOG11_LA_TO_REGION:E08000009")),
                `str:field:UC_Households:F_UC_HH_DATE:DATE_NAME` = list(
                  map = as.list(paste0("str:value:UC_Households:F_UC_HH_DATE:DATE_NAME:C_UC_HH_DATE:",c(202001,202002,202003,202004,202005,202006,202007,202008,202009,202010,202011,202012,202101,202102,202103,202104,202105,202106,202107,202108,202109,202110,202111,202112,202201,202202,202203,202204,202205,202206,202207,202208,202210,202211,202212,202301,202302,202303,202304,202305,202306,202307,202308,202309,202310,202311,202312,202401,202402,202403,202404,202405)))),
                `str:field:UC_Households:V_F_UC_HOUSEHOLDS:HCCARER_ENTITLEMENT` = list(
                  map = as.list(paste0("str:value:UC_Households:V_F_UC_HOUSEHOLDS:HCCARER_ENTITLEMENT:C_UC_CARER_ENTITLEMENT:",c(0,1))))
                
              )) %>% toJSON()
request <- POST(
  url = path,
  body = query,
  config = add_headers(APIKey = api_key),
  encode = "json")
response <- fromJSON(content(request, as = "text"), flatten = TRUE)
# extract list items and convert to a dataframe
tabnames <- response$fields$items %>% purrr::map(~.$labels %>% unlist)
values <- response$cubes[[1]]$values
dimnames(values) <- tabnames

ent_carer <- as.data.frame.table(values, stringsAsFactors = FALSE) %>%
  as_tibble() %>%
  set_names(c(response$fields$label,"value")) %>%
  mutate(indicator = "Carer Entitlement") %>%
  rename(category = `Carer Entitlement`)

uc_entitlement_trend <- bind_rows(ent_wa,ent_hous,ent_dis_child, ent_child, ent_carer) %>%
#uc_entitlement_trend <- bind_rows(ent_hous, ent_child, ent_carer) %>%
  mutate(category = ifelse(category %in% c("LCW", "LCWRA"), "LCW or LCWRA",category)) %>%
  select(area_name = `National - Regional - LA - OAs`, date = Month, everything()) %>%
  group_by(area_name, date, category, indicator) %>%
  summarise(value = sum (value)) %>%
  ungroup() %>%
  group_by(area_name, date, indicator) %>%
  mutate(percentage = round(value/sum(value)*100,1)) %>%
  filter(!category %in% c("None", "No"))%>%
  ungroup() %>%
  mutate(date = as.Date(paste('01', date), format='%d %B %Y')) %>%
  pivot_longer(value:percentage, names_to = "measure", values_to = "value") %>%
  mutate(measure = ifelse(measure == "value", "Count", "Percentage"))

entitlement_stats <- uc_entitlement_trend %>%
  filter(date %in% c("2024-02-01", "2023-02-01", "2022-02-01")) %>%
  group_by(indicator, measure) %>%
  mutate(prev_year = value - lag(value),
         prev_baseYear = value - value[1]) %>%
  filter(measure == "Percentage")

```

Households on Universal Credit can access additional entitlements for children, health and disabilities, housing, and carers, at the top of the standard allowance. Each household can receive more then one entitlement. As a consequence of the Covid-19 pandemic the number of entitlements showed a sudden increase whilst at the same time the proportion of household receiving entitlements decreased indicating a larger number of claims not receiving additional entitlements. 

As claimants are migrating from legacy benefits to Universal Credit and new claims are reported, the number of household receiving an entitlement have been increasing.


<div class="row">
<div class = "col-sm-12">
   
```{r UCEntitlementLine, fig.height= 5, fig.width=12}

#From Feb 2023 to Feb 2024, both the proportion of household receiving Limited Capability for Work Entitlement and Child Entitlement have increased 1.8% points. The previous year, from Feb 2022 to Feb 2023, the Limited Capability for Work Entitlement proportion increased 3.7% points whilst the Child Entitlement proportion increased 1.5% points.

uc_plot_entitlement_c <- uc_entitlement_trend %>%
  mutate(indicator = ordered(indicator, levels = pull(arrange(filter(., date == "2024-02-01", measure == "Count"), desc(value)),indicator))) 

  

 ggplot(uc_plot_entitlement_c, aes(x = date, y = value, colour = indicator)) +
  geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +
  geom_line(size = 1) +
   geom_text(uc_plot_entitlement_c %>% filter(date == max(date)), mapping = aes(x = max(date) + 120, label = value, fontface = "bold", hjust = 1), direction = "y", show.legend = FALSE) +
  facet_wrap(~measure, scales = "free") +
   scale_color_brewer(palette = "Set2") +
  scale_x_date(breaks = c(min(uc_plot_entitlement_c$date), as.Date(as.yearmon("Jan 2021"),format="%b %Y"), as.Date(as.yearmon("Jan 2022"),format="%b %Y"),  as.Date(as.yearmon("Jan 2023"),format="%b %Y"), as.Date(as.yearmon("Jan 2024"),format="%b %Y")), date_labels = "%b %y") +
  scale_y_continuous(expand = c(0.005, 0.005)) +
  labs(title = "Households on Universal Credit by entitlements",
       subtitle = paste("Trafford, ", base, " - ",format(max(uc_plot_entitlement_c$date),format="%b %Y")),
       caption = "Source: DWP",
       x = NULL,
       y = NULL,
       colour = NULL) +
  theme_minimal(base_size = 10) +
  theme(plot.title = element_text(size=14,face = "bold"),
        plot.subtitle = element_text(size=10),
    plot.margin = unit(rep(0.5,4), "cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks.x = element_line(colour = "#333333", size = 0.5),
    panel.spacing = unit(2, "lines"),
    legend.position = "top",
    strip.text = element_text(size=12,face = "bold", hjust = 0),
    strip.position = "left") +
  coord_cartesian(clip = "off")

```

</div>
</div>

</section>

</main>



```{css}

@import url('https://fonts.googleapis.com/css?family=Roboto');

html, body
{
    padding: 0;
    margin: 0;
    height: 100%;
    font-family: 'Open Sans', sans-serif;
    color: #212121;
}

h1, h2, h3, h4, h5, h6, header
{
    font-family: 'Roboto', sans-serif;
    color: #707070;
}

a
{
    color: #046dc3;
    text-decoration: none;
}

a:hover
{
    text-decoration: underline;
}

h1, h2, h3, h4, h5, h6, .header
{
    font-family: 'Roboto', sans-serif;
    color: #757575;
}

p {
  line-height: 1.8em;
}

.main-container {
  max-width: 1000px;
}
.header {
  border-bottom: 2px solid #555;
  font-weight: 400;
  text-transform: uppercase;
}

.header:hover {
  background-color: #eee;
}

.bar-cell {
  display: flex;
  align-items: center;
}

.number {
  width: 35px;
  text-align: right;
}

.bar-chart {
  display: flex;
  flex-grow: 1;
  height: 100%;
}

.bar-neg-container {
  display: flex;
  justify-content: flex-end;
}

.bar-pos-container {
  display: flex;
  justify-content: flex-start;
}

.bar {
  height: 12px;
}

.large-table {
  font-size: 13px;
}

.table {
  font-size: 12px;
}

.small-table,  .xsmall-table {
  font-size: 11px;
}

.small-table .rt-td {
  padding: 3px 3px;
}
.xsmall-table .rt-td .rt-td-inner{  
  padding: 0px 0px;
  font-size: 11px;
}

.large-table .rt-td {
  white-space: nowrap;
  overflow: hidden;
  text-overflow: ellipsis;
}

.tableTitle {
  font-weight: bold;
  margin: 0px;
}

```
