---
title: "Change on unemployment during COVID-19"
output:
    pdf_document:
        latex_engine: xelatex
geometry: "left=3cm,right=3cm,top=2cm,bottom=2cm"
header-includes:
  - \usepackage[sfdefault]{arimo}
  - \usepackage[T1]{fontenc}
---

```{r setup, include=FALSE, echo=FALSE, results="hide",message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse) ; library(httr) ; library(readxl) ;   library(lubridate)
library(zoo) ;library(sf) ; library(jsonlite) ; library(zoo) ; library(scales)

```

# Claimant Count by age

The [Claimant Count](https://www.nomisweb.co.uk/sources/cc) indicates the number of people claiming benefits principally for the reason of being unemployed. From March 2020 to May 2020 the largest percentage of change  occurred on the band age 25 to 29 years with 146% change and the lowest percentage of change occurred on the band age 60 to 64. When considering the number of additional claims the band age 18 to 24 years with 785 new claims.




```{r fig.show="hold", out.width="50%"}
#Claimant Count
# Source: ONS
# URL: https://www.nomisweb.co.uk/sources/cc
# Licence: Open Government Licence

Claimant <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_162_1.data.csv?geography=1811939363&date=latestMINUS2-latest&gender=0&age=0,2,10,11,3,12...16,4,17...20&measure=1&measures=20100") 

claimant_age <- Claimant %>%
  select(period = DATE_NAME,
         area_code = GEOGRAPHY_CODE,
         area_name = GEOGRAPHY_NAME,
         age = AGE_NAME,
         age_type = AGE_TYPE,
         count = OBS_VALUE) %>%
  mutate(period = format(as.yearmon(period, "%b %Y"),"%b %y"),
         Age =  gsub("Aged |All categories: Age ", "", age))

age_bands <- claimant_age %>%
  filter(age_type=="Ageband") %>%
  select(period, Age, count) %>%
  spread(period,count) %>%
  mutate(`Additional claims`= `May 20`-`Mar 20`,
         `% change` = round(`Additional claims`/`Mar 20` *100,0)) %>%
  mutate(Age=factor(Age, levels=c("16-24", "25-49", "50+", "16+"))) %>%
  arrange(Age)

quinary <- claimant_age %>%
  filter(age_type=="Quinary") %>%
  select(period, Age, count) %>%
  spread(period,count) %>%
  mutate(`Additional claims`= `May 20`-`Mar 20`,
         `% change` = round(`Additional claims`/`Mar 20` *100,0)) %>%
  mutate_if(is.numeric, list(~na_if(., Inf)))


knitr::kable(quinary,caption = "Change on the Claimant Count from March 2020 to May 2020 by age bands")
knitr::kable(age_bands, caption = "Change on the Claimant Count from March 2020 to May 2020 by broad age groups")

ggplot(drop_na(quinary), aes(x = Age, y = `% change`)) + 
  geom_col(alpha = 0.8, fill= "#66c2a4" ) +
  scale_y_continuous( expand = c(0, 0)) +
  labs(title = "Percentage of change on Claimant Count by Age",
       subtitle = "March to May 2020, Trafford",
       caption = "Source: ONS",
       x = NULL,
       y = "% change",
       fill = NULL) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "none")

ggplot(quinary, aes(x = Age, y = `Additional claims`)) + 
  geom_col(alpha = 0.8, fill= "#66c2a4" ) +
  scale_y_continuous( expand = c(0, 0)) +
  labs(title = "Additional claims on Claimant Count by Age",
       subtitle = "March to May 2020, Trafford",
       caption = "Source: ONS",
       x = NULL,
       y = "Additional claims",
       fill = NULL) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "none")

```

# Claimant Count by Ward

The ward with the largest percentage of change from March 2020 to May 2020 is Timperley with 236% of change and the ward with the lowest percentage of change is Bucklow-St Martins with 64% of change. However, the ward with more additional claims is Clifford with 420, followed by Longford (340), Gorse Hill(320), Stretford(275) and Bucklow-St Martins(235).

```{r claimantWard, echo=FALSE, fig.show="hold", out.width="50%", fig.width=4}

claimant_ward <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_162_1.data.csv?geography=1660945005...1660945019,1660945021,1660945020,1660945022...1660945025&date=latestMINUS2-latest&gender=0&age=0&measure=1,2&measures=20100")

claimant_area <- claimant_ward %>%
  select(period = DATE_NAME,
         area_code = GEOGRAPHY_CODE,
         area_name = GEOGRAPHY_NAME,
         measure = MEASURE_NAME,
         count = OBS_VALUE) %>%
  mutate(period = as.yearmon(period, "%b %Y")) %>%
  spread(period,count) %>%
  mutate(`Additional claims`= `May 2020`-`Mar 2020`,
         `% change` = ifelse(measure=="Claimant count",round(`Additional claims`/`Mar 2020` *100,0),`May 2020`-`Mar 2020`))

count <- claimant_area %>%
  filter(measure=="Claimant count")

knitr::kable(count %>% select(Ward=area_name,`Mar 2020`:`% change`), caption = "Change on the Claimant Count from March 2020 to May 2020 by broad age wards")

wards <- st_read("https://www.traffordDataLab.io/spatial_data/ward/2017/trafford_ward_generalised.geojson", quiet = TRUE) %>% 
  left_join(., select(count,`% change`,area_code)  , by = "area_code") 

wards2 <- st_read("https://www.traffordDataLab.io/spatial_data/ward/2017/trafford_ward_generalised.geojson", quiet = TRUE) %>% 
  left_join(., select(count,`Additional claims`,area_code)  , by = "area_code") 

ggplot() +
  geom_sf(data = wards, aes(fill = `% change`), alpha = 1, colour = "#FFFFFF",  size = 0.5) +
  scale_fill_gradientn(colours = c("#edf8fb","#b2e2e2","#66c2a4","#2ca25f","#006d2c"), 
                       na.value = "#f0f0f0",
                       guide = guide_legend(keyheight = unit(3, units = "mm"), 
                                            keywidth=unit(12, units = "mm"), 
                                            label.position = "bottom", 
                                            title.position = 'top', 
                                            nrow = 1)) +
  labs(x = NULL, y = NULL, title = "Percentage of change on Claimant Count",
       subtitle = "March to May 2020, Trafford",
       caption = "Source: ONS.\nContains OS data © Crown copyright and database right 2020") +
  theme(plot.margin = unit(c(0.5,0,1,0), "cm"),
        plot.title = element_text(size=10,face = "bold"),
        plot.subtitle = element_text(size=9),
        axis.text = element_blank(),
        plot.caption = element_text(size=6,margin = margin(t = 55)),
        legend.position = c(0.3,-0.05),
        legend.title = element_text(size = 9),
        legend.text = element_text(size = 8),
        plot.background = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        axis.ticks = element_blank())

ggplot() +
  geom_sf(data = wards2, aes(fill = `Additional claims`), alpha = 1, colour = "#FFFFFF",  size = 0.5) +
  scale_fill_gradientn(colours = c("#edf8fb","#b2e2e2","#66c2a4","#2ca25f","#006d2c"), 
                       na.value = "#f0f0f0",
                       guide = guide_legend(keyheight = unit(3, units = "mm"), 
                                            keywidth=unit(12, units = "mm"), 
                                            label.position = "bottom", 
                                            title.position = 'top', 
                                            nrow = 1)) +
  labs(x = NULL, y = NULL, title = "Additional claims on Claimant Count",
       subtitle = "March to May 2020, Trafford",
       caption = "Source: ONS.\nContains OS data © Crown copyright and database right 2020") +
  theme(plot.margin = unit(c(0.5,0,1,0), "cm"),
        plot.title = element_text(size=10,face = "bold"),
        plot.subtitle = element_text(size=9),
        axis.text = element_blank(),
        plot.caption = element_text(size=6,margin = margin(t = 55)),
        legend.position = c(0.35,-0.05),
        legend.title = element_text(size = 9),
        legend.text = element_text(size = 8),
        plot.background = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        axis.ticks = element_blank())

```

&nbsp;


# Universal Credit change by age
&nbsp;



The largest percentage of change on claims of [Universal Credit](https://stat-xplore.dwp.gov.uk/webapi/metadata/UC_Monthly/UC_Monthly.html) during March 2020 to May 2020 occured on the age band of over 65 years with 333% of change with the claims raising from 9 on March to 39 on May. The age band with more additional claims is the 25 to 29 years with an increase of 982 claims.

&nbsp;

&nbsp;



```{r uc, echo=FALSE, fig.show="hold", out.width="50%"}

#Universal Credit

# Source: DWP
# URL: https://stat-xplore.dwp.gov.uk/webapi/metadata/UC_Monthly/UC_Monthly.html
# Licence: Open Government Licence

#House of Commons Library MSOA Names
# URL: https://visual.parliament.uk/msoanames

lookup <- read_csv("https://visual.parliament.uk/msoanames/static/MSOA-Names-v1.1.0.csv") %>%
  filter(Laname=="Trafford")

# add your API key
api_key <- ""

# API endpoint
path <- "https://stat-xplore.dwp.gov.uk/webapi/rest/v1/table"

query <- list(database = unbox("str:database:UC_Monthly"),
              measures = "str:count:UC_Monthly:V_F_UC_CASELOAD_FULL",
              dimensions = c("str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE",
                             "str:field:UC_Monthly:F_UC_DATE:DATE_NAME",
                             "str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:AGE_CODE") %>% matrix(),
              recodes = list(
                `str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE` = list(
                  map = list("str:value:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE:V_C_MASTERGEOG11_LA_TO_REGION:E08000009")),
                `str:field:UC_Monthly:F_UC_DATE:DATE_NAME` = list(
                  map = as.list(paste0("str:value:UC_Monthly:F_UC_DATE:DATE_NAME:C_UC_DATE:",c(202003,202004,202005)))),
                `str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:AGE_CODE` = list(
                  map = as.list(paste0("str:value:UC_Monthly:V_F_UC_CASELOAD_FULL:AGE_CODE:C_UC_AGE_BAND:",c(1,2,3,4,5,6,7,8,9,10,999))))
                
              )) %>% toJSON()

request <- POST(
  url = path,
  body = query,
  config = add_headers(APIKey = api_key),
  encode = "json")

response <- fromJSON(content(request, as = "text"), flatten = TRUE)

# extract list items and convert to a dataframe
tabnames <- response$fields$items %>% map(~.$labels %>% unlist)
values <- response$cubes[[1]]$values
dimnames(values) <- tabnames

traf_uc <- as.data.frame.table(values, stringsAsFactors = FALSE) %>%
  as_tibble() %>% 
  set_names(c(response$fields$label,"value")) %>%
  mutate(indicator="People on Universal Credit", 
         period=format(as.yearmon(`Month`, "%B %Y"),"%b %y")) %>%
  select(area_name = `National - Regional - LA - OAs`, indicator, Age="Age (bands and single year)", period, value) %>%
  spread(period,value) %>%
  mutate(`Additional claims`= `May 20`-`Mar 20`,
         `% change` = round(`Additional claims`/`Mar 20` *100,0))


knitr::kable(select(traf_uc,Age,`Mar 20`,`Apr 20`,`May 20`,`Additional claims`,`% change`),caption = "Change on Universal Credit claims from March 2020 to May 2020 by age bands")

ggplot(traf_uc, aes(x = Age, y = `% change`)) + 
  geom_col(alpha = 0.8, fill= "#66c2a4" ) +
  scale_y_continuous( expand = c(0, 0)) +
  labs(title = "Percentage of change on Universal Credit claims by Age",
       subtitle = "March to May 2020, Trafford",
       caption = "Source: DWP",
       x = NULL,
       y = "% change",
       fill = NULL) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "none")

ggplot(traf_uc, aes(x = Age, y = `Additional claims`)) + 
  geom_col(alpha = 0.8, fill= "#66c2a4" ) +
  scale_y_continuous( expand = c(0, 0)) +
  labs(title = "Additional claims on Universal Credit by Age",
       subtitle = "March to May 2020, Trafford",
       caption = "Source: DWP",
       x = NULL,
       y = "Additional claims",
       fill = NULL) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "none")

```

&nbsp;

&nbsp;

# Universal Credit change by MSOA

Davyhulme MSOA presents the largest percentage of change (163%) followed by Timperley South(141%) from March 2020 to May 2020. There were 601 additional claims on Old Trafford MSOA  which represent the largest increase among the Trafford MSOAs. 

```{r uc2, echo=FALSE, fig.show="hold", out.width="50%", fig.width=4}

#Universal Credit

# Source: DWP
# URL: https://stat-xplore.dwp.gov.uk/webapi/metadata/UC_Monthly/UC_Monthly.html
# Licence: Open Government Licence

query <- list(database = unbox("str:database:UC_Monthly"),
              measures = "str:count:UC_Monthly:V_F_UC_CASELOAD_FULL",
              dimensions = c("str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE",
                             "str:field:UC_Monthly:F_UC_DATE:DATE_NAME") %>% matrix(),
              recodes = list(
                `str:field:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE` = list(
                  map = as.list(paste0("str:value:UC_Monthly:V_F_UC_CASELOAD_FULL:COA_CODE:V_C_MASTERGEOG11_MSOA_TO_LA:E0", seq(2001259, 2001286, 1)))),
                `str:field:UC_Monthly:F_UC_DATE:DATE_NAME` = list(
                  map = as.list(paste0("str:value:UC_Monthly:F_UC_DATE:DATE_NAME:C_UC_DATE:",c(202003,202004,202005))))
              )) %>% toJSON()

request <- POST(
  url = path,
  body = query,
  config = add_headers(APIKey = api_key),
  encode = "json")

response <- fromJSON(content(request, as = "text"), flatten = TRUE)

# extract list items and convert to a dataframe
tabnames <- response$fields$items %>% map(~.$labels %>% unlist)
values <- response$cubes[[1]]$values
dimnames(values) <- tabnames

uc <- as.data.frame.table(values, stringsAsFactors = FALSE) %>%
  as_tibble() %>% 
  set_names(c(response$fields$label,"value")) %>%
  left_join(lookup%>%select(msoa11cd,msoa11nm,msoa11hclnm), by = c("National - Regional - LA - OAs" = "msoa11nm")) %>%
  mutate(indicator="People on Universal Credit", 
         period=format(as.yearmon(`Month`, "%B %Y"),"%b %y")) %>%
  select(area_code=msoa11cd, area_name = msoa11hclnm, indicator, period, value) %>%
  spread(period,value) %>%
  mutate(`Additional claims`= `May 20`-`Mar 20`,
         `% change` = round(`Additional claims`/`Mar 20` *100,0))

knitr::kable(select(uc, MSOA=area_name, `Mar 20`,`Apr 20`,`May 20`,`Additional claims`,`% change`),caption = "Change on Universal Credit claims from March 2020 to May 2020 by age MSOA")

msoa_uc <- st_read("https://www.traffordDataLab.io/spatial_data/msoa/2011/trafford_msoa_generalised.geojson", quiet = TRUE) %>% 
  left_join(., select(uc,`Additional claims`,`% change`,area_code)  , by = "area_code")

ggplot() +
  geom_sf(data = msoa_uc, aes(fill = `% change`), alpha = 1, colour = "#FFFFFF",  size = 0.5) +
  scale_fill_gradientn(colours = c("#edf8fb","#b2e2e2","#66c2a4","#2ca25f","#006d2c"), 
                       na.value = "#f0f0f0",
                       guide = guide_legend(keyheight = unit(3, units = "mm"), 
                                            keywidth=unit(12, units = "mm"), 
                                            label.position = "bottom", 
                                            title.position = 'top', 
                                            nrow = 1)) +
  labs(x = NULL, y = NULL, title = "Percentage of change on Universal Credit",
       subtitle = "March to May 2020, Trafford",
       caption = "Source: DWP.\nContains OS data © Crown copyright and database right 2020") +
  theme(plot.margin = unit(c(0.5,0,1,0), "cm"),
        plot.title = element_text(size=10,face = "bold"),
        plot.subtitle = element_text(size=9),
        axis.text = element_blank(),
        plot.caption = element_text(size=6,margin = margin(t = 55)),
        legend.position = c(0.4,-0.05),
        legend.title = element_text(size = 9),
        legend.text = element_text(size = 8),
        plot.background = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        axis.ticks = element_blank())

ggplot() +
  geom_sf(data = msoa_uc, aes(fill = `Additional claims`), alpha = 1, colour = "#FFFFFF",  size = 0.5) +
  scale_fill_gradientn(colours = c("#edf8fb","#b2e2e2","#66c2a4","#2ca25f","#006d2c"), 
                       na.value = "#f0f0f0",
                       guide = guide_legend(keyheight = unit(3, units = "mm"), 
                                            keywidth=unit(12, units = "mm"), 
                                            label.position = "bottom", 
                                            title.position = 'top', 
                                            nrow = 1)) +
  labs(x = NULL, y = NULL, title = "Additional claims on Universal Credit",
       subtitle = "March to May 2020, Trafford",
       caption = "Source: DWP.\nContains OS data © Crown copyright and database right 2020") +
  theme(plot.margin = unit(c(0.5,0,1,0), "cm"),
        plot.title = element_text(size=10,face = "bold"),
        plot.subtitle = element_text(size=9),
        axis.text = element_blank(),
        plot.caption = element_text(size=6,margin = margin(t = 55)),
        legend.position = c(0.5,-0.05),
        legend.title = element_text(size = 9),
        legend.text = element_text(size = 8),
        plot.background = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        axis.ticks = element_blank())

```
