---
title: "Relative deprivation in Trafford"
output:
  html_document:
    theme: flatly
    self_contained: TRUE
    toc: true
    toc_depth: 2
    toc_float: true
    df_print: paged
    includes:
      in_header: rmd_report_includes_head.html
      before_body: rmd_report_includes_body.html
---

Last updated: `r format(Sys.time(), '%d %B %Y')`

<style>

p.caption {
 font-size: 0.75em;
 text-align: right; 
 color: #7f7f7f;
}

p.captionLeft {
 font-size: 0.75em;
 color: #7f7f7f;
}

p.medFont {
font-size: 16px;
}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse) ; library(sf) ; library(knitr) ; library(kableExtra) ; library(scales) ; library(shadowtext) ; library(rio) ; library(jsonlite) ; library(ggbeeswarm)

source("https://github.com/traffordDataLab/assets/raw/master/theme/ggplot2/theme_lab.R")

gm <- c("Bolton","Bury","Manchester","Oldham","Rochdale","Salford","Stockport","Tameside","Trafford","Wigan")

cipfa <- c("Swindon","Warrington","Milton Keynes","Solihull","South Gloucestershire","Stockport","Telford and Wrekin","Bedford","Cheshire West and Chester","North Northamptonshire","Bury","Derby","Peterborough","Stockton-on-Tees","Halton", "Trafford")

cssn <- c("Hertfordshire","Stockport","Buckinghamshire","Oxfordshire","Surrey","Bromley","Cheshire East","Cambridgeshire","Gloucestershire","Solihull")

nhspg <- c("Bracknell Forest","Windsor and Maidenhead","Bury","Stockport","Solihull","Barnet","Bexley","Bromley","Croydon","Havering","Hillingdon","Kingston upon Thames","Merton","Richmond upon Thames","Sutton", "Trafford")



Trafford_IoD_LSOA <- read_csv("https://assets.publishing.service.gov.uk/media/691ded56d140bbbaa59a2a7d/File_7_IoD2025_All_Ranks_Scores_Deciles_Population_Denominators.csv") %>%
  filter(`Local Authority District name (2024)` == "Trafford") %>% 
  select(lsoa21cd = 1, lad24cd = 3, lad24nm = 4, 5:52) %>% 
  gather(variable, value, -lsoa21cd, -lad24cd, -lad24nm) %>% 
  mutate(measure = case_when(str_detect(variable, "Score") ~ "score", 
                             str_detect(variable, "Decile") ~ "decile", 
                             str_detect(variable, "Rank") ~ "rank"),
         index_domain = case_when(str_detect(variable, "Multiple") ~ "Index of Multiple Deprivation", 
                                  str_detect(variable, "Employment") ~ "Employment",
                                  str_detect(variable, "Education") ~ "Education, Skills and Training",
                                  str_detect(variable, "Health") ~ "Health and Disability",
                                  str_detect(variable, "Crime") ~ "Crime",
                                  str_detect(variable, "Barriers to Housing") ~ "Barriers to Housing and Services",
                                  str_detect(variable, "Living") ~ "Living Environment",
                                  str_detect(variable, "IDACI") ~ "Income Deprivation Affecting Children",
                                  str_detect(variable, "IDAOPI") ~ "Income Deprivation Affecting Older People",
                                  str_detect(variable, "Children") ~ "Children and Young People",
                                  str_detect(variable, "Adult Skills") ~ "Adult Skills",
                                  str_detect(variable, "Geographical Barriers") ~ "Geographical Barriers",
                                  str_detect(variable, "Wider Barriers") ~ "Wider Barriers",
                                  str_detect(variable, "Indoors") ~ "Indoors",
                                  str_detect(variable, "Outdoors") ~ "Outdoors",
                                  TRUE ~ "Income")) %>% 
  select(lsoa21cd, lad24cd, lad24nm, measure, value, index_domain) %>% 
  spread(measure, value) %>% 
  mutate(year = "2025")

IoD_local_authority_2015 <- read_csv("https://www.trafforddatalab.io/open_data/IndicesOfDeprivation/IoD2015/local_authority_district/IoD2015.csv")

IoD_local_authority_2019 <- read_csv("https://www.trafforddatalab.io/open_data/IndicesOfDeprivation/IoD2019/local_authority_district/IoD2019.csv")

IoD_local_authority_2025 <- read_csv("https://www.trafforddatalab.io/open_data/IndicesOfDeprivation/IoD2025/local_authority_district/indices_of_deprivation_2025_lad.csv")

IoD_local_authority <- IoD_local_authority_2025 %>%
  mutate(year = 2025) %>%
  select(area_code, area_name, index_domain, rank_average_score, proportion_bottom_decile, year) %>%
  bind_rows(IoD_local_authority_2019 %>%
  select(area_code = lad19cd, area_name = lad19nm, index_domain, rank_average_score, proportion_bottom_decile, year)) %>%
  bind_rows(IoD_local_authority_2015 %>%
  select(area_code = lad19cd, area_name = lad19nm, index_domain, rank_average_score, proportion_bottom_decile, year)) %>%
  filter(area_name %in% c(gm,cipfa,cssn,nhspg)) %>%
  mutate(index_domain= if_else (index_domain %in% c("Index of Multiple Deprivation", "Income Deprivation Affecting Children", "Income Deprivation Affecting Older People"),index_domain, str_replace(index_domain," Deprivation", ""))) 
  


#trafford_lsoa <- st_read("data/trafford_lsoa.geojson", quiet = TRUE)
trafford_lsoa <- st_read("https://www.trafforddatalab.io/council_open_data/boundaries/lsoa.geojson", quiet = TRUE)

#trafford_wards <- st_read("data/trafford_wards.geojson", quiet = TRUE)
trafford_wards <- st_read("https://www.trafforddatalab.io/council_open_data/boundaries/electoral_ward.geojson", quiet = TRUE)

lookup_wards_lsoa <- fromJSON("https://services1.arcgis.com/ESMARspQHYMw9BZ9/arcgis/rest/services/LSOA21_WD25_LAD25_EW_LU_v2/FeatureServer/0/query?where=LAD25NM%20%3D%20'TRAFFORD'&outFields=*&outSR=4326&f=json", flatten = TRUE) %>% 
  pluck("features") %>% 
  as_tibble() %>% 
  select(lsoa21cd = attributes.LSOA21CD, ward = attributes.WD25NM)

# dumbbell plot
dumbbell_plot <- function(data){
  data <- data %>%
  arrange(desc(`2025`)) %>%
  mutate(`Local authority` = factor(`Local authority`, levels = `Local authority`))
  
  ggplot() +
  geom_segment(
    data = gather(data, year, rank, -`Local authority`) %>% 
      group_by(`Local authority`) %>% 
      summarise(start = range(rank)[1], end = range(rank)[2]) %>% 
      ungroup(),
    aes(x = start, xend = end, y = `Local authority`, yend = `Local authority`), 
    color = "#e3e2e1", size = 2) +
  geom_point(
    data = gather(data, year, rank, -`Local authority`),
    aes(x = rank, y = `Local authority`, color = year), 
    size = 4) +
  scale_x_continuous(expand = c(0.005, 0.005), 
                     limits = c(0, 317),
                     breaks = c(1, 100, 200, 300)) +
  scale_color_manual(values = c("#a6cee3", "#1f78b4"), 
                     labels = c("2019", "2025")) +
    #scale_y_discrete(limits = rev(levels(data$`Local authority`))) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, colour = NULL) +
  theme_lab() +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y = element_text(hjust = 0),
        legend.position = c(0.5,1),
        legend.text = element_text(size = 10)) +
  guides(color = guide_legend(nrow = 1))
}

# change plot
change_plot <- function(data,nrows = 2){
  
  y_scale_range <- c(0,max(data$proportion_bottom_decile, na.rm = TRUE)+6)
  ggplot(data = data, aes(x = year)) +
  geom_ribbon(aes(ymin = 0, ymax = proportion_bottom_decile), fill = "#1f78b4") +#"#f768a1") +
  geom_point(aes(y = proportion_bottom_decile), fill = "#212121", colour = "#FFFFFF", pch = 21, size = 3) +
  geom_text(aes(label = sprintf('%0.1f%%', proportion_bottom_decile), y = proportion_bottom_decile), vjust = -1, size = 2.6) +
  scale_x_continuous(breaks = unique(data$year), expand = c(0.26, 0)) +
  scale_y_continuous(expand = c(0.005, 0.005), limits = y_scale_range) +
  facet_wrap(facets = ~`Local authority`, nrow = nrows, strip.position = "bottom", scales = "free_x", labeller = label_wrap_gen(width = 20)) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, fill = NULL) +
  theme_lab() +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  theme(panel.grid.major.y = element_blank(),
        panel.spacing = unit(1, "lines"),
        strip.text = element_text(face = "plain", hjust = 0.5, vjust = 1),
        strip.placement = "outside",
        axis.text.y = element_blank(),
        axis.text.x = element_text(size = 9),
        legend.position = "none")
}

# chorolpleth map
choropleth_map <- function(data){
  ggplot(data) +
  geom_sf(aes(fill = factor(decile)), colour = "#FFFFFF", size = 0.25) +
  geom_sf(data = trafford_wards, fill = NA, alpha = 1, colour = "#000000",  size = 0.8) +
  scale_fill_manual(breaks = 1:10,
                    values = c("#453B52", "#454F69", "#3F657E", "#317B8D", "#239296", "#26A898", "#43BD93", "#6AD189", "#98E37D", "#CAF270"),
                    labels = c("Most\ndeprived", 2:9, "Least\ndeprived")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, 
       caption = "Contains Ordnance Survey data © Crown copyright and database right 2019", 
       fill = "") +
  coord_sf(crs = st_crs(4326), datum = NA) +
  theme_lab() +
  theme(legend.position = "bottom",
        plot.caption = element_text(hjust = 0)) +
  guides(fill = guide_legend(label.position = "bottom", 
                             label.hjust = 0,
                             direction = "horizontal",
                             nrow = 1,
                             keyheight = unit(2, units = "mm"), 
                             keywidth = unit(5, units = "mm")))
}

#swarm plot
swarn_plot <- function(data,title){
ggplot(data,aes(x=rank,y='')) +
  #geom_beeswarm() 
  geom_quasirandom(aes(color = factor(decile)), size = 3, width = 0.3) +
    scale_color_manual(breaks = 1:10,
                    values = c("#453B52", "#454F69", "#3F657E", "#317B8D", "#239296", "#26A898", "#43BD93", "#6AD189", "#98E37D", "#CAF270"),
                    labels = c("Most\ndeprived", 2:9, "Least\ndeprived")) +

    labs(x = NULL, y = NULL, title = title, subtitle = "LSOAs in Trafford", caption = "The lowest the rank, the most deprived\nSource: English Indices of Deprivation, MHCLG | @traffordDataLab", colour = NULL) +
    scale_x_continuous(limits = c(1,32844 )) +
  theme_lab()  +
    theme(plot.title = element_text(size = 16),
          plot.subtitle = element_text(size = 14),
      axis.text.x = element_text(size = 8),
          legend.text = element_text(size = 8),
          panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()) +
  guides(color = guide_legend(nrow = 1))
}

#decile plot
gap_plot <- function(data,title){
  ggplot(data, aes(y = percent, x = factor(decile), fill = factor(decile))) + 
  geom_col() + 
  geom_text(aes(label = paste0(percent, "%"), y = percent + 1),
            face = "bold", size = 6.5) +
  scale_fill_manual(
                    values = c("#453B52", "#454F69", "#3F657E", "#317B8D", "#239296", "#26A898", "#43BD93", "#6AD189", "#98E37D", "#CAF270"),
                    guide="none") +
    labs(x = NULL, y = NULL, title = title, subtitle = "% of Trafford's LSOAs by decile",caption = "IoD 2025 | @traffordDataLab",
       fill = "") +
    theme_lab() +
    theme(plot.title = element_text(size=28, margin = margin(t=100)),
        
        plot.subtitle = element_text(size=24),
        plot.caption = element_text(size=20),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5, size=18),
        #axis.title.y = element_text(hjust = 0.5, size=16),
        panel.grid.major.y = element_blank(),
       #panel.grid.major.y = element_line(size = 0.5, color = "#cbcbcb"), 
       panel.grid.minor.x  = element_blank(),
       axis.text.y = element_blank(), #size = 18, hjust = 0),
       legend.text = element_text(size = 18))
}

```

This report explores the [2025 English Indices of Deprivation](https://www.gov.uk/government/statistics/english-indices-of-deprivation-2025) for Trafford. It compares Trafford with other local authorities in Greater Manchester (GM) and similar authorities in England using the rank of average score, tracks relative change in ranking between 2019 and 2025, indicates the percentage of Lower-layer Super Output Areas (LSOAs) in the most deprived decile, identifies the most deprived small areas and present the disparities within Trafford.

## Key findings

- Trafford is ranked 201 on the 2025 Index of Multiple Deprivation (IMD) out of 296 local authority districts in England where 1 is the most deprived and 296 is the least deprived. Trafford ranked 191 out of 317 local authorities in 2019.
- Trafford decreased in relative deprivation between 2019 and 2025 on the IMD.
- The lowest rank of Trafford was on the Living environment deprivation domain ranking 79 among local authorities in 2025, the largest change from 2019 when Trafford ranked 131. These may be due to the change in datasets and methodology to produce the Living enviromnent domain.
- The proportion of Trafford's Lower-layer Super Output areas in the most deprived 10% nationally was 4.3% on the 2025 IMD. This represents a decrease of 0.8% since the 2019 IMD.
- Trafford was ranked the least deprived of all Greater Manchester local authorities in the 2025 IMD and across all domains except living Environment where ranked 4th least deprived and Barriers to Housing & Services and the Income Deprivation Affecting Older People sub-domain where Trafford ranked 2nd least deprived.
- Trafford was ranked the second least deprived when compared to the 15 CIPFA similar authorities in the 2025 IMD and 1st to 3rd least deprived in the other domains except for the Living Environment where is the 2nd most deprived, Crime where is 7th least deprived and Income Deprivation Affecting Older people subdomain where is 7th least deprived.
- Trafford ranked in the middle compared to the 15 NHS peer group authorities in the 2025 IMD and ranged from 3rd less deprived to 5th most deprived on the other domains and subdomains.
- Trafford 017E in Bucklow-St Martins ward is the most deprived LSOA in Trafford in the IMD ranking 382 out of 33,755 LSOAs in England. It sits within the 5% most deprived LSOAs nationally for Income, Employment, Health & Disability, Crime and both subdomains  Income Deprivation Affecting Children and Income Deprivation Affecting Older People.
- Trafford 019E in Manor ward and Trafford 017E in Bucklow-St Martins ward rank 699 and 712 respectively out of 33,755 LSOAs nationally on the Income Deprivation Affecting Children sub-domain with with more than 80% of children aged 0 to 15 living in income deprived families.
- Trafford 001G in Gorse Hill & Cornbrook ward rank 699 and 712 out of 33,755 LSOAs nationally on the Income Deprivation Affecting Older people sub-domain with with more than 96% of all those aged 60 or over experiencing income deprivation.
- Trafford 017E in in Bucklow-St Martins ward rank 196 out of 33,755 LSOAs nationally on the Employment domain with 41.5% of working-age residents involuntarily excluded from the labour market.

## Background

The [2025 English Indices of Deprivation](https://www.gov.uk/government/statistics/english-indices-of-deprivation-2025) were published by the [Ministry of Housing, Communities & Local Government](https://www.gov.uk/government/organisations/ministry-of-housing-communities-and-local-government) in October 2025. The Indices measure relative levels of deprivation in 33,755 Lower-layer Super Output Areas (LSOAs) in England. Each LSOA contains approximately 1,500 residents or 650 households. They are ranked from 1 (most deprived) to 33,755 (least deprived) and assigned to a decile ranging from 1 (most deprived 10%) to 10 (least deprived 10%). Local authorities in England are also ranked from 1 to 317 based on a population weighted average rank or score of the LSOA data.

Relative deprivation is measured using 55 indicators which are grouped across seven different domains:

* **Income** - the proportion of the population experiencing deprivation relating to low income, including those people-out-of-work, and those who have low earnings    
* **Employment** - the proportion of the working age population in an area involuntarily excluded from the labour market including unemployment, sickness or disabiliry, or caring responsabilities.     
* **Education, Skills and Training** - the lack of attainment and skills in the local population   
* **Health and Disability** - the risk of premature death and the impairment of quality of life through poor physical or mental health     
* **Crime** - the risk of personal and material victimisation at a local level  
* **Barriers to Housing and Services** - the physical and financial accessibility of housing and local services     
* **Living Environment** - the quality of the local environment    

Each domain is combined and weighted to create the **Index of Multiple Deprivation** (IMD), an overall measure of relative deprivation. 

A further 2 supplementary indices of income deprivation are also provided:    

* **Income Deprivation Affecting Children Index** - the proportion of all children aged 0 to 15 living in income deprived families          
* **Income Deprivation Affecting Older People Index** - the proportion of all those aged 60 or over who experience income deprivation

In this report, Trafford is compared with the other 9 Greater Manchester local authorities, the 15 [CIPFA Nearest Neighbours](https://www.cipfa.org/services/cipfastats/nearest-neighbour-model) and the 15 [NHS England Near Neighbour Peer Group](https://github.com/NHSDigital/ASC_LA_Peer_Groups) local authorities for Trafford.

#### Please note
- New datasets, methodologies and geographies have been used to produce the IoD25, meaning outputs are less directly comparable to previous Indices.
- The number of component indicators used to produce the indices has changed from 39 in 2019 to 54 in 2025.
- The Indices of Deprivation measure *relative* deprivation so a change in rank between 2019 and 2015 releases does not necessarily indicate a change in *absolute* levels of deprivation.
- Full information on the indices's methodology and dataset can be found in the [Technical Report](https://assets.publishing.service.gov.uk/media/68ff59c80f801e57b5bef907/ID_2025_Technical_Report.pdf)

#### Further resources
- A web application which interactively visualises the proportion of LSOAs in the most deprived decile and the spatial distribution of relative deprivation for every local authority district in England is available [here](https://trafforddatalab.shinyapps.io/local_deprivation/).
- A collection of outputs related to the English index of deprivation can be found [here](https://www.trafforddatalab.io/deprivation.html)

***

## Index of Multiple Deprivation
Trafford is ranked 201 out of 296 local authority districts in England on the average 2025 Index of Multiple Deprivation (IMD) score, where a rank of 1 is the most deprived. Trafford's relative position has raised from 191 out of 317 local authorities in 2019. In GM, 6 out of 10 authorities including Trafford reported a decrease in relative deprivation rankings since 2019. Trafford remains the least deprived local authority in Greater Manchester.

Half of CIPFA neighbours raised their relative position in ranking among other local authorities becoming relatively less deprived in 2025 compared to 2019. Around 30% of the NHS peer authorities become relatively less deprived in 2025 compared to 2019.

### {.tabset}

Click the tabs below for:

#### GM

<br />
<h4 style="text-align: left;">Relative change in rank of average IMD score, GM authorities</h4>
<div class = "row">
<div class = "col-md-8">
```{r warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Index of Multiple Deprivation", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(gm)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score)

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Index of Multiple Deprivation", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(gm)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

#### CIPFA

<br />
<h4 style="text-align: left;">Relative change in rank of average IMD score, CIPFA statistical neighbours</h4>
<div class = "row">
<div class = "col-md-8">
```{r fig.height=8, warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Index of Multiple Deprivation", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(cipfa)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score) %>%
  filter(!is.na(`2019`))

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Index of Multiple Deprivation", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(cipfa)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

#### NHS Peers

<br />
<h4 style="text-align: left;">Relative change in rank of average IMD score, NHS Near Neighbour peer group</h4>
<div class = "row">
<div class = "col-md-8">
```{r fig.height=8, warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Index of Multiple Deprivation", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(nhspg)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score) %>%
  filter(!is.na(`2019`))

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Index of Multiple Deprivation", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(nhspg)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

### {-}


### {.tabset}

The proportion of Trafford's LSOAs ranked in the 1st decile or 10% most deprived in England was 4.3% (6 out of 139) in 2025. This represents a relative decrease compared to 2019 when 5.1% (7 out of 138) of LSOAs were ranked within the 1st decile. However, the proportion of LSOAs in Trafford in the bottom decile is much lower than other GM local authorities, 3rd lowest among CIPFA neighbours and 4th highest among NHS peers.
<br/>

Click the tabs below for:

#### GM

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of IMD, GM authorities</h4>
```{r fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Index of Multiple Deprivation") %>% 
    filter(area_name %in% c(gm)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 2)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

#### CIPFA

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of IMD, CIPFA statistical neighbours</h4>
```{r fig.height=8, fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Index of Multiple Deprivation") %>% 
    filter(area_name %in% c(cipfa)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 4)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

#### NHS Peers

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of IMD, NHS Near Neighbour peer group</h4>
```{r fig.height=8, fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Index of Multiple Deprivation") %>% 
    filter(area_name %in% c(nhspg)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 4)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

### {-}

2 LSOAs in Bucklow-St Martins, 1 in Old Trafford, 2 in Manor and 1 in Longford ward were amongst the 10% most deprived LSOAs in England in 2025.

<br />
<h4 style="text-align: left;">Index of Multiple Deprivation 2025, LSOAs in Trafford</h4>
<div class = "row">
<div class = "col-md-6">
```{r fig.height=8}
imd_lsoas <- left_join(trafford_lsoa, 
          select(filter(Trafford_IoD_LSOA, year == 2025, index_domain == "Index of Multiple Deprivation"), lsoa21cd, index_domain, decile), by = c("area_code" = "lsoa21cd")) 

choropleth_map(data = imd_lsoas)
```
</div>
<div class = "col-md-6">
```{r}
left_join(trafford_lsoa, 
          select(filter(Trafford_IoD_LSOA, year == 2025, index_domain == "Index of Multiple Deprivation"), lsoa21cd, decile, rank), by = c("area_code" = "lsoa21cd")) %>%
  filter(decile == 1) %>%  
  arrange(rank) %>% 
  left_join(lookup_wards_lsoa, by = c("area_code" = "lsoa21cd")) %>%
  select(LSOA = area_code, Ward = ward, Decile = decile, Rank = rank) %>% 
  mutate(Rank = comma(Rank)) %>% 
  st_set_geometry(NULL) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">Source: English Indices of Deprivation (2025), MHCLG. | @traffordDataLab</p>
<br />

<div class = "row">
<div class = "col-md-8">
```{r fig.height=4}
lsoa_gap <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Index of Multiple Deprivation")

swarn_plot(lsoa_gap, "Index of Multiple Deprivation 2025")
```

</div>

<div class = "col-md-4">
```{r fig.height=8}
lsoa_3dec <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Index of Multiple Deprivation") %>%
  group_by(decile) %>%
  summarise(count = n()) %>%
  mutate(group = ifelse(decile %in% c(1,2,3),"3most", ifelse(decile %in% c(8,9,10),"3least","other"))) %>%
  group_by(group) %>%
  summarise(sum = sum(count)) %>%
  mutate(percent = round(sum/sum(sum)*100,1))

lsoa_perDeci <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Index of Multiple Deprivation") %>%
    group_by(decile) %>%
  summarise(count = n()) %>%
  mutate(percent = round(count/sum(count)*100,0))
gap_plot(lsoa_perDeci, "Index of Multiple Deprivation")
```

</div>
</div>




<p class="medFont">The difference between the rank of the most deprived LSOA (`r format(arrange(lsoa_gap, rank)$rank[1],scientific = FALSE, big.mark=",")`) and the least deprived LSOA (`r format(arrange(lsoa_gap, desc(rank))$rank[1], scientific = FALSE, big.mark=",")`) on the Index of Multiple Deprivation is `r format((arrange(lsoa_gap, desc(rank))$rank[1] - arrange(lsoa_gap, rank)$rank[1]), scientific = FALSE,big.mark=",")` </p> 

<p class="medFont">`r filter(lsoa_3dec, group == "3most")$percent[1]` % of the LSOAs in Trafford are among the 3 most deprived deciles in England, `r filter(lsoa_3dec, group == "3least")$percent[1]` % of the LSOAs in Trafford are  among the 3 least deprived deciles in England </p> 


***

## Income
The Income domain measures the proportion of the population experiencing deprivation relating to low income, including those people-out-of-work, and those who have low earnings.      

Trafford is ranked 189 out of 296 local authorities in England on the average 2025 Income domain score. Trafford’s relative position was 166 out of 317 local authorities in 2019.

### {.tabset}

Click the tabs below for:

#### GM

<br />
<h4 style="text-align: left;">Relative change in rank of average Income score, GM authorities</h4>
<div class = "row">
<div class = "col-md-8">
```{r warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Income", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(gm)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score)

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Income", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(gm)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

#### CIPFA

<br />
<h4 style="text-align: left;">Relative change in rank of average Income score, CIPFA statistical neighbours</h4>
<div class = "row">
<div class = "col-md-8">
```{r fig.height=8, warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Income", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(cipfa)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score) %>%
  filter(!is.na(`2019`))

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Income", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(cipfa)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

#### NHS Peers

<br />
<h4 style="text-align: left;">Relative change in rank of average Income score, NHS Near Neighbour peer group</h4>
<div class = "row">
<div class = "col-md-8">
```{r fig.height=8, warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Income", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(nhspg)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score) %>%
  filter(!is.na(`2019`))

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Income", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(nhspg)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />


### {-}

The proportion of Trafford’s LSOAs ranked in the 1st decile or 10% most deprived in England was 5.8% (8 out of 139) in 2025. This represents a 1.4% increase compared to 2019 when 7.2% (10 out of 138) of LSOAs were ranked within the 1st decile. 

<br />

### {.tabset}

Click the tabs below for:

#### GM

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Income domain, GM authorities</h4>
```{r fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Income")%>% 
    filter(area_name %in% c(gm)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 2)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

#### CIPFA

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Income domain, CIPFA statistical neighbours</h4>
```{r fig.height=8, fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Income") %>% 
    filter(area_name %in% c(cipfa)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 4)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

#### NHS Peers

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Income domain, NHS Near Neighbour peer group</h4>
```{r fig.height=8, fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Income") %>% 
    filter(area_name %in% c(nhspg)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 4)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

### {-}

2 LSOAs in Bucklow-St Martins, 2 in Manor, 3 in Old Trafford and 1 Gorse Hill & Cornbrook ward were amongst the 10% most deprived LSOAs on the Income domain in England in 2019. 4 LSOAs sit within the 5% most deprived LSOAs nationally.

Trafford 017E in Bucklow-St Martins ward 256 out of 33,755 LSOAs nationally with 68.5% of residents experiencing deprivation relating to low income, including those people-out-of-work, and those who have low earnings. In 2019, the same LSOA was ranked 487 out of 32,844 LSOAs nationally with 39.1% of residents experiencing deprivation relating to low income.

<br />
<h4 style="text-align: left;">Income domain 2025, LSOAs in Trafford</h4>
<div class = "row">
<div class = "col-md-6">
```{r fig.height=8}
income_lsoas <- left_join(trafford_lsoa, 
          select(filter(Trafford_IoD_LSOA, year == 2025, index_domain == "Income"), lsoa21cd, index_domain, decile), by = c("area_code" = "lsoa21cd")) 

choropleth_map(data = income_lsoas)
```
</div>
<div class = "col-md-6">
```{r}
left_join(trafford_lsoa, 
          select(filter(Trafford_IoD_LSOA, year == 2025, index_domain == "Income"), lsoa21cd, decile, rank), by = c("area_code" = "lsoa21cd")) %>%
  filter(decile == 1) %>%  
  arrange(rank) %>%  
  left_join(lookup_wards_lsoa, by = c("area_code" = "lsoa21cd")) %>%
  select(LSOA = area_code, Ward = ward, Decile = decile, Rank = rank) %>% 
  mutate(Rank = comma(Rank)) %>% 
  st_set_geometry(NULL) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">Source: English Indices of Deprivation (2025), MHCLG | @traffordDataLab</p>
<br />


<div class = "row">
<div class = "col-md-8">
```{r fig.height=4}
lsoa_gap <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Income")

swarn_plot(lsoa_gap, "Income deprivation 2025")
```

</div>

<div class = "col-md-4">
```{r fig.height=8}
lsoa_3dec <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Income") %>%
  group_by(decile) %>%
  summarise(count = n()) %>%
  mutate(group = ifelse(decile %in% c(1,2,3),"3most", ifelse(decile %in% c(8,9,10),"3least","other"))) %>%
  group_by(group) %>%
  summarise(sum = sum(count)) %>%
  mutate(percent = round(sum/sum(sum)*100,1))

lsoa_perDeci <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Income") %>%
    group_by(decile) %>%
  summarise(count = n()) %>%
  mutate(percent = round(count/sum(count)*100,0))
gap_plot(lsoa_perDeci,"Income Deprivation")
```

</div>
</div>




<p class="medFont">The difference between the rank of the most deprived LSOA (`r format(arrange(lsoa_gap, rank)$rank[1],scientific = FALSE,big.mark=",")`) and the least deprived LSOA (`r format(arrange(lsoa_gap, desc(rank))$rank[1], scientific = FALSE,big.mark=",")`) on the Index of Multiple Deprivation is `r format((arrange(lsoa_gap, desc(rank))$rank[1] - arrange(lsoa_gap, rank)$rank[1]), scientific = FALSE,big.mark=",")` </p> 

<p class="medFont">`r filter(lsoa_3dec, group == "3most")$percent[1]` % of the LSOAs in Trafford are among the 3 most deprived deciles in England, `r filter(lsoa_3dec, group == "3least")$percent[1]` % of the LSOAs in Trafford are  among the 3 least deprived deciles in England </p> 


***

## Employment

The Employment domain measures the proportion of the working age population in an area involuntarily excluded from the labour market.      

Trafford is ranked 162 out of 296 local authorities in England on the average 2025 Employment domain score. Trafford’s relative position was 155 out of 317 local authorities in 2019.

### {.tabset}

Click the tabs below for:

#### GM

<br />
<h4 style="text-align: left;">Relative change in rank of average Employment score, GM authorities</h4>
<div class = "row">
<div class = "col-md-8">
```{r warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Employment", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(gm)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score)

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Employment", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(gm)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

#### CIPFA

<br />
<h4 style="text-align: left;">Relative change in rank of average Employment score, CIPFA statistical neighbours</h4>
<div class = "row">
<div class = "col-md-8">
```{r fig.height=8, warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Employment", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(cipfa)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score) %>%
  filter(!is.na(`2019`))

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Employment", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(cipfa)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

#### NHS Peers

<br />
<h4 style="text-align: left;">Relative change in rank of average Employment score, NHS Near Neighbour peer group</h4>
<div class = "row">
<div class = "col-md-8">
```{r fig.height=8, warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Employment", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(nhspg)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score) %>%
  filter(!is.na(`2019`))

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Employment", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(nhspg)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />


### {-}

The proportion of Trafford’s LSOAs ranked in the 1st decile or 10% most deprived in England was 5.8% (8 out of 139) in 2025. This represents no change compared to 2019 when 5.8% (8 out of 138) of LSOAs were ranked within the 1st decile.

<br />

### {.tabset}

Click the tabs below for:

#### GM

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Employment domain, GM authorities</h4>
```{r fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Employment") %>% 
    filter(area_name %in% c(gm)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 2)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

#### CIPFA

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Employment domain, CIPFA statistical neighbours</h4>
```{r fig.height=8, fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Employment") %>% 
    filter(area_name %in% c(cipfa)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 4)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

#### NHS Peers

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Employment domain, NHS Near Neighbour peer group</h4>
```{r fig.height=8, fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Employment") %>% 
    filter(area_name %in% c(nhspg)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 4)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

### {-}

2 LSOA in Bucklow-St Martins, 2 in Old Trafford, 2 in Manor, 1 in Longford and 1 in Timperley Central ward were amongst the 10% most deprived LSOAs on the Employment domain in England in 2025. 5 LSOAs sit within the 5% most deprived LSOAs nationally.

Trafford 017E in Bucklow-St Martins ward ranks 196 out of 33,755 LSOAs nationally with 41.5% of working-age residents involuntarily excluded from the labour market. In 2019, the same LSOA was ranked 572 out of 32,844 LSOAs nationally with 28.9% of residents involuntarily excluded from the labour market.

<br />
<h4 style="text-align: left;">Employment domain 2025, LSOAs in Trafford</h4>
<div class = "row">
<div class = "col-md-6">
```{r fig.height=8}
employment_lsoas <- left_join(trafford_lsoa, 
          select(filter(Trafford_IoD_LSOA, year == 2025, index_domain == "Employment"), lsoa21cd, index_domain, decile), by = c("area_code" = "lsoa21cd")) 

choropleth_map(data = employment_lsoas)
```
</div>
<div class = "col-md-6">
```{r}
left_join(trafford_lsoa, 
          select(filter(Trafford_IoD_LSOA, year == 2025, index_domain == "Employment"), lsoa21cd, decile, rank), by = c("area_code" = "lsoa21cd")) %>%
  filter(decile == 1) %>%  
  arrange(rank) %>% 
  left_join(lookup_wards_lsoa, by = c("area_code" = "lsoa21cd")) %>%
  select(LSOA = area_code, Ward = ward, Decile = decile, Rank = rank) %>% 
  mutate(Rank = comma(Rank)) %>% 
  st_set_geometry(NULL) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">Source: English Indices of Deprivation (2025), MHCLG | @traffordDataLab</p>
<br />


<div class = "row">
<div class = "col-md-8">
```{r fig.height=4}
lsoa_gap <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Employment")

swarn_plot(lsoa_gap, "Employment deprivation 2025")
```

</div>

<div class = "col-md-4">
```{r fig.height=7}
lsoa_3dec <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Employment") %>%
  group_by(decile) %>%
  summarise(count = n()) %>%
  mutate(group = ifelse(decile %in% c(1,2,3),"3most", ifelse(decile %in% c(8,9,10),"3least","other"))) %>%
  group_by(group) %>%
  summarise(sum = sum(count)) %>%
  mutate(percent = round(sum/sum(sum)*100,1))

lsoa_perDeci <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Employment") %>%
    group_by(decile) %>%
  summarise(count = n()) %>%
  mutate(percent = round(count/sum(count)*100,0))
gap_plot(lsoa_perDeci, "Employment Deprivation")
```

</div>
</div>




<p class="medFont">The difference between the rank of the most deprived LSOA (`r format(arrange(lsoa_gap, rank)$rank[1],scientific = FALSE,big.mark=",")`) and the least deprived LSOA (`r format(arrange(lsoa_gap, desc(rank))$rank[1], scientific = FALSE,big.mark=",")`) on the Index of Multiple Deprivation is `r format((arrange(lsoa_gap, desc(rank))$rank[1] - arrange(lsoa_gap, rank)$rank[1]), scientific = FALSE,big.mark=",")` </p> 

<p class="medFont">`r filter(lsoa_3dec, group == "3most")$percent[1]` % of the LSOAs in Trafford are among the 3 most deprived deciles in England, `r filter(lsoa_3dec, group == "3least")$percent[1]` % of the LSOAs in Trafford are  among the 3 least deprived deciles in England </p> 


***

## Education, Skills and Training
The Education, Skills and Training domain measures the lack of attainment and skills in the local population.    

Trafford is ranked 276 out of 296 local authorities in England on the average 2025 Education, Skills and Training domain score. Trafford’s relative position was 275 out of 317 local authorities in 2019.


<br />

### {.tabset}

Click the tabs below for:

#### GM

<br />
<h4 style="text-align: left;">Relative change in rank of average Education, Skills and Training score, GM authorities</h4>
<div class = "row">
<div class = "col-md-8">
```{r warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Education, Skills and Training", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(gm)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score)

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Education, Skills and Training", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(gm)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

#### CIPFA

<br />
<h4 style="text-align: left;">Relative change in rank of average Education, Skills and Training score, CIPFA statistical neighbours</h4>
<div class = "row">
<div class = "col-md-8">
```{r fig.height=8, warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Education, Skills and Training", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(cipfa)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score) %>%
  filter(!is.na(`2019`))

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Education, Skills and Training", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(cipfa)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

#### NHS Peers

<br />
<h4 style="text-align: left;">Relative change in rank of average Education, Skills and Training score, NHS Near Neighbour peer group</h4>
<div class = "row">
<div class = "col-md-8">
```{r fig.height=8, warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Education, Skills and Training", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(nhspg)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score) %>%
  filter(!is.na(`2019`))

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Education, Skills and Training", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(nhspg)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />


### {-}

The proportion of Trafford’s LSOAs ranked in the 1st decile or 10% most deprived in England was 1.4% (2 out of 139) in 2025. This represents practically no change since 2019 when 1.5% (2 out of 138) of LSOAs were ranked within the 1st decile. 

<br />

### {.tabset}

Click the tabs below for:

#### GM

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Education, Skills and Training domain, GM authorities</h4>
```{r fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Education, Skills and Training") %>% 
    filter(area_name %in% c(gm)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 2)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

#### CIPFA

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Education, Skills and Training domain, CIPFA statistical neighbours</h4>
```{r fig.height=8, fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Education, Skills and Training") %>% 
    filter(area_name %in% c(cipfa)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 4)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

#### NHS Peers

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Education, Skills and Training domain, NHS Near Neighbour peer group</h4>
```{r fig.height=8, fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Education, Skills and Training") %>% 
    filter(area_name %in% c(nhspg)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 4)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

### {-}


2 LSOAs in Bucklow-St Martins ward were among the 10% most deprived LSOAs on the Education, Skills and Training domain in England in 2025.


<br />
<h4 style="text-align: left;">Education, Skills and Training domain 2025, LSOAs in Trafford</h4>
<div class = "row">
<div class = "col-md-6">
```{r fig.height=8}
education_lsoas <- left_join(trafford_lsoa, 
          select(filter(Trafford_IoD_LSOA, year == 2025, index_domain == "Education, Skills and Training"), lsoa21cd, index_domain, decile), by = c("area_code" = "lsoa21cd")) 

choropleth_map(data = education_lsoas)
```
</div>
<div class = "col-md-6">
```{r}
left_join(trafford_lsoa, 
          select(filter(Trafford_IoD_LSOA, year == 2025, index_domain == "Education, Skills and Training"), lsoa21cd, decile, rank), by = c("area_code" = "lsoa21cd")) %>%
  filter(decile == 1) %>%  
  arrange(rank) %>% 
  left_join(lookup_wards_lsoa, by = c("area_code" = "lsoa21cd")) %>%
  select(LSOA = area_code, Ward = ward, Decile = decile, Rank = rank) %>% 
  mutate(Rank = comma(Rank)) %>% 
  st_set_geometry(NULL) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">Source: English Indices of Deprivation (2025), MHCLG | @traffordDataLab</p>
<br />


<div class = "row">
<div class = "col-md-8">
```{r fig.height=4}
lsoa_gap <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Education, Skills and Training")

swarn_plot(lsoa_gap, "Education, Skills and Training Deprivation 2025")
```

</div>

<div class = "col-md-4">
```{r fig.height=8}
lsoa_3dec <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Education, Skills and Training") %>%
  group_by(decile) %>%
  summarise(count = n()) %>%
  mutate(group = ifelse(decile %in% c(1,2,3),"3most", ifelse(decile %in% c(8,9,10),"3least","other"))) %>%
  group_by(group) %>%
  summarise(sum = sum(count)) %>%
  mutate(percent = round(sum/sum(sum)*100,1))

lsoa_perDeci <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Education, Skills and Training") %>%
    group_by(decile) %>%
  summarise(count = n()) %>%
  mutate(percent = round(count/sum(count)*100,0))
gap_plot(lsoa_perDeci,"Education, Skills and Training\nDeprivation")
```

</div>
</div>




<p class="medFont">The difference between the rank of the most deprived LSOA (`r format(arrange(lsoa_gap, rank)$rank[1],scientific = FALSE,big.mark=",")`) and the least deprived LSOA (`r format(arrange(lsoa_gap, desc(rank))$rank[1], scientific = FALSE,big.mark=",")`) on the Index of Multiple Deprivation is `r format((arrange(lsoa_gap, desc(rank))$rank[1] - arrange(lsoa_gap, rank)$rank[1]), scientific = FALSE,big.mark=",")` </p> 

<p class="medFont">`r filter(lsoa_3dec, group == "3most")$percent[1]` % of the LSOAs in Trafford are among the 3 most deprived deciles in England, `r filter(lsoa_3dec, group == "3least")$percent[1]` % of the LSOAs in Trafford are  among the 3 least deprived deciles in England </p> 


***

## Health and Disability
The Health and Disability domain measures the risk of premature death and the impairment of quality of life through poor physical or mental health.     

Trafford is ranked 173 out of 296 local authorities in England on the average 2025 Health and Disability domain score. Trafford’s relative position was 147 out of 317 local authorities in 2019.


<br />

### {.tabset}

Click the tabs below for:

#### GM

<br />
<h4 style="text-align: left;">Relative change in rank of average Health and Disability score, GM authorities</h4>
<div class = "row">
<div class = "col-md-8">
```{r warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Health and Disability", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(gm)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score)

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Health and Disability", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(gm)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

#### CIPFA

<br />
<h4 style="text-align: left;">Relative change in rank of average Health and Disability score, CIPFA statistical neighbours</h4>
<div class = "row">
<div class = "col-md-8">
```{r fig.height=8, warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Health and Disability", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(cipfa)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score) %>%
  filter(!is.na(`2019`))

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Health and Disability", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(cipfa)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

#### NHS Peers

<br />
<h4 style="text-align: left;">Relative change in rank of average Health and Disability score, NHS Near Neighbour peer group</h4>
<div class = "row">
<div class = "col-md-8">
```{r fig.height=8, warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Health and Disability", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(nhspg)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score) %>%
  filter(!is.na(`2019`))

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Health and Disability", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(nhspg)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />


### {-}


The proportion of Trafford’s LSOAs ranked in the 1st decile or 10% most deprived in England was 4.3% (6 out of 139) in 2025. This represents a decline in relative deprivation since 2019 when 5.1% (7 out of 138) of LSOAs were ranked within the 1st decile.

<br />

### {.tabset}

Click the tabs below for:

#### GM

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Health and Disability domain, GM authorities</h4>
```{r fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Health and Disability") %>% 
    filter(area_name %in% c(gm)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 2)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

#### CIPFA

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Health and Disability domain, CIPFA statistical neighbours</h4>
```{r fig.height=8, fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Health and Disability") %>% 
    filter(area_name %in% c(cipfa)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 4)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

#### NHS Peers

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Health and Disability domain, NHS Near Neighbour peer group</h4>
```{r fig.height=8, fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Health and Disability") %>% 
    filter(area_name %in% c(nhspg)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 4)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

### {-}


3 LSOAs in Bucklow-St Martins, 2 in St Mary's, 1 in Stretford and 1 in Clifford ward were amongst the 10% most deprived LSOAs on the Health and Disability domain in England in 2019. 2 LSOAs sit within the 5% most deprived LSOAs nationally.

<br />
<h4 style="text-align: left;">Health and Disability domain 2025, LSOAs in Trafford</h4>
<div class = "row">
<div class = "col-md-6">
```{r fig.height=8}
health_lsoas <- left_join(trafford_lsoa, 
          select(filter(Trafford_IoD_LSOA, year == 2025, index_domain == "Health and Disability"), lsoa21cd, index_domain, decile), by = c("area_code" = "lsoa21cd")) 

choropleth_map(data = health_lsoas)
```
</div>
<div class = "col-md-6">
```{r}
left_join(trafford_lsoa, 
          select(filter(Trafford_IoD_LSOA, year == 2025, index_domain == "Health and Disability"), lsoa21cd, decile, rank), by = c("area_code" = "lsoa21cd")) %>%
  filter(decile == 1) %>%  
  arrange(rank) %>% 
  left_join(lookup_wards_lsoa, by = c("area_code" = "lsoa21cd")) %>%
  select(LSOA = area_code, Ward = ward, Decile = decile, Rank = rank) %>% 
  mutate(Rank = comma(Rank)) %>% 
  st_set_geometry(NULL) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">Source: English Indices of Deprivation (2025), MHCLG | @traffordDataLab</p>
<br />


<div class = "row">
<div class = "col-md-8">
```{r fig.height=4}
lsoa_gap <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Health and Disability")

swarn_plot(lsoa_gap, "Health and Disability Deprivation 2025")
```

</div>

<div class = "col-md-4">
```{r fig.height=8}
lsoa_3dec <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Health and Disability") %>%
  group_by(decile) %>%
  summarise(count = n()) %>%
  mutate(group = ifelse(decile %in% c(1,2,3),"3most", ifelse(decile %in% c(8,9,10),"3least","other"))) %>%
  group_by(group) %>%
  summarise(sum = sum(count)) %>%
  mutate(percent = round(sum/sum(sum)*100,1))

lsoa_perDeci <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Health and Disability") %>%
    group_by(decile) %>%
  summarise(count = n()) %>%
  mutate(percent = round(count/sum(count)*100,0))
gap_plot(lsoa_perDeci, "Health and Disability\nDeprivation")
```

</div>
</div>



<p class="medFont">The difference between the rank of the most deprived LSOA (`r format(arrange(lsoa_gap, rank)$rank[1],scientific = FALSE,big.mark=",")`) and the least deprived LSOA (`r format(arrange(lsoa_gap, desc(rank))$rank[1], scientific = FALSE,big.mark=",")`) on the Index of Multiple Deprivation is `r format((arrange(lsoa_gap, desc(rank))$rank[1] - arrange(lsoa_gap, rank)$rank[1]), scientific = FALSE,big.mark=",")` </p> 

<p class="medFont">`r filter(lsoa_3dec, group == "3most")$percent[1]` % of the LSOAs in Trafford are among the 3 most deprived deciles in England, `r filter(lsoa_3dec, group == "3least")$percent[1]` % of the LSOAs in Trafford are  among the 3 least deprived deciles in England </p> 


***

## Crime
The Crime domain measures the risk of personal and material victimisation at a local level.      

Trafford is ranked 140 out of 296 local authorities in England on the average 2025 Crime domain score. Trafford’s relative position was 138 out of 317 local authorities in 2019.


<br />

### {.tabset}

Click the tabs below for:

#### GM

<br />
<h4 style="text-align: left;">Relative change in rank of average Crime score, GM authorities</h4>
<div class = "row">
<div class = "col-md-8">
```{r warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Crime", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(gm)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score)

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Crime", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(gm)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

#### CIPFA

<br />
<h4 style="text-align: left;">Relative change in rank of average Crime score, CIPFA statistical neighbours</h4>
<div class = "row">
<div class = "col-md-8">
```{r fig.height=8, warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Crime", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(cipfa)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score) %>%
  filter(!is.na(`2019`))

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Crime", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(cipfa)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

#### NHS Peers

<br />
<h4 style="text-align: left;">Relative change in rank of average Crime score, NHS Near Neighbour peer group</h4>
<div class = "row">
<div class = "col-md-8">
```{r fig.height=8, warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Crime", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(nhspg)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score) %>%
  filter(!is.na(`2019`))

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Crime", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(nhspg)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />


### {-}

The proportion of Trafford’s LSOAs ranked in the 1st decile or 10% most deprived in England was 5.8% (8 out of 139) in 2025. This represents a relative increase compared to 2019 when 2.2% (3 out of 138) of LSOAs were ranked within the 1st decile.

<br />

### {.tabset}

Click the tabs below for:

#### GM

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Crime domain, GM authorities</h4>
```{r fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Crime") %>% 
    filter(area_name %in% c(gm)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 2)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

#### CIPFA

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Crime domain, CIPFA statistical neighbours</h4>
```{r fig.height=8, fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Crime") %>% 
    filter(area_name %in% c(cipfa)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 4)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

#### NHS Peers

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Crime domain, NHS Near Neighbour peer group</h4>
```{r fig.height=8, fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Crime") %>% 
    filter(area_name %in% c(nhspg)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 4)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

### {-}

1 LSOA in Old Trafford, 2 in Bucklow-St Martins, 1 in Longford, 1 in Timperley Central and 2 in Manor ward were amongst the 10% most deprived LSOAs on the Crime domain in England in 2025. 3 LSOA sits within the 5% most deprived LSOAs nationally.

<br />
<h4 style="text-align: left;">Crime domain 2025, LSOAs in Trafford</h4>
<div class = "row">
<div class = "col-md-6">
```{r fig.height=8}
crime_lsoas <- left_join(trafford_lsoa, 
          select(filter(Trafford_IoD_LSOA, year == 2025, index_domain == "Crime"), lsoa21cd, index_domain, decile), by = c("area_code" = "lsoa21cd")) 

choropleth_map(data = crime_lsoas)
```
</div>
<div class = "col-md-6">
```{r}
left_join(trafford_lsoa, 
          select(filter(Trafford_IoD_LSOA, year == 2025, index_domain == "Crime"), lsoa21cd, decile, rank), by = c("area_code" = "lsoa21cd")) %>%
  filter(decile == 1) %>%  
  arrange(rank) %>% 
  left_join(lookup_wards_lsoa, by = c("area_code" = "lsoa21cd")) %>%
  select(LSOA = area_code, Ward = ward, Decile = decile, Rank = rank) %>% 
  mutate(Rank = comma(Rank)) %>% 
  st_set_geometry(NULL) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">Source: English Indices of Deprivation (2025), MHCLG | @traffordDataLab</p>
<br />


<div class = "row">
<div class = "col-md-8">
```{r fig.height=4}
lsoa_gap <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Crime")

swarn_plot(lsoa_gap, "Crime Deprivation 2025")
```

</div>

<div class = "col-md-4">
```{r fig.height=8}
lsoa_3dec <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Crime") %>%
  group_by(decile) %>%
  summarise(count = n()) %>%
  mutate(group = ifelse(decile %in% c(1,2,3),"3most", ifelse(decile %in% c(8,9,10),"3least","other"))) %>%
  group_by(group) %>%
  summarise(sum = sum(count)) %>%
  mutate(percent = round(sum/sum(sum)*100,1))

lsoa_perDeci <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Crime") %>%
    group_by(decile) %>%
  summarise(count = n()) %>%
  mutate(percent = round(count/sum(count)*100,0))
gap_plot(lsoa_perDeci, "Crime Deprivation")
```

</div>
</div>




<p class="medFont">The difference between the rank of the most deprived LSOA (`r format(arrange(lsoa_gap, rank)$rank[1],scientific = FALSE,big.mark=",")`) and the least deprived LSOA (`r format(arrange(lsoa_gap, desc(rank))$rank[1], scientific = FALSE,big.mark=",")`) on the Index of Multiple Deprivation is `r format((arrange(lsoa_gap, desc(rank))$rank[1] - arrange(lsoa_gap, rank)$rank[1]), scientific = FALSE,big.mark=",")` </p> 

<p class="medFont">`r filter(lsoa_3dec, group == "3most")$percent[1]` % of the LSOAs in Trafford are among the 3 most deprived deciles in England, `r filter(lsoa_3dec, group == "3least")$percent[1]` % of the LSOAs in Trafford are  among the 3 least deprived deciles in England </p> 


***

## Barriers to Housing and Services
The Barriers to Housing and Services domain measures the physical and financial accessibility of housing and local services.     

Trafford is ranked 287 out of 296 local authorities in England on the average 2025 Barriers to Housing and Services domain score. Trafford’s relative position was 289 out of 317 local authorities in 2019.

<br />

### {.tabset}

Click the tabs below for:

#### GM

<br />
<h4 style="text-align: left;">Relative change in rank of average Barriers to Housing and Services score, GM authorities</h4>
<div class = "row">
<div class = "col-md-8">
```{r warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Barriers to Housing and Services", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(gm)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score)

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Barriers to Housing and Services", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(gm)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

#### CIPFA

<br />
<h4 style="text-align: left;">Relative change in rank of average Barriers to Housing and Services score, CIPFA statistical neighbours</h4>
<div class = "row">
<div class = "col-md-8">
```{r fig.height=8, warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Barriers to Housing and Services", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(cipfa)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score) %>%
  filter(!is.na(`2019`))

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Barriers to Housing and Services", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(cipfa)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

#### NHS Peers

<br />
<h4 style="text-align: left;">Relative change in rank of average Barriers to Housing and Services score, NHS Near Neighbour peer group</h4>
<div class = "row">
<div class = "col-md-8">
```{r fig.height=8, warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Barriers to Housing and Services", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(nhspg)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score) %>%
  filter(!is.na(`2019`))

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Barriers to Housing and Services", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(nhspg)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />


### {-}


None of the Trafford’s LSOAs ranked in the 1st decile or 10% most deprived in England in 2025. This represents a decrease compared to 2019 when 0.7% (1 out of 138) of LSOAs were ranked within the 1st decile.

<br />

### {.tabset}

Click the tabs below for:

#### GM

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Barriers to Housing and Services domain, GM authorities</h4>
```{r fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Barriers to Housing and Services") %>% 
    filter(area_name %in% c(gm)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 2)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

#### CIPFA

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Barriers to Housing and Services domain, CIPFA statistical neighbours</h4>
```{r fig.height=8, fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Barriers to Housing and Services") %>% 
    filter(area_name %in% c(cipfa)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 4)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

#### NHS Peers

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Barriers to Housing and Services domain, NHS Near Neighbour peer group</h4>
```{r fig.height=8, fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Barriers to Housing and Services") %>% 
    filter(area_name %in% c(nhspg)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 4)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

### {-}


There are no LSOA in Trafford amongst the 10% most deprived LSOAs on the Barriers to Housing and Services domain in England in 2025.

<br />
<h4 style="text-align: left;">Barriers to Housing and Services domain 2025, LSOAs in Trafford</h4>
<div class = "row">
<div class = "col-md-6">
```{r fig.height=8}
housing_lsoas <- left_join(trafford_lsoa, 
          select(filter(Trafford_IoD_LSOA, year == 2025, index_domain == "Barriers to Housing and Services"), lsoa21cd, index_domain, decile), by = c("area_code" = "lsoa21cd")) 

choropleth_map(data = housing_lsoas)
```
</div>
<div class = "col-md-6">
```{r}
# left_join(trafford_lsoa, 
#           select(filter(Trafford_IoD_LSOA, year == 2025, index_domain == "Barriers to Housing and Services"), lsoa21cd, decile, rank), by = c("area_code" = "lsoa21cd"))%>%
#   filter(decile == 1) %>%  
#   arrange(rank) %>% 
#   left_join(lookup_wards_lsoa, by = c("area_code" = "lsoa21cd")) %>%
#   select(LSOA = area_code, Ward = ward, Decile = decile, Rank = rank) %>% 
#   mutate(Rank = comma(Rank)) %>% 
#   st_set_geometry(NULL) %>% 
#   kable() %>% 
#   kable_styling(font_size = 12) %>%
#   row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">Source: English Indices of Deprivation (2025), MHCLG | @traffordDataLab</p>
<br />


<div class = "row">
<div class = "col-md-8">
```{r fig.height=4}
lsoa_gap <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Barriers to Housing and Services")

swarn_plot(lsoa_gap, "Barriers to Housing and Services Deprivation 2025")
```

</div>

<div class = "col-md-4">
```{r fig.height=8}
lsoa_3dec <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Barriers to Housing and Services") %>%
  group_by(decile) %>%
  summarise(count = n()) %>%
  mutate(group = ifelse(decile %in% c(1,2,3),"3most", ifelse(decile %in% c(8,9,10),"3least","other"))) %>%
  group_by(group) %>%
  summarise(sum = sum(count)) %>%
  mutate(percent = round(sum/sum(sum)*100,1))

lsoa_perDeci <- Trafford_IoD_LSOA%>%
  filter(index_domain == "Barriers to Housing and Services") %>%
    group_by(decile) %>%
  summarise(count = n()) %>%
  mutate(percent = round(count/sum(count)*100,0)) %>%
  add_row(decile = 1, count = 0, percent = 0)
gap_plot(lsoa_perDeci, "Barriers to Housing and Services\nDeprivation")
```

</div>
</div>




<p class="medFont">The difference between the rank of the most deprived LSOA (`r format(arrange(lsoa_gap, rank)$rank[1],scientific = FALSE,big.mark=",")`) and the least deprived LSOA (`r format(arrange(lsoa_gap, desc(rank))$rank[1], scientific = FALSE,big.mark=",")`) on the Index of Multiple Deprivation is `r format((arrange(lsoa_gap, desc(rank))$rank[1] - arrange(lsoa_gap, rank)$rank[1]), scientific = FALSE,big.mark=",")` </p> 

<p class="medFont">`r filter(lsoa_3dec, group == "3most")$percent[1]` % of the LSOAs in Trafford are among the 3 most deprived deciles in England, `r filter(lsoa_3dec, group == "3least")$percent[1]` % of the LSOAs in Trafford are  among the 3 least deprived deciles in England </p> 


***

## Living Environment
The Living Environment domain measures the quality of the local environment. This includes an ‘indoors’ (Housing in poor condition, Housing Energy Performance Score, Housing lacking private outdoor space) and an ‘outdoors’ (Air quality, Road traffic accidents and Noise pollution) subdomain.

The lowest rank of Trafford was on the Living environment deprivation domain ranking 79 among local authorities in 2025, the largest change from 2019 when Trafford ranked 131. These may be due to the change in datasets with the introduction of the Housing Energy Performance Score, Housing Energy Performance Score and Noise pollution, and elimination of Houses without central heating.
<br />

### {.tabset}

Click the tabs below for:

#### GM

<br />
<h4 style="text-align: left;">Relative change in rank of average Living Environment score, GM authorities</h4>
<div class = "row">
<div class = "col-md-8">
```{r warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Living Environment", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(gm)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score)

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Living Environment", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(gm)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

#### CIPFA

<br />
<h4 style="text-align: left;">Relative change in rank of average Living Environment score, CIPFA statistical neighbours</h4>
<div class = "row">
<div class = "col-md-8">
```{r fig.height=8, warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Living Environment", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(cipfa)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score) %>%
  filter(!is.na(`2019`))

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Living Environment", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(cipfa)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

#### NHS Peers

<br />
<h4 style="text-align: left;">Relative change in rank of average Living Environment score, NHS Near Neighbour peer group</h4>
<div class = "row">
<div class = "col-md-8">
```{r fig.height=8, warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Living Environment", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(nhspg)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score) %>%
  filter(!is.na(`2019`))

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Living Environment", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(nhspg)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />


### {-}

The proportion of Trafford’s LSOAs ranked in the 1st decile or 10% most deprived in England was 5.8% (8 out of 139) in 2025. This represents a relative increase compared to 2019 when 2.9% (4 out of 138) of LSOAs were ranked within the 1st decile. 

<br />

### {.tabset}

Click the tabs below for:

#### GM

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Living Environment domain, GM authorities</h4>
```{r fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Living Environment") %>% 
    filter(area_name %in% c(gm)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 2)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

#### CIPFA

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Living Environment domain, CIPFA statistical neighbours</h4>
```{r fig.height=8, fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Living Environment") %>% 
    filter(area_name %in% c(cipfa)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 4)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

#### NHS Peers

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Living Environment domain, NHS Near Neighbour peer group</h4>
```{r fig.height=8, fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Living Environment") %>% 
    filter(area_name %in% c(nhspg)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 4)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

### {-}


1 LSOA in Stretford & Humphrey Park, 2 in Old Trafford, 1 in Sale Central, 1 in Gorse Hill & Cornbrook, 1 in Hale, 1 in Urmston and 1 in Ashton upon Mersey ward  were amongst the 10% most deprived LSOAs on the Living Environment domain in England in 2025.    

<br />
<h4 style="text-align: left;">Living Environment domain 2025, LSOAs in Trafford</h4>
<div class = "row">
<div class = "col-md-6">
```{r fig.height=8}
environment_lsoas <- left_join(trafford_lsoa, 
          select(filter(Trafford_IoD_LSOA, year == 2025, index_domain == "Living Environment"), lsoa21cd, index_domain, decile), by = c("area_code" = "lsoa21cd")) 

choropleth_map(data = environment_lsoas)
```
</div>
<div class = "col-md-6">
```{r}
left_join(trafford_lsoa, 
          select(filter(Trafford_IoD_LSOA, year == 2025, index_domain == "Living Environment"), lsoa21cd, decile, rank), by = c("area_code" = "lsoa21cd")) %>%
  filter(decile == 1) %>%  
  arrange(rank) %>% 
  left_join(lookup_wards_lsoa, by = c("area_code" = "lsoa21cd")) %>%
  select(LSOA = area_code, Ward = ward, Decile = decile, Rank = rank) %>% 
  mutate(Rank = comma(Rank)) %>% 
  st_set_geometry(NULL) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">Source: English Indices of Deprivation (2025), MHCLG | @traffordDataLab</p>
<br />

<div class = "row">
<div class = "col-md-8">
```{r fig.height=4}
lsoa_gap <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Living Environment")

swarn_plot(lsoa_gap, "Living Environment Deprivation 2025")
```

</div>

<div class = "col-md-4">
```{r fig.height=8}
lsoa_3dec <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Living Environment") %>%
  group_by(decile) %>%
  summarise(count = n()) %>%
  mutate(group = ifelse(decile %in% c(1,2,3),"3most", ifelse(decile %in% c(8,9,10),"3least","other"))) %>%
  group_by(group) %>%
  summarise(sum = sum(count)) %>%
  mutate(percent = round(sum/sum(sum)*100,1))

lsoa_perDeci <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Living Environment") %>%
    group_by(decile) %>%
  summarise(count = n()) %>%
  mutate(percent = round(count/sum(count)*100,0))
gap_plot(lsoa_perDeci, "Living Environment Deprivation")
```

</div>
</div>




<p class="medFont">The difference between the rank of the most deprived LSOA (`r format(arrange(lsoa_gap, rank)$rank[1],scientific = FALSE,big.mark=",")`) and the least deprived LSOA (`r format(arrange(lsoa_gap, desc(rank))$rank[1], scientific = FALSE,big.mark=",")`) on the Index of Multiple Deprivation is `r format((arrange(lsoa_gap, desc(rank))$rank[1] - arrange(lsoa_gap, rank)$rank[1]), scientific = FALSE,big.mark=",")` </p> 

<p class="medFont">`r filter(lsoa_3dec, group == "3most")$percent[1]` % of the LSOAs in Trafford are among the 3 most deprived deciles in England, `r filter(lsoa_3dec, group == "3least")$percent[1]` % of the LSOAs in Trafford are  among the 3 least deprived deciles in England </p> 


***

## Income Deprivation Affecting Children
The Income Deprivation Affecting Children sub-domain measures the proportion of all children aged 0 to 15 living in income deprived families.   

Trafford is ranked 251 out of 296 local authorities in England on the average 2025 Income Deprivation Affecting Children Index score. Trafford’s relative position was 214 out of 317 local authorities in 2019.


<br />

### {.tabset}

Click the tabs below for:

#### GM

<br />
<h4 style="text-align: left;">Relative change in rank of average IDACI score, GM authorities</h4>
<div class = "row">
<div class = "col-md-8">
```{r warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Children", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(gm)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score)

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Children", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(gm)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

#### CIPFA

<br />
<h4 style="text-align: left;">Relative change in rank of average IDACI score, CIPFA statistical neighbours</h4>
<div class = "row">
<div class = "col-md-8">
```{r fig.height=8, warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Children", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(cipfa)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score) %>%
  filter(!is.na(`2019`))

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Children", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(cipfa)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

#### NHS Peers

<br />
<h4 style="text-align: left;">Relative change in rank of average IDACI score, NHS Near Neighbour peer group</h4>
<div class = "row">
<div class = "col-md-8">
```{r fig.height=8, warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Children", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(nhspg)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score) %>%
  filter(!is.na(`2019`))

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Children", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(nhspg)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />


### {-}

The proportion of Trafford’s LSOAs ranked in the 1st decile or 10% most deprived in England was 4.3% (6 out of 139) in 2025. This represents a rise in relative deprivation since 2019 when 3.6% (5 out of 138) of LSOAs were ranked within the 1st decile.

<br />

### {.tabset}

Click the tabs below for:

#### GM

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of IDACI supplementary index, GM authorities</h4>
```{r fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Children") %>% 
    filter(area_name %in% c(gm)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 2)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

#### CIPFA

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of IDACI supplementary index, CIPFA statistical neighbours</h4>
```{r fig.height=8, fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Children") %>% 
    filter(area_name %in% c(cipfa)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 4)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

#### NHS Peers

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of IDACI supplementary index, NHS Near Neighbour peer group</h4>
```{r fig.height=8, fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Children") %>% 
    filter(area_name %in% c(nhspg)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 4)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

### {-}


2 LSOAs in Manor, 3 in Bucklow-St Martins and 1 in Old Trafford ward were amongst the 10% most deprived LSOAs on the Income Deprivation Affecting Children sub-domain in England in 2025.

Trafford 019E in Manor ward and Trafford 017E in Bucklow-St Martins ward rank 699 and 712 respectively out of 33,755 LSOAs nationally on the Income Deprivation Affecting Children sub-domain with with more than 80% of children aged 0 to 15 living in income deprived families. In 2019, the same LSOAs was ranked 1,899 and 805 out of 32,844 LSOAs respectivley nationally with 37.8% and 44.1% of children aged 0 to 15 living in income deprived families.

<br />
<h4 style="text-align: left;">IDACI 2025, LSOAs in Trafford</h4>
<div class = "row">
<div class = "col-md-6">
```{r fig.height=8}
idaci_lsoas <- left_join(trafford_lsoa, 
          select(filter(Trafford_IoD_LSOA, year == 2025, index_domain == "Income Deprivation Affecting Children"), lsoa21cd, index_domain, decile), by = c("area_code" = "lsoa21cd")) 

choropleth_map(data = idaci_lsoas)
```
</div>
<div class = "col-md-6">
```{r}
left_join(trafford_lsoa, 
          select(filter(Trafford_IoD_LSOA, year == 2025, index_domain == "Income Deprivation Affecting Children"), lsoa21cd, decile, rank), by = c("area_code" = "lsoa21cd")) %>%
  filter(decile == 1) %>%  
  arrange(rank) %>% 
  left_join(lookup_wards_lsoa, by = c("area_code" = "lsoa21cd")) %>%
  select(LSOA = area_code, Ward = ward, Decile = decile, Rank = rank) %>% 
  mutate(Rank = comma(Rank)) %>% 
  st_set_geometry(NULL) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">Source: English Indices of Deprivation (2025), MHCLG | @traffordDataLab</p>
<br />

<div class = "row">
<div class = "col-md-8">
```{r fig.height=4}
lsoa_gap <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Income Deprivation Affecting Children")

swarn_plot(lsoa_gap, "Income Deprivation Affecting Children 2025")
```

</div>

<div class = "col-md-4">
```{r fig.height=8}
lsoa_3dec <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Income Deprivation Affecting Children") %>%
  group_by(decile) %>%
  summarise(count = n()) %>%
  mutate(group = ifelse(decile %in% c(1,2,3),"3most", ifelse(decile %in% c(8,9,10),"3least","other"))) %>%
  group_by(group) %>%
  summarise(sum = sum(count)) %>%
  mutate(percent = round(sum/sum(sum)*100,1))

lsoa_perDeci <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Income Deprivation Affecting Children") %>%
    group_by(decile) %>%
  summarise(count = n()) %>%
  mutate(percent = round(count/sum(count)*100,0))
gap_plot(lsoa_perDeci, "IDACI")
```

</div>
</div>




<p class="medFont">The difference between the rank of the most deprived LSOA (`r format(arrange(lsoa_gap, rank)$rank[1],scientific = FALSE,big.mark=",")`) and the least deprived LSOA (`r format(arrange(lsoa_gap, desc(rank))$rank[1], scientific = FALSE,big.mark=",")`) on the Index of Multiple Deprivation is `r format((arrange(lsoa_gap, desc(rank))$rank[1] - arrange(lsoa_gap, rank)$rank[1]), scientific = FALSE,big.mark=",")` </p> 

<p class="medFont">`r filter(lsoa_3dec, group == "3most")$percent[1]` % of the LSOAs in Trafford are among the 3 most deprived deciles in England, `r filter(lsoa_3dec, group == "3least")$percent[1]` % of the LSOAs in Trafford are  among the 3 least deprived deciles in England </p> 


***

## Income Deprivation Affecting Older People	
The Income Deprivation Affecting Older People sub-domain measures the proportion of all those aged 60 or over who experience income deprivation.     

Trafford is ranked 138 out of 296 local authorities in England on the average 2025 Income Deprivation Affecting Older People Index score. Trafford’s relative position was 134 out of 317 local authorities in 2019.



<br />

### {.tabset}

Click the tabs below for:

#### GM

<br />
<h4 style="text-align: left;">Relative change in rank of average IDAOPI score, GM authorities</h4>
<div class = "row">
<div class = "col-md-8">
```{r warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Older People", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(gm)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score)

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Older People", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(gm)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

#### CIPFA

<br />
<h4 style="text-align: left;">Relative change in rank of average IDAOPI score, CIPFA statistical neighbours</h4>
<div class = "row">
<div class = "col-md-8">
```{r fig.height=8, warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Older People", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(cipfa)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score) %>%
  filter(!is.na(`2019`))

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Older People", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(cipfa)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

#### NHS Peers

<br />
<h4 style="text-align: left;">Relative change in rank of average IDAOPI score, NHS Near Neighbour peer group</h4>
<div class = "row">
<div class = "col-md-8">
```{r fig.height=8, warning = FALSE, echo = FALSE}
imd_rank <- filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Older People", year %in% c(2019,2025)) %>%
  filter(area_name %in% c(nhspg)) %>%
  select(`Local authority` = area_name, rank_average_score, year) %>%
  spread(year, rank_average_score) %>%
  filter(!is.na(`2019`))

dumbbell_plot(data = imd_rank)
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Older People", year %in% c(2025,2019)) %>% 
    filter(area_name %in% c(nhspg)) %>%
  select(`Local authority` = area_name,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2025`-`2019`) %>% 
  arrange(`2025`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">2019: 317 authorities, 2025: 296 authorities. The lowest the rank, the most deprived</br>Source: English Indices of Deprivation, MHCLG | @traffordDataLab</p>
<br />

### {-}

The proportion of Trafford’s LSOAs ranked in the 1st decile or 10% most deprived in England was 7.2% (10 out of 139) in 2025. This represents no change on 2019 when 7.2% (10 out of 138) of LSOAs were ranked within the 1st decile.

<br />

### {.tabset}

Click the tabs below for:

#### GM

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of IDAOPI supplementary index, GM authorities</h4>
```{r fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Older People") %>% 
    filter(area_name %in% c(gm)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 2)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

#### CIPFA

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of IDAOPI supplementary index, CIPFA statistical neighbours</h4>
```{r fig.height=8, fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Older People") %>% 
    filter(area_name %in% c(cipfa)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 4)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

#### NHS Peers

<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of IDAOPI supplementary index, NHS Near Neighbour peer group</h4>
```{r fig.height=8, fig.align='center'}
imd_decile <- filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Older People") %>% 
    filter(area_name %in% c(nhspg)) %>%
  select(`Local authority`= area_name,
         proportion_bottom_decile,
         year) %>% 
  spread(year, proportion_bottom_decile) %>% 
  mutate(change = round((`2025` - `2019`)*100,1),
         status = case_when(
           `2025` > `2019` ~ "Increase",
           `2025` < `2019` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, proportion_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         proportion_bottom_decile = proportion_bottom_decile*100) %>% 
  select(`Local authority`, year, proportion_bottom_decile, change, status)

change_plot(data = imd_decile, nrows = 4)
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG | @traffordDataLab</h6>
<br />

### {-}


2 LSOAs in Gorse Hill & Cornbrook, 3 in Old Trafford, 1 in Bucklow-St Martins, 2 in Manor, 1 in Sale More and 1 in Broadheath ward were amongst the 10% most deprived LSOAs on the Income Deprivation Affecting Older People sub-domain in England in 2019. 6 LSOAs sit within the 5% most deprived LSOAs nationally.

Trafford 001G in Gorse Hill & Cornbrook ward rank 699 and 712 out of 33,755 LSOAs nationally on the Income Deprivation Affecting Older people sub-domain with with more than 96% of all those aged 60 or over experiencing income deprivation.

<br />
<h4 style="text-align: left;">IDAOPI 2025, LSOAs in Trafford</h4>
<div class = "row">
<div class = "col-md-6">
```{r fig.height=8}
idaopi_lsoas <- left_join(trafford_lsoa, 
          select(filter(Trafford_IoD_LSOA, year == 2025, index_domain == "Income Deprivation Affecting Older People"), lsoa21cd, index_domain, decile), by = c("area_code" = "lsoa21cd")) 

choropleth_map(data = idaopi_lsoas)
```
</div>
<div class = "col-md-6">
```{r}
left_join(trafford_lsoa, 
          select(filter(Trafford_IoD_LSOA, year == 2025, index_domain == "Income Deprivation Affecting Older People"), lsoa21cd, decile, rank), by = c("area_code" = "lsoa21cd")) %>%
  filter(decile == 1) %>%  
  arrange(rank) %>% 
  left_join(lookup_wards_lsoa, by = c("area_code" = "lsoa21cd")) %>%
  select(LSOA = area_code, Ward = ward, Decile = decile, Rank = rank) %>% 
  mutate(Rank = comma(Rank)) %>% 
  st_set_geometry(NULL) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<p class="caption">Source: English Indices of Deprivation (2025), MHCLG | @traffordDataLab</p>
<br />


<div class = "row">
<div class = "col-md-8">
```{r fig.height=4}
lsoa_gap <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Income Deprivation Affecting Older People")

swarn_plot(lsoa_gap, "Income Deprivation Affecting Older People 2025")
```

</div>

<div class = "col-md-4">
```{r fig.height=8}
lsoa_3dec <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Income Deprivation Affecting Older People") %>%
  group_by(decile) %>%
  summarise(count = n()) %>%
  mutate(group = ifelse(decile %in% c(1,2,3),"3most", ifelse(decile %in% c(8,9,10),"3least","other"))) %>%
  group_by(group) %>%
  summarise(sum = sum(count)) %>%
  mutate(percent = round(sum/sum(sum)*100,1))

lsoa_perDeci <- Trafford_IoD_LSOA %>%
  filter(index_domain == "Income Deprivation Affecting Older People") %>%
    group_by(decile) %>%
  summarise(count = n()) %>%
  mutate(percent = round(count/sum(count)*100,0))
gap_plot(lsoa_perDeci, "IDAOPI")
```

</div>
</div>



<p class="medFont">The difference between the rank of the most deprived LSOA (`r format(arrange(lsoa_gap, rank)$rank[1],scientific = FALSE,big.mark=",")`) and the least deprived LSOA (`r format(arrange(lsoa_gap, desc(rank))$rank[1], scientific = FALSE,big.mark=",")`) on the Index of Multiple Deprivation is `r format((arrange(lsoa_gap, desc(rank))$rank[1] - arrange(lsoa_gap, rank)$rank[1]), scientific = FALSE,big.mark=",")` </p> 

<p class="medFont">`r filter(lsoa_3dec, group == "3most")$percent[1]` % of the LSOAs in Trafford are among the 3 most deprived deciles in England, `r filter(lsoa_3dec, group == "3least")$percent[1]` % of the LSOAs in Trafford are  among the 3 least deprived deciles in England </p> 
