---
title: "Indices of Deprivation 2019"
output:
  html_document:
    theme: flatly
    self_contained: TRUE
    toc: true
    toc_depth: 2
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse) ; library(sf) ; library(knitr) ; library(kableExtra) ; library(scales) ; library(shadowtext)
source("https://github.com/traffordDataLab/assets/raw/master/theme/ggplot2/theme_lab.R")

trafford_IoD <- read_csv("data/trafford_IoD.csv")
IoD_local_authority <- read_csv("data/IoD_local_authority.csv")
trafford_lsoa <- st_read("data/trafford_lsoa.geojson", quiet = TRUE)
trafford_wards <- st_read("data/trafford_wards.geojson", quiet = TRUE)
```

This report explores the English Indices of Deprivation 2019 for Trafford. It compares Trafford with other local authorities in Greater Manchester using the  the rank of average score, tracks relative change in ranking between 2015 and 2019, assesses the percentage of LSOAs in the most deprived decile, and identifies the most deprived small areas.

## Background

The [English Indices of Deprivation](https://www.gov.uk/government/statistics/english-indices-of-deprivation-2019) were published by the [Ministry of Housing, Communities & Local Government](https://www.gov.uk/government/organisations/ministry-of-housing-communities-and-local-government) in September 2019. The Indices measure relative levels of deprivation in 32,844 Lower-layer Super Output Areas (LSOAs) in England. Each LSOA contains approximately 1,500 residents or 650 households. They are ranked from 1 (most deprived) to 32,844 (least deprived) and assigned to a decile ranging from 1 (most deprived 10%) to 10 (least deprived 10%). Local authorities in England are also ranked from 1 to 317 based on a population weighted average rank or score of the LSOA data.

Relative deprivation is measured using 39 indicators which are grouped across seven different domains:

* **Income** - the proportion of the population experiencing deprivation relating to low income    
* **Employment** - the proportion of the working age population who would like to work but are unable to do because of unemployment, sickness or disability, or caring responsibilities     
* **Education, Skills and Training** - the lack of attainment and skills in the local population    
* **Health and Disability** - the  risk  of  premature  death  and  the impairment of quality of life through poor physical or mental health     
* **Crime** - the risk of personal and material victimisation   
* **Barriers to Housing and Services** - the physical and financial accessibility of housing and local services     
* **Living Environment** - the  quality  of  the  local  environment. This includes an 'indoors' (quality  of housing) and an 'outdoors' (air quality and road traffic accidents) subdomain    

Each domain is combined and weighted to create the **Index of Multiple Deprivation** (IMD), an overall relative measure of deprivation. 

A further 2 supplementary indices of income deprivation are also provided:    

* **Income Deprivation Affecting Children Index** - the proportion of all children aged 0 to 15 living in income deprived families          
* **Income Deprivation Affecting Older People Index** - the proportion of all those aged 60 or over who experience income deprivation   

**Please note**      
- The Indices of Deprivation measure *relative* deprivation so a change in rank between 2019 and 2015 releases does not necessarily indicate a change in *absolute* levels of deprivation.     
- Local authority summaries for 2015 have been reaggregated to 2019 Local Authority District boundaries to enable direct comparisons to be made. The methodology followed is available in Appendix N of the IoD2019 [technical report](https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/833951/IoD2019_Technical_Report.pdf).

***

## Index of Multiple Deprivation
Trafford is ranked 191 out of 317 local authority districts in England on the average 2019 Index of Multiple Deprivation (IMD) score, where a rank of 1 is the most deprived. Trafford's relative position has fallen from 199 in 2015. This reflects a general trend in relative deprivation rankings across Greater Manchester where all but 1 local authority districts reported a fall in rank since 2015. However, Trafford remains the least deprived local authority in Greater Manchester.

<br />
<h4 style="text-align: center;">Relative change in rank of average IMD score</h4>
<div class = "row">
<div class = "col-md-8">
```{r warning = FALSE, echo = FALSE}
df <- filter(IoD_local_authority, index_domain == "Index of Multiple Deprivation") %>% 
  select(`Local authority` = lad19nm,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score)

ggplot() +
  geom_segment(
    data = gather(df, year, rank, -`Local authority`) %>% 
      group_by(`Local authority`) %>% 
      summarise(start = range(rank)[1], end = range(rank)[2]) %>% 
      ungroup(),
    aes(x = start, xend = end, y = fct_reorder(`Local authority`, end, .desc = TRUE), yend = `Local authority`), 
    color = "#e3e2e1", size = 2) +
  geom_point(
    data = gather(df, year, rank, -`Local authority`),
    aes(x = rank, y = fct_reorder(`Local authority`, rank, .desc = TRUE), color = year), 
    size = 4) +
  scale_x_continuous(expand = c(0.005, 0.005), 
                     limits = c(0, 317),
                     breaks = c(1, 100, 200, 300)) +
  scale_color_manual(values = c("#a6cee3", "#1f78b4"), 
                     labels = c("2015", "2019")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, colour = NULL) +
  theme_lab() +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y = element_text(hjust = 0),
        legend.position = c(0.5,1),
        legend.text = element_text(size = 10)) +
  guides(color = guide_legend(nrow = 1))
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Index of Multiple Deprivation") %>% 
  select(`Local authority` = lad19nm,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2019`-`2015`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG</h6>
<br />

The proportion of Trafford's LSOAs ranked in the 1st decile or 10% most deprived in England was 5.1% (7 out of 138) in 2019. This represents a relative increase on 2015 when 2.9% (4 out of 138) of LSOAs were ranked within the 1st decile. However, the proportion of LSOAs in Trafford in the bottom decile is much lower than other Greater Manchester local authorities.

<br />
<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of IMD</h4>
```{r fig.align='center'}
df <- filter(IoD_local_authority, index_domain == "Index of Multiple Deprivation") %>% 
  select(`Local authority` = lad19nm,
         percent_bottom_decile,
         year) %>% 
  spread(year, percent_bottom_decile) %>% 
  mutate(change = round((`2019` - `2015`)*100,1),
         status = case_when(
           `2019` > `2015` ~ "Increase",
           `2019` < `2015` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, percent_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         percent_bottom_decile = percent_bottom_decile*100) %>% 
  select(`Local authority`, year, percent_bottom_decile, change, status)

ggplot(data = df, aes(x = year)) +
  geom_ribbon(aes(ymin = 0, ymax = percent_bottom_decile, fill =  factor(status))) +
  geom_point(aes(y = percent_bottom_decile), fill = "#212121", colour = "#FFFFFF", pch = 21, size = 3) +
  geom_text(aes(label = sprintf('%0.1f%%', percent_bottom_decile), y = percent_bottom_decile), vjust = -1, size = 3) +
  geom_text(aes(label = sprintf('%+0.1f%%', change)), x = 2017, y = 0, vjust = -1, fontface = 'bold', size = 3) +
  scale_fill_manual(values = c('Increase' = '#ff6961', 'Decrease' = "#77dd77", 'No change' = '#ffe662')) +
  scale_x_continuous(breaks = unique(df$year), expand = c(0.26, 0)) +
  scale_y_continuous(expand = c(0.005, 0.005), limits = c(0, 50)) +
  facet_wrap(facets = ~`Local authority`, nrow = 2) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, fill = NULL) +
  theme_lab() +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  theme(panel.grid.major.y = element_blank(),
        panel.spacing = unit(1, "lines"),
        strip.text = element_text(face = "plain", hjust = 0.5, vjust = 1),
        axis.text.y = element_blank(),
        legend.position = "none")
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG</h6>
<br />

The map below shows that 2 LSOAs in Bucklow-St Martins, 1 in Clifford, 2 in St-Mary’s, 1 in Stretford and 1 in Village ward were amongst the 10% most deprived LSOAs in England in 2019.

<br />
<h4 style="text-align: center;">Index of Multiple Deprivation 2019, LSOAs in Trafford</h4>
<div class = "row">
<div class = "col-md-7">
```{r fig.height=8}
left_join(trafford_lsoa, 
          select(filter(trafford_IoD, year == 2019, index_domain == "Index of Multiple Deprivation"), lsoa11cd, index_domain, decile), by = "lsoa11cd")  %>% 
  ggplot() +
  geom_sf(aes(fill = factor(decile)), colour = "#FFFFFF", size = 0.25) +
  geom_sf(data = trafford_wards, fill = NA, alpha = 1, colour = "#000000",  size = 0.8) +
  # geom_shadowtext(data = trafford_wards, aes(x = lon, y = lat, label = wd18nm), colour = "#FFFFFF", fontface = "bold", size = 3, bg.colour = "#212121", nudge_y = 0.002) +
  scale_fill_manual(breaks = 1:10,
                    values = c("#453B52", "#454F69", "#3F657E", "#317B8D", "#239296", "#26A898", "#43BD93", "#6AD189", "#98E37D", "#CAF270"),
                    labels = c("Most\ndeprived", 2:9, "Least\ndeprived")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, 
       caption = "Contains Ordnance Survey data © Crown copyright and database right 2019", 
       fill = "") +
  coord_sf(crs = st_crs(4326), datum = NA) +
  theme_lab() +
  theme(legend.position = "bottom",
        plot.caption = element_text(hjust = 0)) +
  guides(fill = guide_legend(label.position = "bottom", 
                             label.hjust = 0,
                             direction = "horizontal",
                             nrow = 1,
                             keyheight = unit(2, units = "mm"), 
                             keywidth = unit(5, units = "mm")))
```
</div>
<div class = "col-md-5">
```{r}
left_join(trafford_lsoa, 
          select(filter(trafford_IoD, year == 2019, index_domain == "Index of Multiple Deprivation"), lsoa11cd, decile, rank), by = "lsoa11cd") %>% 
  filter(decile == 1) %>%  
  arrange(rank) %>% 
  select(LSOA = lsoa11nm, Ward = wd18nm, Decile = decile, Rank = rank) %>% 
  mutate(Rank = comma(Rank)) %>% 
  st_set_geometry(NULL) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation (2019), MHCLG.</h6>
<br />

***

## Income
Trafford is ranked 166 out of 317 local authorities in England on the average 2019 Income domain score. This represents a fall (-3) from 169 in 2015.

<br />
<h4 style="text-align: center;">Relative change in rank of average Income score</h4>
<div class = "row">
<div class = "col-md-8">
```{r warning = FALSE, echo = FALSE}
df <- filter(IoD_local_authority, index_domain == "Income") %>% 
  select(`Local authority` = lad19nm,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score)

ggplot() +
  geom_segment(
    data = gather(df, year, rank, -`Local authority`) %>% 
      group_by(`Local authority`) %>% 
      summarise(start = range(rank)[1], end = range(rank)[2]) %>% 
      ungroup(),
    aes(x = start, xend = end, y = fct_reorder(`Local authority`, end, .desc = TRUE), yend = `Local authority`), 
    color = "#e3e2e1", size = 2) +
  geom_point(
    data = gather(df, year, rank, -`Local authority`),
    aes(x = rank, y = fct_reorder(`Local authority`, rank, .desc = TRUE), color = year), 
    size = 4) +
  scale_x_continuous(expand = c(0.005, 0.005), 
                     limits = c(0, 317),
                     breaks = c(1, 100, 200, 300)) +
  scale_color_manual(values = c("#a6cee3", "#1f78b4"), 
                     labels = c("2015", "2019")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, colour = NULL) +
  theme_lab() +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y = element_text(hjust = 0),
        legend.position = c(0.5,1),
        legend.text = element_text(size = 10)) +
  guides(color = guide_legend(nrow = 1))
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Income") %>% 
  select(`Local authority` = lad19nm,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2019`-`2015`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG</h6>
<br />

<br />
<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Income domain</h4>
```{r fig.align='center'}
df <- filter(IoD_local_authority, index_domain == "Income") %>% 
  select(`Local authority` = lad19nm,
         percent_bottom_decile,
         year) %>% 
  spread(year, percent_bottom_decile) %>% 
  mutate(change = round((`2019` - `2015`)*100,1),
         status = case_when(
           `2019` > `2015` ~ "Increase",
           `2019` < `2015` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, percent_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         percent_bottom_decile = percent_bottom_decile*100) %>% 
  select(`Local authority`, year, percent_bottom_decile, change, status)

ggplot(data = df, aes(x = year)) +
  geom_ribbon(aes(ymin = 0, ymax = percent_bottom_decile, fill =  factor(status))) +
  geom_point(aes(y = percent_bottom_decile), fill = "#212121", colour = "#FFFFFF", pch = 21, size = 3) +
  geom_text(aes(label = sprintf('%0.1f%%', percent_bottom_decile), y = percent_bottom_decile), vjust = -1, size = 3) +
  geom_text(aes(label = sprintf('%+0.1f%%', change)), x = 2017, y = 0, vjust = -1, fontface = 'bold', size = 3) +
  scale_fill_manual(values = c('Increase' = '#ff6961', 'Decrease' = "#77dd77", 'No change' = '#ffe662')) +
  scale_x_continuous(breaks = unique(df$year), expand = c(0.26, 0)) +
  scale_y_continuous(expand = c(0.005, 0.005), limits = c(0, 50)) +
  facet_wrap(facets = ~`Local authority`, nrow = 2) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, fill = NULL) +
  theme_lab() +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  theme(panel.grid.major.y = element_blank(),
        panel.spacing = unit(1, "lines"),
        strip.text = element_text(face = "plain", hjust = 0.5, vjust = 1),
        axis.text.y = element_blank(),
        legend.position = "none")
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG</h6>
<br />

<br />
<h4 style="text-align: center;">Income domain 2019, LSOAs in Trafford</h4>
<div class = "row">
<div class = "col-md-7">
```{r fig.height=8}
left_join(trafford_lsoa, 
          select(filter(trafford_IoD, year == 2019, index_domain == "Income"), lsoa11cd, index_domain, decile), by = "lsoa11cd")  %>% 
  ggplot() +
  geom_sf(aes(fill = factor(decile)), colour = "#FFFFFF", size = 0.25) +
  geom_sf(data = trafford_wards, fill = NA, alpha = 1, colour = "#000000",  size = 0.8) +
  # geom_shadowtext(data = trafford_wards, aes(x = lon, y = lat, label = wd18nm), colour = "#FFFFFF", fontface = "bold", size = 3, bg.colour = "#212121", nudge_y = 0.002) +
  scale_fill_manual(breaks = 1:10,
                    values = c("#453B52", "#454F69", "#3F657E", "#317B8D", "#239296", "#26A898", "#43BD93", "#6AD189", "#98E37D", "#CAF270"),
                    labels = c("Most\ndeprived", 2:9, "Least\ndeprived")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, 
       caption = "Contains Ordnance Survey data © Crown copyright and database right 2019", 
       fill = "") +
  coord_sf(crs = st_crs(4326), datum = NA) +
  theme_lab() +
  theme(legend.position = "bottom",
        plot.caption = element_text(hjust = 0)) +
  guides(fill = guide_legend(label.position = "bottom", 
                             label.hjust = 0,
                             direction = "horizontal",
                             nrow = 1,
                             keyheight = unit(2, units = "mm"), 
                             keywidth = unit(5, units = "mm")))
```
</div>
<div class = "col-md-5">
```{r}
left_join(trafford_lsoa, 
          select(filter(trafford_IoD, year == 2019, index_domain == "Income"), lsoa11cd, decile, rank), by = "lsoa11cd") %>% 
  filter(decile == 1) %>%  
  arrange(rank) %>% 
  select(LSOA = lsoa11nm, Ward = wd18nm, Decile = decile, Rank = rank) %>% 
  mutate(Rank = comma(Rank)) %>% 
  st_set_geometry(NULL) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation (2019), MHCLG.</h6>
<br />

***

## Employment
Trafford is ranked 155 out of 317 local authorities in England on the average 2019 Employment domain score. This represents a rise (+5) from 150 in 2015.

<br />
<h4 style="text-align: center;">Relative change in rank of average Employment score</h4>
<div class = "row">
<div class = "col-md-8">
```{r warning = FALSE, echo = FALSE}
df <- filter(IoD_local_authority, index_domain == "Employment") %>% 
  select(`Local authority` = lad19nm,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score)

ggplot() +
  geom_segment(
    data = gather(df, year, rank, -`Local authority`) %>% 
      group_by(`Local authority`) %>% 
      summarise(start = range(rank)[1], end = range(rank)[2]) %>% 
      ungroup(),
    aes(x = start, xend = end, y = fct_reorder(`Local authority`, end, .desc = TRUE), yend = `Local authority`), 
    color = "#e3e2e1", size = 2) +
  geom_point(
    data = gather(df, year, rank, -`Local authority`),
    aes(x = rank, y = fct_reorder(`Local authority`, rank, .desc = TRUE), color = year), 
    size = 4) +
  scale_x_continuous(expand = c(0.005, 0.005), 
                     limits = c(0, 317),
                     breaks = c(1, 100, 200, 300)) +
  scale_color_manual(values = c("#a6cee3", "#1f78b4"), 
                     labels = c("2015", "2019")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, colour = NULL) +
  theme_lab() +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y = element_text(hjust = 0),
        legend.position = c(0.5,1),
        legend.text = element_text(size = 10)) +
  guides(color = guide_legend(nrow = 1))
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Employment") %>% 
  select(`Local authority` = lad19nm,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2019`-`2015`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG</h6>
<br />

<br />
<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Employment domain</h4>
```{r fig.align='center'}
df <- filter(IoD_local_authority, index_domain == "Employment") %>% 
  select(`Local authority` = lad19nm,
         percent_bottom_decile,
         year) %>% 
  spread(year, percent_bottom_decile) %>% 
  mutate(change = round((`2019` - `2015`)*100,1),
         status = case_when(
           `2019` > `2015` ~ "Increase",
           `2019` < `2015` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, percent_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         percent_bottom_decile = percent_bottom_decile*100) %>% 
  select(`Local authority`, year, percent_bottom_decile, change, status)

ggplot(data = df, aes(x = year)) +
  geom_ribbon(aes(ymin = 0, ymax = percent_bottom_decile, fill =  factor(status))) +
  geom_point(aes(y = percent_bottom_decile), fill = "#212121", colour = "#FFFFFF", pch = 21, size = 3) +
  geom_text(aes(label = sprintf('%0.1f%%', percent_bottom_decile), y = percent_bottom_decile), vjust = -1, size = 3) +
  geom_text(aes(label = sprintf('%+0.1f%%', change)), x = 2017, y = 0, vjust = -1, fontface = 'bold', size = 3) +
  scale_fill_manual(values = c('Increase' = '#ff6961', 'Decrease' = "#77dd77", 'No change' = '#ffe662')) +
  scale_x_continuous(breaks = unique(df$year), expand = c(0.26, 0)) +
  scale_y_continuous(expand = c(0.005, 0.005), limits = c(0, 50)) +
  facet_wrap(facets = ~`Local authority`, nrow = 2) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, fill = NULL) +
  theme_lab() +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  theme(panel.grid.major.y = element_blank(),
        panel.spacing = unit(1, "lines"),
        strip.text = element_text(face = "plain", hjust = 0.5, vjust = 1),
        axis.text.y = element_blank(),
        legend.position = "none")
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG</h6>
<br />

<br />
<h4 style="text-align: center;">Employment domain 2019, LSOAs in Trafford</h4>
<div class = "row">
<div class = "col-md-7">
```{r fig.height=8}
left_join(trafford_lsoa, 
          select(filter(trafford_IoD, year == 2019, index_domain == "Employment"), lsoa11cd, index_domain, decile), by = "lsoa11cd")  %>% 
  ggplot() +
  geom_sf(aes(fill = factor(decile)), colour = "#FFFFFF", size = 0.25) +
  geom_sf(data = trafford_wards, fill = NA, alpha = 1, colour = "#000000",  size = 0.8) +
  # geom_shadowtext(data = trafford_wards, aes(x = lon, y = lat, label = wd18nm), colour = "#FFFFFF", fontface = "bold", size = 3, bg.colour = "#212121", nudge_y = 0.002) +
  scale_fill_manual(breaks = 1:10,
                    values = c("#453B52", "#454F69", "#3F657E", "#317B8D", "#239296", "#26A898", "#43BD93", "#6AD189", "#98E37D", "#CAF270"),
                    labels = c("Most\ndeprived", 2:9, "Least\ndeprived")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, 
       caption = "Contains Ordnance Survey data © Crown copyright and database right 2019", 
       fill = "") +
  coord_sf(crs = st_crs(4326), datum = NA) +
  theme_lab() +
  theme(legend.position = "bottom",
        plot.caption = element_text(hjust = 0)) +
  guides(fill = guide_legend(label.position = "bottom", 
                             label.hjust = 0,
                             direction = "horizontal",
                             nrow = 1,
                             keyheight = unit(2, units = "mm"), 
                             keywidth = unit(5, units = "mm")))
```
</div>
<div class = "col-md-5">
```{r}
left_join(trafford_lsoa, 
          select(filter(trafford_IoD, year == 2019, index_domain == "Employment"), lsoa11cd, decile, rank), by = "lsoa11cd") %>% 
  filter(decile == 1) %>%  
  arrange(rank) %>% 
  select(LSOA = lsoa11nm, Ward = wd18nm, Decile = decile, Rank = rank) %>% 
  mutate(Rank = comma(Rank)) %>% 
  st_set_geometry(NULL) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation (2019), MHCLG.</h6>
<br />

***

## Education, Skills and Training
Trafford is ranked 275 out of 317 local authorities in England on the average 2019 Education, Skills and Training domain score. This represents a rise (+1) from 274 in 2015.

<br />
<h4 style="text-align: center;">Relative change in rank of average Education, Skills and Training score</h4>
<div class = "row">
<div class = "col-md-8">
```{r warning = FALSE, echo = FALSE}
df <- filter(IoD_local_authority, index_domain == "Education, Skills and Training") %>% 
  select(`Local authority` = lad19nm,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score)

ggplot() +
  geom_segment(
    data = gather(df, year, rank, -`Local authority`) %>% 
      group_by(`Local authority`) %>% 
      summarise(start = range(rank)[1], end = range(rank)[2]) %>% 
      ungroup(),
    aes(x = start, xend = end, y = fct_reorder(`Local authority`, end, .desc = TRUE), yend = `Local authority`), 
    color = "#e3e2e1", size = 2) +
  geom_point(
    data = gather(df, year, rank, -`Local authority`),
    aes(x = rank, y = fct_reorder(`Local authority`, rank, .desc = TRUE), color = year), 
    size = 4) +
  scale_x_continuous(expand = c(0.005, 0.005), 
                     limits = c(0, 317),
                     breaks = c(1, 100, 200, 300)) +
  scale_color_manual(values = c("#a6cee3", "#1f78b4"), 
                     labels = c("2015", "2019")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, colour = NULL) +
  theme_lab() +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y = element_text(hjust = 0),
        legend.position = c(0.5,1),
        legend.text = element_text(size = 10)) +
  guides(color = guide_legend(nrow = 1))
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Education, Skills and Training") %>% 
  select(`Local authority` = lad19nm,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2019`-`2015`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG</h6>
<br />

<br />
<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Education, Skills and Training domain</h4>
```{r fig.align='center'}
df <- filter(IoD_local_authority, index_domain == "Education, Skills and Training") %>% 
  select(`Local authority` = lad19nm,
         percent_bottom_decile,
         year) %>% 
  spread(year, percent_bottom_decile) %>% 
  mutate(change = round((`2019` - `2015`)*100,1),
         status = case_when(
           `2019` > `2015` ~ "Increase",
           `2019` < `2015` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, percent_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         percent_bottom_decile = percent_bottom_decile*100) %>% 
  select(`Local authority`, year, percent_bottom_decile, change, status)

ggplot(data = df, aes(x = year)) +
  geom_ribbon(aes(ymin = 0, ymax = percent_bottom_decile, fill =  factor(status))) +
  geom_point(aes(y = percent_bottom_decile), fill = "#212121", colour = "#FFFFFF", pch = 21, size = 3) +
  geom_text(aes(label = sprintf('%0.1f%%', percent_bottom_decile), y = percent_bottom_decile), vjust = -1, size = 3) +
  geom_text(aes(label = sprintf('%+0.1f%%', change)), x = 2017, y = 0, vjust = -1, fontface = 'bold', size = 3) +
  scale_fill_manual(values = c('Increase' = '#ff6961', 'Decrease' = "#77dd77", 'No change' = '#ffe662')) +
  scale_x_continuous(breaks = unique(df$year), expand = c(0.26, 0)) +
  scale_y_continuous(expand = c(0.005, 0.005), limits = c(0, 50)) +
  facet_wrap(facets = ~`Local authority`, nrow = 2) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, fill = NULL) +
  theme_lab() +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  theme(panel.grid.major.y = element_blank(),
        panel.spacing = unit(1, "lines"),
        strip.text = element_text(face = "plain", hjust = 0.5, vjust = 1),
        axis.text.y = element_blank(),
        legend.position = "none")
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG</h6>
<br />

<br />
<h4 style="text-align: center;">Education, Skills and Training domain 2019, LSOAs in Trafford</h4>
<div class = "row">
<div class = "col-md-7">
```{r fig.height=8}
left_join(trafford_lsoa, 
          select(filter(trafford_IoD, year == 2019, index_domain == "Education, Skills and Training"), lsoa11cd, index_domain, decile), by = "lsoa11cd")  %>% 
  ggplot() +
  geom_sf(aes(fill = factor(decile)), colour = "#FFFFFF", size = 0.25) +
  geom_sf(data = trafford_wards, fill = NA, alpha = 1, colour = "#000000",  size = 0.8) +
  # geom_shadowtext(data = trafford_wards, aes(x = lon, y = lat, label = wd18nm), colour = "#FFFFFF", fontface = "bold", size = 3, bg.colour = "#212121", nudge_y = 0.002) +
  scale_fill_manual(breaks = 1:10,
                    values = c("#453B52", "#454F69", "#3F657E", "#317B8D", "#239296", "#26A898", "#43BD93", "#6AD189", "#98E37D", "#CAF270"),
                    labels = c("Most\ndeprived", 2:9, "Least\ndeprived")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, 
       caption = "Contains Ordnance Survey data © Crown copyright and database right 2019", 
       fill = "") +
  coord_sf(crs = st_crs(4326), datum = NA) +
  theme_lab() +
  theme(legend.position = "bottom",
        plot.caption = element_text(hjust = 0)) +
  guides(fill = guide_legend(label.position = "bottom", 
                             label.hjust = 0,
                             direction = "horizontal",
                             nrow = 1,
                             keyheight = unit(2, units = "mm"), 
                             keywidth = unit(5, units = "mm")))
```
</div>
<div class = "col-md-5">
```{r}
left_join(trafford_lsoa, 
          select(filter(trafford_IoD, year == 2019, index_domain == "Education, Skills and Training"), lsoa11cd, decile, rank), by = "lsoa11cd") %>% 
  filter(decile == 1) %>%  
  arrange(rank) %>% 
  select(LSOA = lsoa11nm, Ward = wd18nm, Decile = decile, Rank = rank) %>% 
  mutate(Rank = comma(Rank)) %>% 
  st_set_geometry(NULL) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation (2019), MHCLG.</h6>
<br />

***

## Health and Disability
Trafford is ranked 147 out of 317 local authorities in England on the average 2019 Health and Disability domain score. This represents a rise (+20) from 127 in 2015.

<br />
<h4 style="text-align: center;">Relative change in rank of average Health and Disability score</h4>
<div class = "row">
<div class = "col-md-8">
```{r warning = FALSE, echo = FALSE}
df <- filter(IoD_local_authority, index_domain == "Health and Disability") %>% 
  select(`Local authority` = lad19nm,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score)

ggplot() +
  geom_segment(
    data = gather(df, year, rank, -`Local authority`) %>% 
      group_by(`Local authority`) %>% 
      summarise(start = range(rank)[1], end = range(rank)[2]) %>% 
      ungroup(),
    aes(x = start, xend = end, y = fct_reorder(`Local authority`, end, .desc = TRUE), yend = `Local authority`), 
    color = "#e3e2e1", size = 2) +
  geom_point(
    data = gather(df, year, rank, -`Local authority`),
    aes(x = rank, y = fct_reorder(`Local authority`, rank, .desc = TRUE), color = year), 
    size = 4) +
  scale_x_continuous(expand = c(0.005, 0.005), 
                     limits = c(0, 317),
                     breaks = c(1, 100, 200, 300)) +
  scale_color_manual(values = c("#a6cee3", "#1f78b4"), 
                     labels = c("2015", "2019")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, colour = NULL) +
  theme_lab() +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y = element_text(hjust = 0),
        legend.position = c(0.5,1),
        legend.text = element_text(size = 10)) +
  guides(color = guide_legend(nrow = 1))
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Health and Disability") %>% 
  select(`Local authority` = lad19nm,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2019`-`2015`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG</h6>
<br />

<br />
<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Health and Disability domain</h4>
```{r fig.align='center'}
df <- filter(IoD_local_authority, index_domain == "Health and Disability") %>% 
  select(`Local authority` = lad19nm,
         percent_bottom_decile,
         year) %>% 
  spread(year, percent_bottom_decile) %>% 
  mutate(change = round((`2019` - `2015`)*100,1),
         status = case_when(
           `2019` > `2015` ~ "Increase",
           `2019` < `2015` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, percent_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         percent_bottom_decile = percent_bottom_decile*100) %>% 
  select(`Local authority`, year, percent_bottom_decile, change, status)

ggplot(data = df, aes(x = year)) +
  geom_ribbon(aes(ymin = 0, ymax = percent_bottom_decile, fill =  factor(status))) +
  geom_point(aes(y = percent_bottom_decile), fill = "#212121", colour = "#FFFFFF", pch = 21, size = 3) +
  geom_text(aes(label = sprintf('%0.1f%%', percent_bottom_decile), y = percent_bottom_decile), vjust = -1, size = 3) +
  geom_text(aes(label = sprintf('%+0.1f%%', change)), x = 2017, y = 0, vjust = -1, fontface = 'bold', size = 3) +
  scale_fill_manual(values = c('Increase' = '#ff6961', 'Decrease' = "#77dd77", 'No change' = '#ffe662')) +
  scale_x_continuous(breaks = unique(df$year), expand = c(0.26, 0)) +
  scale_y_continuous(expand = c(0.005, 0.005), limits = c(0, 50)) +
  facet_wrap(facets = ~`Local authority`, nrow = 2) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, fill = NULL) +
  theme_lab() +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  theme(panel.grid.major.y = element_blank(),
        panel.spacing = unit(1, "lines"),
        strip.text = element_text(face = "plain", hjust = 0.5, vjust = 1),
        axis.text.y = element_blank(),
        legend.position = "none")
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG</h6>
<br />

<br />
<h4 style="text-align: center;">Health and Disability domain 2019, LSOAs in Trafford</h4>
<div class = "row">
<div class = "col-md-7">
```{r fig.height=8}
left_join(trafford_lsoa, 
          select(filter(trafford_IoD, year == 2019, index_domain == "Health and Disability"), lsoa11cd, index_domain, decile), by = "lsoa11cd")  %>% 
  ggplot() +
  geom_sf(aes(fill = factor(decile)), colour = "#FFFFFF", size = 0.25) +
  geom_sf(data = trafford_wards, fill = NA, alpha = 1, colour = "#000000",  size = 0.8) +
  # geom_shadowtext(data = trafford_wards, aes(x = lon, y = lat, label = wd18nm), colour = "#FFFFFF", fontface = "bold", size = 3, bg.colour = "#212121", nudge_y = 0.002) +
  scale_fill_manual(breaks = 1:10,
                    values = c("#453B52", "#454F69", "#3F657E", "#317B8D", "#239296", "#26A898", "#43BD93", "#6AD189", "#98E37D", "#CAF270"),
                    labels = c("Most\ndeprived", 2:9, "Least\ndeprived")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, 
       caption = "Contains Ordnance Survey data © Crown copyright and database right 2019", 
       fill = "") +
  coord_sf(crs = st_crs(4326), datum = NA) +
  theme_lab() +
  theme(legend.position = "bottom",
        plot.caption = element_text(hjust = 0)) +
  guides(fill = guide_legend(label.position = "bottom", 
                             label.hjust = 0,
                             direction = "horizontal",
                             nrow = 1,
                             keyheight = unit(2, units = "mm"), 
                             keywidth = unit(5, units = "mm")))
```
</div>
<div class = "col-md-5">
```{r}
left_join(trafford_lsoa, 
          select(filter(trafford_IoD, year == 2019, index_domain == "Health and Disability"), lsoa11cd, decile, rank), by = "lsoa11cd") %>% 
  filter(decile == 1) %>%  
  arrange(rank) %>% 
  select(LSOA = lsoa11nm, Ward = wd18nm, Decile = decile, Rank = rank) %>% 
  mutate(Rank = comma(Rank)) %>% 
  st_set_geometry(NULL) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation (2019), MHCLG.</h6>
<br />

***

## Crime
Trafford is ranked 138 out of 317 local authorities in England on the average 2019 Crime domain score. This represents a fall (-66) from 204 in 2015.

<br />
<h4 style="text-align: center;">Relative change in rank of average Crime score</h4>
<div class = "row">
<div class = "col-md-8">
```{r warning = FALSE, echo = FALSE}
df <- filter(IoD_local_authority, index_domain == "Crime") %>% 
  select(`Local authority` = lad19nm,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score)

ggplot() +
  geom_segment(
    data = gather(df, year, rank, -`Local authority`) %>% 
      group_by(`Local authority`) %>% 
      summarise(start = range(rank)[1], end = range(rank)[2]) %>% 
      ungroup(),
    aes(x = start, xend = end, y = fct_reorder(`Local authority`, end, .desc = TRUE), yend = `Local authority`), 
    color = "#e3e2e1", size = 2) +
  geom_point(
    data = gather(df, year, rank, -`Local authority`),
    aes(x = rank, y = fct_reorder(`Local authority`, rank, .desc = TRUE), color = year), 
    size = 4) +
  scale_x_continuous(expand = c(0.005, 0.005), 
                     limits = c(0, 317),
                     breaks = c(1, 100, 200, 300)) +
  scale_color_manual(values = c("#a6cee3", "#1f78b4"), 
                     labels = c("2015", "2019")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, colour = NULL) +
  theme_lab() +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y = element_text(hjust = 0),
        legend.position = c(0.5,1),
        legend.text = element_text(size = 10)) +
  guides(color = guide_legend(nrow = 1))
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Crime") %>% 
  select(`Local authority` = lad19nm,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2019`-`2015`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG</h6>
<br />

<br />
<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Crime domain</h4>
```{r fig.align='center'}
df <- filter(IoD_local_authority, index_domain == "Crime") %>% 
  select(`Local authority` = lad19nm,
         percent_bottom_decile,
         year) %>% 
  spread(year, percent_bottom_decile) %>% 
  mutate(change = round((`2019` - `2015`)*100,1),
         status = case_when(
           `2019` > `2015` ~ "Increase",
           `2019` < `2015` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, percent_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         percent_bottom_decile = percent_bottom_decile*100) %>% 
  select(`Local authority`, year, percent_bottom_decile, change, status)

ggplot(data = df, aes(x = year)) +
  geom_ribbon(aes(ymin = 0, ymax = percent_bottom_decile, fill =  factor(status))) +
  geom_point(aes(y = percent_bottom_decile), fill = "#212121", colour = "#FFFFFF", pch = 21, size = 3) +
  geom_text(aes(label = sprintf('%0.1f%%', percent_bottom_decile), y = percent_bottom_decile), vjust = -1, size = 3) +
  geom_text(aes(label = sprintf('%+0.1f%%', change)), x = 2017, y = 0, vjust = -1, fontface = 'bold', size = 3) +
  scale_fill_manual(values = c('Increase' = '#ff6961', 'Decrease' = "#77dd77", 'No change' = '#ffe662')) +
  scale_x_continuous(breaks = unique(df$year), expand = c(0.26, 0)) +
  scale_y_continuous(expand = c(0.005, 0.005), limits = c(0, 70)) +
  facet_wrap(facets = ~`Local authority`, nrow = 2) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, fill = NULL) +
  theme_lab() +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  theme(panel.grid.major.y = element_blank(),
        panel.spacing = unit(1, "lines"),
        strip.text = element_text(face = "plain", hjust = 0.5, vjust = 1),
        axis.text.y = element_blank(),
        legend.position = "none")
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG</h6>
<br />

<br />
<h4 style="text-align: center;">Crime domain 2019, LSOAs in Trafford</h4>
<div class = "row">
<div class = "col-md-7">
```{r fig.height=8}
left_join(trafford_lsoa, 
          select(filter(trafford_IoD, year == 2019, index_domain == "Crime"), lsoa11cd, index_domain, decile), by = "lsoa11cd")  %>% 
  ggplot() +
  geom_sf(aes(fill = factor(decile)), colour = "#FFFFFF", size = 0.25) +
  geom_sf(data = trafford_wards, fill = NA, alpha = 1, colour = "#000000",  size = 0.8) +
  # geom_shadowtext(data = trafford_wards, aes(x = lon, y = lat, label = wd18nm), colour = "#FFFFFF", fontface = "bold", size = 3, bg.colour = "#212121", nudge_y = 0.002) +
  scale_fill_manual(breaks = 1:10,
                    values = c("#453B52", "#454F69", "#3F657E", "#317B8D", "#239296", "#26A898", "#43BD93", "#6AD189", "#98E37D", "#CAF270"),
                    labels = c("Most\ndeprived", 2:9, "Least\ndeprived")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, 
       caption = "Contains Ordnance Survey data © Crown copyright and database right 2019", 
       fill = "") +
  coord_sf(crs = st_crs(4326), datum = NA) +
  theme_lab() +
  theme(legend.position = "bottom",
        plot.caption = element_text(hjust = 0)) +
  guides(fill = guide_legend(label.position = "bottom", 
                             label.hjust = 0,
                             direction = "horizontal",
                             nrow = 1,
                             keyheight = unit(2, units = "mm"), 
                             keywidth = unit(5, units = "mm")))
```
</div>
<div class = "col-md-5">
```{r}
left_join(trafford_lsoa, 
          select(filter(trafford_IoD, year == 2019, index_domain == "Crime"), lsoa11cd, decile, rank), by = "lsoa11cd") %>% 
  filter(decile == 1) %>%  
  arrange(rank) %>% 
  select(LSOA = lsoa11nm, Ward = wd18nm, Decile = decile, Rank = rank) %>% 
  mutate(Rank = comma(Rank)) %>% 
  st_set_geometry(NULL) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation (2019), MHCLG.</h6>
<br />

***

## Barriers to Housing and Services
Trafford is ranked 289 out of 317 local authorities in England on the average 2019 Barriers to Housing and Services domain score. This represents a fall (-3) from 292 in 2015.

<br />
<h4 style="text-align: center;">Relative change in rank of average Barriers to Housing and Services score</h4>
<div class = "row">
<div class = "col-md-8">
```{r warning = FALSE, echo = FALSE}
df <- filter(IoD_local_authority, index_domain == "Barriers to Housing and Services") %>% 
  select(`Local authority` = lad19nm,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score)

ggplot() +
  geom_segment(
    data = gather(df, year, rank, -`Local authority`) %>% 
      group_by(`Local authority`) %>% 
      summarise(start = range(rank)[1], end = range(rank)[2]) %>% 
      ungroup(),
    aes(x = start, xend = end, y = fct_reorder(`Local authority`, end, .desc = TRUE), yend = `Local authority`), 
    color = "#e3e2e1", size = 2) +
  geom_point(
    data = gather(df, year, rank, -`Local authority`),
    aes(x = rank, y = fct_reorder(`Local authority`, rank, .desc = TRUE), color = year), 
    size = 4) +
  scale_x_continuous(expand = c(0.005, 0.005), 
                     limits = c(0, 317),
                     breaks = c(1, 100, 200, 300)) +
  scale_color_manual(values = c("#a6cee3", "#1f78b4"), 
                     labels = c("2015", "2019")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, colour = NULL) +
  theme_lab() +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y = element_text(hjust = 0),
        legend.position = c(0.5,1),
        legend.text = element_text(size = 10)) +
  guides(color = guide_legend(nrow = 1))
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Barriers to Housing and Services") %>% 
  select(`Local authority` = lad19nm,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2019`-`2015`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG</h6>
<br />

<br />
<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Barriers to Housing and Services domain</h4>
```{r fig.align='center'}
df <- filter(IoD_local_authority, index_domain == "Barriers to Housing and Services") %>% 
  select(`Local authority` = lad19nm,
         percent_bottom_decile,
         year) %>% 
  spread(year, percent_bottom_decile) %>% 
  mutate(change = round((`2019` - `2015`)*100,1),
         status = case_when(
           `2019` > `2015` ~ "Increase",
           `2019` < `2015` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, percent_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         percent_bottom_decile = percent_bottom_decile*100) %>% 
  select(`Local authority`, year, percent_bottom_decile, change, status)

ggplot(data = df, aes(x = year)) +
  geom_ribbon(aes(ymin = 0, ymax = percent_bottom_decile, fill =  factor(status))) +
  geom_point(aes(y = percent_bottom_decile), fill = "#212121", colour = "#FFFFFF", pch = 21, size = 3) +
  geom_text(aes(label = sprintf('%0.1f%%', percent_bottom_decile), y = percent_bottom_decile), vjust = -1, size = 3) +
  geom_text(aes(label = sprintf('%+0.1f%%', change)), x = 2017, y = 0, vjust = -1, fontface = 'bold', size = 3) +
  scale_fill_manual(values = c('Increase' = '#ff6961', 'Decrease' = "#77dd77", 'No change' = '#ffe662')) +
  scale_x_continuous(breaks = unique(df$year), expand = c(0.26, 0)) +
  scale_y_continuous(expand = c(0.005, 0.005), limits = c(0, 50)) +
  facet_wrap(facets = ~`Local authority`, nrow = 2) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, fill = NULL) +
  theme_lab() +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  theme(panel.grid.major.y = element_blank(),
        panel.spacing = unit(1, "lines"),
        strip.text = element_text(face = "plain", hjust = 0.5, vjust = 1),
        axis.text.y = element_blank(),
        legend.position = "none")
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG</h6>
<br />

<br />
<h4 style="text-align: center;">Barriers to Housing and Services domain 2019, LSOAs in Trafford</h4>
<div class = "row">
<div class = "col-md-7">
```{r fig.height=8}
left_join(trafford_lsoa, 
          select(filter(trafford_IoD, year == 2019, index_domain == "Barriers to Housing and Services"), lsoa11cd, index_domain, decile), by = "lsoa11cd")  %>% 
  ggplot() +
  geom_sf(aes(fill = factor(decile)), colour = "#FFFFFF", size = 0.25) +
  geom_sf(data = trafford_wards, fill = NA, alpha = 1, colour = "#000000",  size = 0.8) +
  # geom_shadowtext(data = trafford_wards, aes(x = lon, y = lat, label = wd18nm), colour = "#FFFFFF", fontface = "bold", size = 3, bg.colour = "#212121", nudge_y = 0.002) +
  scale_fill_manual(breaks = 1:10,
                    values = c("#453B52", "#454F69", "#3F657E", "#317B8D", "#239296", "#26A898", "#43BD93", "#6AD189", "#98E37D", "#CAF270"),
                    labels = c("Most\ndeprived", 2:9, "Least\ndeprived")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, 
       caption = "Contains Ordnance Survey data © Crown copyright and database right 2019", 
       fill = "") +
  coord_sf(crs = st_crs(4326), datum = NA) +
  theme_lab() +
  theme(legend.position = "bottom",
        plot.caption = element_text(hjust = 0)) +
  guides(fill = guide_legend(label.position = "bottom", 
                             label.hjust = 0,
                             direction = "horizontal",
                             nrow = 1,
                             keyheight = unit(2, units = "mm"), 
                             keywidth = unit(5, units = "mm")))
```
</div>
<div class = "col-md-5">
```{r}
left_join(trafford_lsoa, 
          select(filter(trafford_IoD, year == 2019, index_domain == "Barriers to Housing and Services"), lsoa11cd, decile, rank), by = "lsoa11cd") %>% 
  filter(decile == 1) %>%  
  arrange(rank) %>% 
  select(LSOA = lsoa11nm, Ward = wd18nm, Decile = decile, Rank = rank) %>% 
  mutate(Rank = comma(Rank)) %>% 
  st_set_geometry(NULL) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation (2019), MHCLG.</h6>
<br />

***

## Living Environment
Trafford is ranked 131 out of 317 local authorities in England on the average 2019 Living Environment domain score. This represents a fall (-35) from 166 in 2015.

<br />
<h4 style="text-align: center;">Relative change in rank of average Living Environment score</h4>
<div class = "row">
<div class = "col-md-8">
```{r warning = FALSE, echo = FALSE}
df <- filter(IoD_local_authority, index_domain == "Living Environment") %>% 
  select(`Local authority` = lad19nm,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score)

ggplot() +
  geom_segment(
    data = gather(df, year, rank, -`Local authority`) %>% 
      group_by(`Local authority`) %>% 
      summarise(start = range(rank)[1], end = range(rank)[2]) %>% 
      ungroup(),
    aes(x = start, xend = end, y = fct_reorder(`Local authority`, end, .desc = TRUE), yend = `Local authority`), 
    color = "#e3e2e1", size = 2) +
  geom_point(
    data = gather(df, year, rank, -`Local authority`),
    aes(x = rank, y = fct_reorder(`Local authority`, rank, .desc = TRUE), color = year), 
    size = 4) +
  scale_x_continuous(expand = c(0.005, 0.005), 
                     limits = c(0, 317),
                     breaks = c(1, 100, 200, 300)) +
  scale_color_manual(values = c("#a6cee3", "#1f78b4"), 
                     labels = c("2015", "2019")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, colour = NULL) +
  theme_lab() +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y = element_text(hjust = 0),
        legend.position = c(0.5,1),
        legend.text = element_text(size = 10)) +
  guides(color = guide_legend(nrow = 1))
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Living Environment") %>% 
  select(`Local authority` = lad19nm,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2019`-`2015`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG</h6>
<br />

<br />
<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of Living Environment domain</h4>
```{r fig.align='center'}
df <- filter(IoD_local_authority, index_domain == "Living Environment") %>% 
  select(`Local authority` = lad19nm,
         percent_bottom_decile,
         year) %>% 
  spread(year, percent_bottom_decile) %>% 
  mutate(change = round((`2019` - `2015`)*100,1),
         status = case_when(
           `2019` > `2015` ~ "Increase",
           `2019` < `2015` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, percent_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         percent_bottom_decile = percent_bottom_decile*100) %>% 
  select(`Local authority`, year, percent_bottom_decile, change, status)

ggplot(data = df, aes(x = year)) +
  geom_ribbon(aes(ymin = 0, ymax = percent_bottom_decile, fill =  factor(status))) +
  geom_point(aes(y = percent_bottom_decile), fill = "#212121", colour = "#FFFFFF", pch = 21, size = 3) +
  geom_text(aes(label = sprintf('%0.1f%%', percent_bottom_decile), y = percent_bottom_decile), vjust = -1, size = 3) +
  geom_text(aes(label = sprintf('%+0.1f%%', change)), x = 2017, y = 0, vjust = -1, fontface = 'bold', size = 3) +
  scale_fill_manual(values = c('Increase' = '#ff6961', 'Decrease' = "#77dd77", 'No change' = '#ffe662')) +
  scale_x_continuous(breaks = unique(df$year), expand = c(0.26, 0)) +
  scale_y_continuous(expand = c(0.005, 0.005), limits = c(0, 50)) +
  facet_wrap(facets = ~`Local authority`, nrow = 2) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, fill = NULL) +
  theme_lab() +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  theme(panel.grid.major.y = element_blank(),
        panel.spacing = unit(1, "lines"),
        strip.text = element_text(face = "plain", hjust = 0.5, vjust = 1),
        axis.text.y = element_blank(),
        legend.position = "none")
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG</h6>
<br />

<br />
<h4 style="text-align: center;">Living Environment domain 2019, LSOAs in Trafford</h4>
<div class = "row">
<div class = "col-md-7">
```{r fig.height=8}
left_join(trafford_lsoa, 
          select(filter(trafford_IoD, year == 2019, index_domain == "Living Environment"), lsoa11cd, index_domain, decile), by = "lsoa11cd")  %>% 
  ggplot() +
  geom_sf(aes(fill = factor(decile)), colour = "#FFFFFF", size = 0.25) +
  geom_sf(data = trafford_wards, fill = NA, alpha = 1, colour = "#000000",  size = 0.8) +
  # geom_shadowtext(data = trafford_wards, aes(x = lon, y = lat, label = wd18nm), colour = "#FFFFFF", fontface = "bold", size = 3, bg.colour = "#212121", nudge_y = 0.002) +
  scale_fill_manual(breaks = 1:10,
                    values = c("#453B52", "#454F69", "#3F657E", "#317B8D", "#239296", "#26A898", "#43BD93", "#6AD189", "#98E37D", "#CAF270"),
                    labels = c("Most\ndeprived", 2:9, "Least\ndeprived")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, 
       caption = "Contains Ordnance Survey data © Crown copyright and database right 2019", 
       fill = "") +
  coord_sf(crs = st_crs(4326), datum = NA) +
  theme_lab() +
  theme(legend.position = "bottom",
        plot.caption = element_text(hjust = 0)) +
  guides(fill = guide_legend(label.position = "bottom", 
                             label.hjust = 0,
                             direction = "horizontal",
                             nrow = 1,
                             keyheight = unit(2, units = "mm"), 
                             keywidth = unit(5, units = "mm")))
```
</div>
<div class = "col-md-5">
```{r}
left_join(trafford_lsoa, 
          select(filter(trafford_IoD, year == 2019, index_domain == "Living Environment"), lsoa11cd, decile, rank), by = "lsoa11cd") %>% 
  filter(decile == 1) %>%  
  arrange(rank) %>% 
  select(LSOA = lsoa11nm, Ward = wd18nm, Decile = decile, Rank = rank) %>% 
  mutate(Rank = comma(Rank)) %>% 
  st_set_geometry(NULL) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation (2019), MHCLG.</h6>
<br />

***

## Income Deprivation Affecting Children
Trafford is ranked 214 out of 317 local authorities in England on the average 2019 Income Deprivation Affecting Children Index score. This represents an increase (+13) from 201 in 2015.

<br />
<h4 style="text-align: center;">Relative change in rank of average IDACI score</h4>
<div class = "row">
<div class = "col-md-8">
```{r warning = FALSE, echo = FALSE}
df <- filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Children") %>% 
  select(`Local authority` = lad19nm,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score)

ggplot() +
  geom_segment(
    data = gather(df, year, rank, -`Local authority`) %>% 
      group_by(`Local authority`) %>% 
      summarise(start = range(rank)[1], end = range(rank)[2]) %>% 
      ungroup(),
    aes(x = start, xend = end, y = fct_reorder(`Local authority`, end, .desc = TRUE), yend = `Local authority`), 
    color = "#e3e2e1", size = 2) +
  geom_point(
    data = gather(df, year, rank, -`Local authority`),
    aes(x = rank, y = fct_reorder(`Local authority`, rank, .desc = TRUE), color = year), 
    size = 4) +
  scale_x_continuous(expand = c(0.005, 0.005), 
                     limits = c(0, 317),
                     breaks = c(1, 100, 200, 300)) +
  scale_color_manual(values = c("#a6cee3", "#1f78b4"), 
                     labels = c("2015", "2019")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, colour = NULL) +
  theme_lab() +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y = element_text(hjust = 0),
        legend.position = c(0.5,1),
        legend.text = element_text(size = 10)) +
  guides(color = guide_legend(nrow = 1))
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Children") %>% 
  select(`Local authority` = lad19nm,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2019`-`2015`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG</h6>
<br />

<br />
<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of IDACI</h4>
```{r fig.align='center'}
df <- filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Children") %>% 
  select(`Local authority` = lad19nm,
         percent_bottom_decile,
         year) %>% 
  spread(year, percent_bottom_decile) %>% 
  mutate(change = round((`2019` - `2015`)*100,1),
         status = case_when(
           `2019` > `2015` ~ "Increase",
           `2019` < `2015` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, percent_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         percent_bottom_decile = percent_bottom_decile*100) %>% 
  select(`Local authority`, year, percent_bottom_decile, change, status)

ggplot(data = df, aes(x = year)) +
  geom_ribbon(aes(ymin = 0, ymax = percent_bottom_decile, fill =  factor(status))) +
  geom_point(aes(y = percent_bottom_decile), fill = "#212121", colour = "#FFFFFF", pch = 21, size = 3) +
  geom_text(aes(label = sprintf('%0.1f%%', percent_bottom_decile), y = percent_bottom_decile), vjust = -1, size = 3) +
  geom_text(aes(label = sprintf('%+0.1f%%', change)), x = 2017, y = 0, vjust = -1, fontface = 'bold', size = 3) +
  scale_fill_manual(values = c('Increase' = '#ff6961', 'Decrease' = "#77dd77", 'No change' = '#ffe662')) +
  scale_x_continuous(breaks = unique(df$year), expand = c(0.26, 0)) +
  scale_y_continuous(expand = c(0.005, 0.005), limits = c(0, 50)) +
  facet_wrap(facets = ~`Local authority`, nrow = 2) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, fill = NULL) +
  theme_lab() +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  theme(panel.grid.major.y = element_blank(),
        panel.spacing = unit(1, "lines"),
        strip.text = element_text(face = "plain", hjust = 0.5, vjust = 1),
        axis.text.y = element_blank(),
        legend.position = "none")
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG</h6>
<br />

<br />
<h4 style="text-align: center;">IDACI 2019, LSOAs in Trafford</h4>
<div class = "row">
<div class = "col-md-7">
```{r fig.height=8}
left_join(trafford_lsoa, 
          select(filter(trafford_IoD, year == 2019, index_domain == "Income Deprivation Affecting Children"), lsoa11cd, index_domain, decile), by = "lsoa11cd")  %>% 
  ggplot() +
  geom_sf(aes(fill = factor(decile)), colour = "#FFFFFF", size = 0.25) +
  geom_sf(data = trafford_wards, fill = NA, alpha = 1, colour = "#000000",  size = 0.8) +
  # geom_shadowtext(data = trafford_wards, aes(x = lon, y = lat, label = wd18nm), colour = "#FFFFFF", fontface = "bold", size = 3, bg.colour = "#212121", nudge_y = 0.002) +
  scale_fill_manual(breaks = 1:10,
                    values = c("#453B52", "#454F69", "#3F657E", "#317B8D", "#239296", "#26A898", "#43BD93", "#6AD189", "#98E37D", "#CAF270"),
                    labels = c("Most\ndeprived", 2:9, "Least\ndeprived")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, 
       caption = "Contains Ordnance Survey data © Crown copyright and database right 2019", 
       fill = "") +
  coord_sf(crs = st_crs(4326), datum = NA) +
  theme_lab() +
  theme(legend.position = "bottom",
        plot.caption = element_text(hjust = 0)) +
  guides(fill = guide_legend(label.position = "bottom", 
                             label.hjust = 0,
                             direction = "horizontal",
                             nrow = 1,
                             keyheight = unit(2, units = "mm"), 
                             keywidth = unit(5, units = "mm")))
```
</div>
<div class = "col-md-5">
```{r}
left_join(trafford_lsoa, 
          select(filter(trafford_IoD, year == 2019, index_domain == "Income Deprivation Affecting Children"), lsoa11cd, decile, rank), by = "lsoa11cd") %>% 
  filter(decile == 1) %>%  
  arrange(rank) %>% 
  select(LSOA = lsoa11nm, Ward = wd18nm, Decile = decile, Rank = rank) %>% 
  mutate(Rank = comma(Rank)) %>% 
  st_set_geometry(NULL) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation (2019), MHCLG.</h6>
<br />

***

## Income Deprivation Affecting Older People	
Trafford is ranked 134 out of 317 local authorities in England on the average 2019 Income Deprivation Affecting Older People Index score. This represents an increase (5) from 129 in 2015.

<br />
<h4 style="text-align: center;">Relative change in rank of average IDAOPI score</h4>
<div class = "row">
<div class = "col-md-8">
```{r warning = FALSE, echo = FALSE}
df <- filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Older People") %>% 
  select(`Local authority` = lad19nm,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score)

ggplot() +
  geom_segment(
    data = gather(df, year, rank, -`Local authority`) %>% 
      group_by(`Local authority`) %>% 
      summarise(start = range(rank)[1], end = range(rank)[2]) %>% 
      ungroup(),
    aes(x = start, xend = end, y = fct_reorder(`Local authority`, end, .desc = TRUE), yend = `Local authority`), 
    color = "#e3e2e1", size = 2) +
  geom_point(
    data = gather(df, year, rank, -`Local authority`),
    aes(x = rank, y = fct_reorder(`Local authority`, rank, .desc = TRUE), color = year), 
    size = 4) +
  scale_x_continuous(expand = c(0.005, 0.005), 
                     limits = c(0, 317),
                     breaks = c(1, 100, 200, 300)) +
  scale_color_manual(values = c("#a6cee3", "#1f78b4"), 
                     labels = c("2015", "2019")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, colour = NULL) +
  theme_lab() +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y = element_text(hjust = 0),
        legend.position = c(0.5,1),
        legend.text = element_text(size = 10)) +
  guides(color = guide_legend(nrow = 1))
```
</div>
<div class = "col-md-4">
```{r warning = FALSE, echo=FALSE}
filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Older People") %>% 
  select(`Local authority` = lad19nm,
         rank_average_score,
         year) %>%
  spread(year, rank_average_score) %>% 
  mutate(Change = `2019`-`2015`) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG</h6>
<br />

<br />
<h4 style="text-align: center;">Change in percentage of LSOAs in 1st decile of IDAOPI</h4>
```{r fig.align='center'}
df <- filter(IoD_local_authority, index_domain == "Income Deprivation Affecting Older People") %>% 
  select(`Local authority` = lad19nm,
         percent_bottom_decile,
         year) %>% 
  spread(year, percent_bottom_decile) %>% 
  mutate(change = round((`2019` - `2015`)*100,1),
         status = case_when(
           `2019` > `2015` ~ "Increase",
           `2019` < `2015` ~ "Decrease",
           TRUE ~ "No change")) %>% 
  gather(year, percent_bottom_decile, -`Local authority`, -status, -change) %>% 
  mutate(year = as.integer(year),
         percent_bottom_decile = percent_bottom_decile*100) %>% 
  select(`Local authority`, year, percent_bottom_decile, change, status)

ggplot(data = df, aes(x = year)) +
  geom_ribbon(aes(ymin = 0, ymax = percent_bottom_decile, fill =  factor(status))) +
  geom_point(aes(y = percent_bottom_decile), fill = "#212121", colour = "#FFFFFF", pch = 21, size = 3) +
  geom_text(aes(label = sprintf('%0.1f%%', percent_bottom_decile), y = percent_bottom_decile), vjust = -1, size = 3) +
  geom_text(aes(label = sprintf('%+0.1f%%', change)), x = 2017, y = 0, vjust = -1, fontface = 'bold', size = 3) +
  scale_fill_manual(values = c('Increase' = '#ff6961', 'Decrease' = "#77dd77", 'No change' = '#ffe662')) +
  scale_x_continuous(breaks = unique(df$year), expand = c(0.26, 0)) +
  scale_y_continuous(expand = c(0.005, 0.005), limits = c(0, 50)) +
  facet_wrap(facets = ~`Local authority`, nrow = 2) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, fill = NULL) +
  theme_lab() +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  theme(panel.grid.major.y = element_blank(),
        panel.spacing = unit(1, "lines"),
        strip.text = element_text(face = "plain", hjust = 0.5, vjust = 1),
        axis.text.y = element_blank(),
        legend.position = "none")
```
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation, MHCLG</h6>
<br />

<br />
<h4 style="text-align: center;">IDAOPI 2019, LSOAs in Trafford</h4>
<div class = "row">
<div class = "col-md-7">
```{r fig.height=8}
left_join(trafford_lsoa, 
          select(filter(trafford_IoD, year == 2019, index_domain == "Income Deprivation Affecting Older People"), lsoa11cd, index_domain, decile), by = "lsoa11cd")  %>% 
  ggplot() +
  geom_sf(aes(fill = factor(decile)), colour = "#FFFFFF", size = 0.25) +
  geom_sf(data = trafford_wards, fill = NA, alpha = 1, colour = "#000000",  size = 0.8) +
  # geom_shadowtext(data = trafford_wards, aes(x = lon, y = lat, label = wd18nm), colour = "#FFFFFF", fontface = "bold", size = 3, bg.colour = "#212121", nudge_y = 0.002) +
  scale_fill_manual(breaks = 1:10,
                    values = c("#453B52", "#454F69", "#3F657E", "#317B8D", "#239296", "#26A898", "#43BD93", "#6AD189", "#98E37D", "#CAF270"),
                    labels = c("Most\ndeprived", 2:9, "Least\ndeprived")) +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, 
       caption = "Contains Ordnance Survey data © Crown copyright and database right 2019", 
       fill = "") +
  coord_sf(crs = st_crs(4326), datum = NA) +
  theme_lab() +
  theme(legend.position = "bottom",
        plot.caption = element_text(hjust = 0)) +
  guides(fill = guide_legend(label.position = "bottom", 
                             label.hjust = 0,
                             direction = "horizontal",
                             nrow = 1,
                             keyheight = unit(2, units = "mm"), 
                             keywidth = unit(5, units = "mm")))
```
</div>
<div class = "col-md-5">
```{r}
left_join(trafford_lsoa, 
          select(filter(trafford_IoD, year == 2019, index_domain == "Income Deprivation Affecting Older People"), lsoa11cd, decile, rank), by = "lsoa11cd") %>% 
  filter(decile == 1) %>%  
  arrange(rank) %>% 
  select(LSOA = lsoa11nm, Ward = wd18nm, Decile = decile, Rank = rank) %>% 
  mutate(Rank = comma(Rank)) %>% 
  st_set_geometry(NULL) %>% 
  kable() %>% 
  kable_styling(font_size = 12) %>%
  row_spec(0, bold = T, color = "#757575")
```
</div>
</div>
<h6 style="text-align: right; color: #7f7f7f;">Source: English Indices of Deprivation (2019), MHCLG.</h6>
<br />