---
title: "Twitter analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

Last updated: `r format(Sys.time(), '%d %B %Y')`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 8, fig.height = 4)

library(knitr) ; library(tidyverse) ; library(rtweet) ; library(tidytext) ; library(reshape2) ; library(igraph) ; library(grid) ; library(ggraph) ; library(lubridate)

# ggplot2 theme
theme_x <- function () { 
  theme_minimal(base_size = 14, base_family = "Open Sans") %+replace% 
    theme(
      panel.grid.major.y = element_blank(),
      panel.grid.minor = element_blank(),
      plot.title = element_text(size = 14, face = "bold", hjust = 0),
      plot.subtitle = element_text(size = 10, hjust = 0, margin = margin(9, 0, 9, 0)),
      axis.title = element_text(size = 10, hjust = 1)
    )
}
```

```{r mentions, include=FALSE}
mentions <- read_csv("data/mentions.csv") %>% 
  mutate(created_at = as.Date(as.POSIXct(created_at)))

# tokenise
tidy_mentions <- mentions %>%
  mutate(text = str_remove_all(text, "&amp;|&lt;|&gt;")) %>% 
  unnest_tokens(word, text, token = "tweets") %>% 
  filter(!word %in% stop_words$word, # remove stop words
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>% # remove numbers
  select(screen_name, word, created_at, is_retweet, hashtags)
```

```{r tweets, include=FALSE}
tweets <- read_csv("data/tweets.csv") %>% 
  mutate(created_at = as.POSIXct(created_at))

# tokenise
tidy_tweets <- tweets %>% 
  mutate(text = str_remove_all(text, "&amp;|&lt;|&gt;")) %>% 
  unnest_tokens(word, text, token = "tweets") %>% 
  filter(!word %in% stop_words$word, # remove stop words
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>% # remove numbers
  select(screen_name, word, created_at, favorite_count, retweet_count, is_retweet, hashtags, media_type)
```

```{r likes, include=FALSE}
likes <- read_csv("data/likes.csv") %>% 
  mutate(created_at = as.Date(as.POSIXct(created_at)))
```

This report is a brief exploratory Twitter analysis. [Part 1](#part1) explores tweets mentioning @TraffordCouncil or the phrase "Trafford Council". [Part 2](#part2) examines the last 1,000 tweets sent by @TraffordCouncil.

### Part 1: tweets mentioning @TraffordCouncil or "Trafford Council" {#part1}

There were `r nrow(mentions) ` tweets - excluding retweets and replies - that mention <a href="https://twitter.com/traffordcouncil" target="_blank">@TraffordCouncil</a> or include the phrase <a href="https://twitter.com/search?q=%22trafford%20council%22" target="_blank">"Trafford Council"</a> between `r paste(format(min(mentions$created_at), "%d"), format(max(mentions$created_at), "%d %B %Y"), sep = " and " )`. You can explore these tweets in an <a href="mentions.html" target="_blank">interactive table</a>.

<br>

<div class = "row">
<div class = "col-md-4">
<span style="font-weight:bold;">Word frequency</span>
<p style="word-wrap:break-word;">The most frequent word in tweets mentioning <a href="https://twitter.com/traffordcouncil" target="_blank">@TraffordCouncil</a> or the phrase <a href="https://twitter.com/search?q=%22trafford%20council%22" target="_blank">"Trafford Council"</a> was "vote" (27). This is unsurprising given the general election was held in the middle of the period examined on 12 December 2019. The prominence of terms such as "bins" (14) and "emptied" (12) suggests that bin collections were important concerns over the period too.</p>
</div>
<div class = "col-md-8">
```{r}
tidy_mentions %>% 
  filter(!str_detect(word, "@\\S+"),
         !word %in% c("council", "trafford")) %>% # excluding "council" and "trafford"
  count(word, sort = TRUE) %>% 
  filter(n > 10) %>% 
  mutate(word = reorder(word, n)) %>% 
  ggplot(aes(word, n)) +
  geom_col(fill = "#a6cee3") +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  scale_y_continuous(expand = c(0.005, 0.005)) +
  labs(x = NULL, y = NULL,
       title = "Most common words",
       subtitle = paste0("@TraffordCouncil mentions, ", format(min(mentions$created_at), "%d"), " - ", format(max(mentions$created_at),"%d %b %Y"))) +
  coord_flip() +
  theme_x()
```
</div>
</div>

<br>

<div class = "row">
<div class = "col-md-4">
<span style="font-weight:bold;">Hashtag frequency</span>
<p style="word-wrap:break-word;">The most commonly used hashtag over the period examined was <a href="https://twitter.com/search?q=%23ge2019&src=typed_query" target="_blank">#ge2019</a>, which refers to the 2019 General Election.</p>
</div>
<div class = "col-md-8">
```{r}
tidy_mentions %>% 
  filter(str_detect(word, "^#")) %>% 
  count(word, sort = TRUE) %>% 
  top_n(5) %>% 
  mutate(word = reorder(word, n)) %>% 
  ggplot(aes(word, n)) +
  geom_col(fill = "#7570b3") +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  scale_y_continuous(expand = c(0.005, 0.005)) +
  labs(x = NULL, y = NULL,
       title = "Most common hashtags",
       subtitle = paste0("@TraffordCouncil mentions, ", format(min(mentions$created_at), "%d"), " - ", format(max(mentions$created_at),"%d %b %Y"))) +
  coord_flip() +
  theme_x()
```
</div>
</div>

<br>

<div class = "row">
<div class = "col-md-4">
<span style="font-weight:bold;">Sentiment analysis</span>
<p style="word-wrap:break-word;">Sentiment analysis is a type of text mining that assigns sentiment to text. Here the most common words in tweets mentioning <a href="https://twitter.com/traffordcouncil" target="_blank">@TraffordCouncil</a> or <a href="https://twitter.com/search?q=%22trafford%20council%22" target="_blank">"Trafford Council"</a> have been classified as either positive or negative. If we discount "hale" which refers to a ward we find positive sentiment in terms such as "love", "beautiful" etc. Negative sentiment can be found in "drains" and "missed". The term "marginal" is likely to refer to the General Election.</p>
</div>
<div class = "col-md-8">
```{r}
tidy_mentions %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  top_n(10) %>%
  mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = word, y = n, fill = sentiment)) +
  geom_col(alpha = 0.7) +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  scale_fill_manual(values = c("positive" = "#31a354", "negative" = "#de2d26")) + 
  scale_y_continuous(expand = c(0.005, 0.005)) +
  coord_flip() +
  labs(x = NULL, y = NULL, 
       title = "Negative and positive sentiment in top words",
       subtitle = paste0("@TraffordCouncil mentions, ", format(min(mentions$created_at), "%d"), " - ", format(max(mentions$created_at),"%d %b %Y"))) +
  theme_x() +
  theme(legend.position = "none")
```
</div>
</div>

<br>

<div class = "row">
<div class = "col-md-4">
<span style="font-weight:bold;">Network analysis</span>
<p style="word-wrap:break-word;">This network graph visualises the relationship between words in tweets that mentioned Trafford Council. The words "Trafford" and "Council" have the strongest relationship as indicated by the darker coloured arrow. Other linked words include "polling station", "monkey puzzle" and "wedding ring found in ballot box".</p>
</div>
<div class = "col-md-8">
```{r}
bigram_graph <- mentions %>%
  mutate(text = str_remove_all(text, "&amp;|&lt;|&gt;"),
         text = str_remove_all(text, "\\s?(f|ht)(tp)(s?)(://)([^\\.]*)[\\.|/](\\S*)")) %>% # remove URLs
  unnest_tokens(word, text, token = "tweets") %>%
  filter(!word %in% stop_words$word, # remove stop words
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]"),
         !str_detect(word, "#\\S+"), # remove hashtags
         !str_detect(word, "@\\S+")) %>% # remove @user
  select(word) %>% 
  unnest_tokens(bigram, word, token = "ngrams", n = 2) %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  count(word1, word2, sort = TRUE) %>%
  select(word1, word2, n) %>%
  filter(n > 2) %>%
  graph_from_data_frame()

#bigram_graph <- bigram_graph -V(bigram_graph)[3] # remove NA node

a <- arrow(type = "closed", 
                 length = unit(.15, "inches"))

ggraph(bigram_graph, layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE,
                 arrow = a, end_cap = circle(.07, 'inches')) +
  geom_node_point(color = "#b2df8a", size = 5) +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
  theme_void()  
```
</div>
</div>

***

### Part 2: tweets by @TraffordCouncil {#part2}

This section explores the last 1,000 tweets of <a href="https://twitter.com/traffordcouncil" target="_blank">@TraffordCouncil</a> which cover the period between `r paste(format(min(tweets$created_at), "%d %B"), format(max(tweets$created_at), "%d %B %Y"), sep = " and " )`. You can explore these tweets in an <a href="tweets.html" target="_blank">interactive table</a>.

<br>

<div class = "row">
<div class = "col-md-4">
<span style="font-weight:bold;">Twitter activity</span>
<p style="word-wrap:break-word;">The bulk of <a href="https://twitter.com/traffordcouncil" target="_blank">@TraffordCouncil</a>'s twitter activity occurs between 0700 and 1700 with a clear mid-afternoon peak.</p>
</div>
<div class = "col-md-8">
```{r}
tweets %>%
  filter(is_retweet == FALSE) %>%
  separate(col = created_at, into = c("date", "time"), sep = " ") %>%
  mutate(time = parse_time(time),
         hour = hour(time)) %>% 
  group_by(hour) %>% 
  tally() %>% 
  ggplot(aes(x = hour, n)) +
  geom_col(fill = "#bdbdbd") +
  scale_y_continuous(expand = c(0.005, 0.005)) +
  labs(x = "Hour", y = NULL,
       title = "Twitter activity",
       subtitle = paste0("@TraffordCouncil tweets, ", format(min(tweets$created_at), "%d %b"), " - ", format(max(tweets$created_at),"%d %b %Y"))) +
  theme_x() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line())
```
</div>
</div>

<br>

<div class = "row">
<div class = "col-md-4">
<span style="font-weight:bold;">Word frequency</span>
<p style="word-wrap:break-word;">The most frequent terms tweeted from the <a href="https://twitter.com/traffordcouncil" target="_blank">@TraffordCouncil</a> Twitter account were "people", "day" and "free".</p>
</div>
<div class = "col-md-8">
```{r}
tidy_tweets %>%
  filter(!str_detect(word, "#\\S+"), # remove hashtags
         !str_detect(word, "@\\S+"), # remove @user
         !str_detect(word, regex("trafford", ignore_case = TRUE))) %>% 
  count(word, sort = TRUE) %>%
  top_n(10) %>%
  ggplot(aes(x = reorder(word, n), y = n)) +
  geom_col(fill = "#fb9a99") +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  scale_y_continuous(expand = c(0.005, 0.005)) +
  labs(x = NULL, y = NULL,
       title = "Most common words",
       subtitle = paste0("@TraffordCouncil tweets, ", format(min(tweets$created_at), "%d %b"), " - ", format(max(tweets$created_at),"%d %b %Y"))) +
  coord_flip() +
  theme_x()
```
</div>
</div>

<br>

<div class = "row">
<div class = "col-md-4">
<span style="font-weight:bold;">Most retweets</span>
<p style="word-wrap:break-word;">The most retweeted tweet from the <a href="https://twitter.com/traffordcouncil" target="_blank">@TraffordCouncil</a> Twitter account was about the loss of a wedding ring in a ballot box. This tweet was retweeted `r pull(top_n(arrange(filter(tweets, !is_retweet), desc(retweet_count)), 1, retweet_count), retweet_count)
` times. This tweet also received the most likes (`r pull(top_n(arrange(filter(tweets, !is_retweet), desc(favorite_count)), 1, favorite_count), favorite_count)
`).</p> 
</div>
<div class = "col-md-8">
<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">We've found a wedding ring in one of our ballot boxes! It was found in the Cloverlea Primary School ballot box, Hale Barns. If it's yours please call Access Trafford on 0161 912 2000 so we can reunite you with it!</p>&mdash; Trafford Council (@TraffordCouncil) <a href="https://twitter.com/TraffordCouncil/status/1205277494619844610">December 13, 2019</a></blockquote>
  <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>
</div>
</div>

<br>

<div class = "row">
<div class = "col-md-4">
<span style="font-weight:bold;">Frequency of user mentions</span>
<p style="word-wrap:break-word;"><a href="https://twitter.com/OneTrafford" target="_blank">@OneTrafford</a> was the most frequently mentioned Twitter account (28) by <a href="https://twitter.com/TraffordCouncil" target="_blank">@TraffordCouncil</a>.</p>
</div>
<div class = "col-md-8">
```{r}
tweets %>% 
  mutate(text = str_remove_all(text, "&amp;|&lt;|&gt;")) %>% 
  unnest_tokens(word, text, token = "tweets", to_lower = FALSE) %>% 
  filter(!word %in% stop_words$word, # remove stop words
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]"),
         str_detect(word, "@\\S+"),
         !str_detect(word, regex("@traffordcouncil", ignore_case = TRUE))) %>% 
  count(word, sort = TRUE) %>%
  top_n(10) %>%
  ggplot(aes(x = reorder(word, n), y = n)) +
  geom_col(fill = "#fdbf6f") +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  scale_y_continuous(expand = c(0.005, 0.005)) +
  labs(x = NULL, y = NULL,
       title = "Most common accounts mentioned in tweets",
       subtitle = paste0("@TraffordCouncil tweets, ", format(min(tweets$created_at), "%d %b"), " - ", format(max(tweets$created_at),"%d %b %Y"))) +
  coord_flip() +
  theme_x()
```
</div>
</div>

<br>

<div class = "row">
<div class = "col-md-4">
<span style="font-weight:bold;">Likes by @TraffordCouncil</span>
<p style="word-wrap:break-word;">The Twitter account whose tweets were most frequently liked by <a href="https://twitter.com/TraffordCouncil" target="_blank">@TraffordCouncil</a> was <a href="https://twitter.com/OneTrafford" target="_blank">@OneTrafford</a> (50). You can explore all the tweets liked by @TraffordCouncil in an <a href="likes.html" target="_blank">interactive table</a>.</p>
</div>
<div class = "col-md-8">
```{r}
likes %>% 
  count(screen_name, sort = TRUE) %>%
  top_n(10) %>%
  mutate(screen_name = paste0("@", screen_name)) %>% 
  ggplot(aes(x = reorder(screen_name, n), y = n)) +
  geom_col(fill = "#cab2d6") +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  scale_y_continuous(expand = c(0.005, 0.005)) +
  labs(x = NULL, y = NULL,
       title = "Frequency of tweets liked by account",
       subtitle = paste0("@TraffordCouncil account, ", format(min(likes$created_at), "%d %b %Y"), " - ", format(max(likes$created_at),"%d %b %Y"))) +
  coord_flip() +
  guides(fill = FALSE) +
  theme_x()
```
</div>
</div>

<br><br>

<p><span style="font-weight:bold;">Credits:</span> this analysis used the <a href="https://cran.r-project.org/web/packages/rtweet" target="_blank">rtweet</a> and <a href="https://cran.r-project.org/web/packages/tidytext" target="_blank">tidytext</a> <a href="https://cran.r-project.org" target="_blank">R</a> packages.</p>