---
title: "Comparing Population Estimates for Trafford"
lang: "en-GB"
output:
  html_document:
    highlight: null
    mathjax: null
    theme: flatly
    css: styles.css
    self_contained: TRUE
    toc: false
    toc_depth: 2
    toc_float: true
    df_print: paged
    includes:
      in_header: head_includes.html
---


<main>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, results=FALSE)
library(tidyverse) ;
library(httr) ; 
library(readxl) ;
library(scales) ; 
library(ggiraph) ;


source("https://github.com/traffordDataLab/assets/raw/master/theme/ggplot2/theme_lab.R")

`%out%` <- function(a,b) ! a %in% b

girafe_options <- list(opts_tooltip(use_fill = FALSE, opacity = 1, css = "background-color: #e7e6e1; color: #212121; padding: 0.1em; border-radius: 0.25em;"),
                      opts_toolbar(saveaspng = FALSE),
                      opts_hover(css = "fill-opacity: 1; stroke:white; stroke-opacity: 1; r: 2.5pt;")
                      )

```

```{r}

mid2022 <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_2002_1.data.csv?geography=1853882369,1807745054,2092957699&date=latest&gender=1,2&c_age=200,1,3...18,186...191&measures=20100") %>%
  select(area_name = GEOGRAPHY_NAME, year = DATE_NAME, gender = GENDER_NAME, ageband = C_AGE_NAME, value = OBS_VALUE)

mid2022Trafford <- mid2022 %>%
  filter(area_name== "Trafford", ageband != "All Ages") %>%
  mutate(ageband = ifelse(ageband %in% c("Age 85", "Age 86", "Age 87", "Age 88", "Age 89"), "Aged 85-89", ageband)) %>%
  group_by(area_name,year,gender, ageband) %>%
  summarise(value = sum(value)) %>%
  ungroup() %>%
  mutate(ageband = str_remove(ageband, "Age |Aged ")) %>%
  mutate(ageband = str_replace(ageband, " - ", "-")) %>%
  mutate(ageband = factor(ageband, levels = c("0-4","5-9", "10-14", "15-19", "20-24", "25-29", "30-34","35-39","40-44","45-49","50-54","55-59","60-64","65-69","70-74","75-79","80-84","85-89","90+"))) %>%
  mutate(gender = factor(gender, levels = c("Male", "Female")))

census2021 <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_2029_1.data.csv?date=latest&geography=1778385132&c_sex=1,2&c2021_age_92=1...91&measures=20100") %>%
  select(area_name = GEOGRAPHY_NAME, year = DATE_NAME, gender = C_SEX_NAME, age = C2021_AGE_92_SORTORDER, value = OBS_VALUE) %>%
  mutate(ageband = cut(age,
                       breaks = c(0,5,10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90,120),
                       labels = c("0-4","5-9","10-14","15-19","20-24","25-29","30-34","35-39",
                                  "40-44","45-49","50-54","55-59","60-64","65-69","70-74",
                                  "75-79","80-84","85-89","90+"),
                       right = FALSE)) %>%
    group_by(area_name,year,gender, ageband) %>%
  summarise(value = sum(value)) %>%
  ungroup() %>%
  mutate(ageband = factor(ageband, levels = c("0-4","5-9", "10-14", "15-19", "20-24", "25-29", "30-34","35-39","40-44","45-49","50-54","55-59","60-64","65-69","70-74","75-79","80-84","85-89","90+"))) %>%
  mutate(gender = factor(gender, levels = c("Male", "Female")))

```


## Mid Year 2022

</b>

<p>ONS has published the Mid-Year 2022 population estimates for England and Wales including local authority levels.</p>

<p>Trafford population was 236,301, 715 more than the population in Mid-Year 2021 and 1,249 more than the population reported by the Census 2021.</p>

<p>From the Census 2021 to Mid-Year 2022 the population has grow 0.53%, this is 1% less than GM at 1.53% grow and 0.56% less than England at 1.09%.</p>

<p>From Mid-Year 2021 to Mid-Year 2022 the population has grow 0.3%, this is 1.16% less than GM at 1.47% grow and 0.67% less than England at 0.97%. Trafford has the lowest growth compared to the rest of GM authorities.</p>

<p>The median age in Trafford was 41 years, 0.2 years more than the median age in Mid-Year 2021. Trafford had the third highest median age among GM authorities. GM median age is 37.3 years and England’s median age is 40.5 years.</p>

<div class = "row">
<div class = "col-md-2">
</div>
<div class = "col-md-8">
```{r}

temp <- bind_rows(mid2022Trafford, census2021) %>%
  mutate(value = case_when(gender == "Male" ~ value * -1, TRUE ~ value))

ggplot() +
  geom_col(data = filter(temp, year == "2022"), 
           aes(x = ageband, y = value, fill = gender), alpha = 0.8)+ 
  geom_line(data = filter(temp, year == "2021"), 
            aes(x = ageband, y = value, group = gender, colour = gender), stat = "identity", size = 1) +
  scale_fill_manual(values = c("#1d85a5", "#194a81")) +
  scale_colour_manual(values = c("#1d85a5", "#194a81")) +
  facet_share(~gender, dir = "h", scales = "free",reverse_num = TRUE) +
  coord_flip() +
  labs(x = NULL, y = NULL, 
       title = "Trafford population pyramid, Census 2021 to Mid year 2022",
       caption = "Census 2021 (line), Mid-year 2022 (bar)\nSource: ONS | @traffordDataLab") +
  theme(
        plot.title = element_text(size=16, face = "bold", hjust = 0),
        plot.subtitle = element_markdown(size=14, hjust = 0),
      plot.caption = element_text(size = 8, hjust = 1, margin = margin(t = 15)),
        plot.title.position = "plot",
        #plot.margin = unit(c(0,0,0,0), "cm"),
        #panel.grid.major.y = element_blank(),
        axis.text.y = element_text(hjust = 1, size=8),
        axis.text.x = element_text(size=8),
        axis.line.y = element_line(colour = "#000000"),
        
              plot.background = element_blank(),
      panel.background = element_blank(),
      panel.border = element_blank(),
            #panel.grid.major.x = element_blank(),
      panel.grid.major.x = element_line(size = 0.3, color = "#cbcbcb"),
      #panel.grid.minor = element_blank(),
      legend.position = "none",
            axis.ticks.x=element_line(size=0.1, color = "transparent"),
      axis.ticks.y=element_line(size=0.1, color = "transparent"),
      axis.line = element_blank(),
      strip.background = element_rect(fill = "transparent", colour = NA),
      strip.text = element_text(size = 9, face = "bold", hjust = .5))

```
</div>
<div class = "col-md-2">
</div>
</div>

<p>From Mid-Year 2021 to Mid-Year 2022 there was an estimate of 2,274 births and 2,103 deaths, with a surplus of births over deaths of 171. The Internal Migration inflow was 13,629 and the Internal Migration outflow was 14,870, with a net Internal Migration of -1,241. The International Migration inflow was 3,048 and the International Migration outflow was 1,261, with a net international migration of 1,787. Other changes account for -2.</p>

<p>Internal migration outflow is larger than the internal migration inflow for Trafford, England and the rest of the GM authorities except for Salford and Wigan which had a large internal migration inflow. Net International migration is the largest component of population growth for Trafford and all GM authorities, and England.</p>

<p>Births minus deaths or natural change has decline over the years. Net Internal Migration have gone from adding over 1,000 residents in 2013 to deducting residents from 2019, reaching -1,241 in 2022. Net International Migration has notably increased particularly in 2022 after negative numbers in 2019 and 2020.</p>

```{r}

tmp <- tempfile(fileext = ".xlsx")
 
 
GET(url = "https://www.ons.gov.uk/file?uri=/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/estimatesofthepopulationforenglandandwales/mid20222023localauthorityboundaires/mye22tablesew2023geogs.xlsx",
    write_disk(tmp))
 
mid2022 <- read_xlsx(tmp, sheet = 11, skip = 7) %>%
  filter(Code %in% c("E92000001", "E11000001", "E08000009") | Name %in% c("Bolton","Bury","Manchester","Oldham","Rochdale","Salford","Stockport","Tameside","Trafford","Wigan")) %>%
  select(Code,Name, `Mid-2022`, `Mid-2021`) %>%
  mutate(popDiff = `Mid-2022`-`Mid-2021`, percChange = (`Mid-2022`-`Mid-2021`)/`Mid-2021`*100)
 
mid2022_comp <- read_xlsx(tmp, sheet = 10, skip = 7) %>%
  filter(Code %in% c("E92000001", "E11000001", "E08000009") | Name %in% c("Bolton","Bury","Manchester","Oldham","Rochdale","Salford","Stockport","Tameside","Trafford","Wigan"))
 
mid2022_midAge <- read_xlsx(tmp, sheet = 13, skip = 7) %>%
  filter(Code %in% c("E92000001", "E11000001", "E08000009") | Name %in% c("Bolton","Bury","Manchester","Oldham","Rochdale","Salford","Stockport","Tameside","Trafford","Wigan")) %>%
  select(Code,Name, `Mid-2022`, `Mid-2021`) %>%
  mutate(Diff = `Mid-2022`-`Mid-2021`, percChange = (`Mid-2022`-`Mid-2021`)/`Mid-2021`*100)
 
tmp2 <- tempfile(fileext = ".xlsx")
 
GET(url = "https://www.ons.gov.uk/file?uri=/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/estimatesofthepopulationforenglandandwales/mid2011tomid2022detailedtimeseries/myebtablesenglandwales20112022v2.xlsx",
    write_disk(tmp2))
 
mid2022b <- read_xlsx(tmp2, sheet = 10, skip = 1) %>%
  filter(ladcode23 %in% c("E92000001", "E11000001", "E08000009") | laname23 %in% c("Bolton","Bury","Manchester","Oldham","Rochdale","Salford","Stockport","Tameside","Trafford","Wigan"))%>%
  select(area_code = ladcode23, area_name = laname23, contains(c("2021","2022"))) %>%
  pivot_longer(-c(area_code,area_name),names_to = c("measure","year"), names_sep = -4) %>%
  pivot_wider(names_from = year, values_from = value)%>%
  mutate(Diff = `2022`-`2021`, percChange = (`2022`-`2021`)/`2021`*100) %>%
  mutate(measure2 = ifelse(measure == "natchange_","natchange_net_",measure))
 
mid2022c <- mid2022_comp %>%
  select(Name, `Births minus Deaths`, `Internal Migration Net`,`International Migration Net`, Other ) %>%
  mutate(sum = rowSums(across(where(is.numeric))))

mid2022d <- read_xlsx(tmp2, sheet = 10, skip = 1) %>%
  filter(ladcode23 %in% c("E08000009"))%>%
  select(area_code = ladcode23, area_name = laname23, -Country, everything(), -Country) %>%
  pivot_longer(-c(area_code,area_name),names_to = c("measure","year"), names_sep = -4) 
```

<div class = "row">
<div class = "col-md-2">
</div>
<div class = "col-md-8">
```{r}

ggplot(mid2022d %>% filter(measure %in% c("natchange_", "internal_net_", "international_net_"))) +
  geom_hline(yintercept = 0, size = 1, colour = "#cbcbcb") +
  geom_line(aes(x =year, y= value, color = measure, group = measure),size = 1) +
  geom_point(aes(x =year, y= value, fill = measure, group = measure), shape = 21, color = "transparent", size = 2) +
  scale_colour_manual(values = c("natchange_" = "#A577A7",
                                 "internal_net_" = "#1d85a5",                                  "international_net_" = "#194a81"),
                      labels=c("Internal Migration Net", "International Migration Net", "Briths minus Deaths")
  ) +
  scale_fill_manual(values = c("natchange_" = "#A577A7", 
                               "internal_net_" = "#1d85a5", 
                               "international_net_" = "#194a81"),
                    labels=c("Trafford", "Similar Local Authorities*", "England")) +

  labs(
    title = "Main components of population change",
    subtitle = "Trafford",
    caption = "Source: ONS | @trafforddatalab",
    x = NULL,
    y = NULL,
    colour = NULL
  ) +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0),
        plot.subtitle = element_text(size = 14, hjust = 0),
        axis.text.x = element_text(angle = 90, margin = margin(r = 0)),
        legend.position = "top",
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(size = 0.5, color = "#cbcbcb"), 
        panel.grid.minor = element_blank()
  ) +
  guides(fill = "none")



```
</div>
<div class = "col-md-2">
</div>
</div>


```{css}
.smtable {
padding-left: 0px;
}
.TableCaption {
font-size: 10px;
}

#popTable td {
padding: 2px 2px;
}

.rightPar {
text-align:right;
color: grey;
font-size: 14px;
}
```
